### 2019-08-21 Analysis.Rmd
Added lots

---
title: "Single Cell Xenopus TMT Analysis"
author: "Kyle Kovary"
date: "5/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RColorBrewer)
library(pheatmap)
library(GGally)
library(VIM)
library(org.Xl.eg.db)
library(GO.db)
library(ggrepel)

# Import Data
files = list.files('TMTdata/', pattern = 'TargetPeptideSpectrumMatch', full.names = T)

for(i in seq_along(files)){
  if(i == 1){
    rm(pepRaw)
  }
  if(!exists('pepRaw')){
    pepRaw = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
  } else{
    temp = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
    colnames(temp) = colnames(pepRaw)
    pepRaw = rbind(pepRaw, temp)
    rm(temp)
  }
}
pepRaw = pepRaw %>% dplyr::select(file, massID, everything()) %>% filter(!is.na(`126`), 
                                                                  `Quan Usage` == 'Use',
                                                                  Confidence != 'Low')

```

## Data Organization

```{r }
pepRaw = pepRaw %>% 
  mutate(run = c('ctrl','a','b','c','d','e')[match(pepRaw$file, files)]) %>% 
  group_by(file, `Annotated Sequence`) %>% 
  mutate(pepID = letters[1:n()]) %>% ungroup() %>%
  dplyr::select(file, run, pepID, everything())

pepTidy = pepRaw %>% 
  gather(key = label, value = intensity, '126':'131') %>% 
  dplyr::select(run, label, intensity, everything(), -file) %>% 
  mutate(condition = ifelse(run == 'ctrl', 'ctrl',c('ctrl',0,20,40,60,80)[match(label,c(126:131))]),
         intensity = as.numeric(intensity)) %>% unite(peptide, c(`Annotated Sequence`,'pepID')) %>%
  dplyr::rename(Uniprot = `Master Protein Accessions`) %>% 
  unite(cell, c('run','label','condition'), remove = F) %>%
  dplyr::select(cell, run, label, condition, peptide, Uniprot, intensity)
```

## Variance and Co-variance Between Cells Before Normalization

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-20,20)
```

The peptide intensities appear to follow a log-normal distribution.

```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
pairs = pepTidy %>% filter(condition == 'ctrl') %>% 
  dplyr::select(cell, peptide, intensity) %>% spread(cell, intensity)

ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
```

```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
meanSD = pepTidy %>% filter(condition != 'ctrl') %>% 
  group_by(peptide) %>% 
  summarise(mean = mean(intensity, na.rm = T), 
            sd = sd(intensity, na.rm = T) / mean)

ggplot(meanSD, aes(x = log10(mean), y = sd)) + geom_point() + geom_smooth(method = lm, formula = y ~ poly(x, 2))
```

## Variance and Co-variance Between Cells After Cell Normalization

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
pepTidy = pepTidy %>% group_by(cell) %>% mutate(intensity = intensity / median(intensity, na.rm = T))

ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-8,8)
```

```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
pairs = pepTidy %>% filter(condition == 'ctrl') %>% 
  dplyr::select(cell, peptide, intensity) %>% spread(cell, intensity)

ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
```

## Variance and Co-variance Between Cells After Peptide Batch Normalization

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
pepTidy = pepTidy %>% group_by(run, peptide) %>% 
  mutate(intensity = intensity / median(intensity, na.rm = T)) %>% ungroup()

ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-8,8)
```

```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
pairs = pepTidy %>% dplyr::filter(condition == 'ctrl') %>% 
  dplyr::select(cell, peptide, intensity) %>% spread(cell, intensity)

ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
```

There are

## Variance and Co-variance Between Cells After Removing Peptides Outside of 2*sd

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
pepTidy %>% dplyr::filter(abs(log2(intensity)) < 2*sd(log2(pepTidy$intensity), na.rm = T))

ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-4,4)
```

The standard deviation of all peptides across eggs was calculated, and every peptide that had a value outside of `r 2*sd(log2(pepTidy$intensity), na.rm = T)`

## Group into proteins based on median

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
protTidy = pepTidy %>% filter(abs(log2(intensity)) < 2*sd(log2(pepTidy$intensity), na.rm = T)) %>% group_by(cell, run, label, condition, Uniprot) %>% 
  summarise(abundance = median(intensity, na.rm = T)) %>% ungroup()

pairs = protTidy %>% dplyr::filter(condition == 'ctrl') %>% 
  dplyr::select(cell, Uniprot, abundance) %>% 
  spread(cell, abundance)

ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()

pairs = protTidy %>% dplyr::select(cell, Uniprot, abundance) %>% spread(cell, abundance)

protImpute = kNN(data = pairs, k = 1) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)

cor = cor(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
colnames(cor) = protImpute$Uniprot
rownames(cor) = protImpute$Uniprot

mypal = rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201))
k = NA
map = pheatmap(cor, show_colnames = F, show_rownames = F, color = mypal, breaks = seq(-1,1, by = 0.01), cutree_cols = k, cutree_rows = k, legend = F)
map
tibble(Uniprot = protImpute$Uniprot, Cluster = cutree(map$tree_row, k = k)) %>% write_csv('/Users/kylekovary/Downloads/Map.csv')

remove = which(is.na(pairs), arr.ind = T)[,1] %>% unique()
pca = prcomp(protImpute[,2:ncol(protImpute)])

pcaDF = tibble(cell = rownames(pca$rotation), PC1 = pca$rotation[,1], PC2 = pca$rotation[,2])
pcaDF = pcaDF %>% separate(cell, into = c('run','label','condition'), remove = F) %>% print()
ggplot(pcaDF, aes(x = PC1, y = PC2, colour = condition)) + geom_point(size = 5)

```

## Map GO Terms to Proteins

```{r eval=F}
columns(org.Xl.eg.db)
mapIds(org.Xl.eg.db, keys = 'A0JPG5', keytype = 'UNIPROT', column = 'GOALL', multiVals = 'list')

keys <- keys(org.Xl.eg.db, keytype = 'UNIPROT')

GO <- protImpute %>% dplyr::select(Uniprot) %>% filter(Uniprot %in% keys) %>% 
  rowwise() %>% 
  mutate(GO = mapIds(org.Xl.eg.db, 
                     keys = Uniprot, 
                     keytype = 'UNIPROT', 
                     column = 'GOALL', 
                     multiVals = 'list')) %>% 
  ungroup()

allGO <- tibble(GOterms = unlist(GO$GO) %>% unique())
allGO <- allGO %>% group_by(GOterms) %>% mutate(Uniprot = GO[grep(GOterms, GO$GO),1] %>% as.list()) %>% ungroup()

write_csv(mutate(GO, GO = paste(unlist(GO), sep = '', collapse = ',')), '/Users/kylekovary/Documents/GitHub/Proteome-Expression-Covariation/analysisData/GO.csv')

write_csv(mutate(allGO, Uniprot = paste(unlist(Uniprot), sep = '', collapse = ',')), '/Users/kylekovary/Documents/GitHub/Proteome-Expression-Covariation/analysisData/allGO.csv')

```

## GO Term Analysis
```{r echo=TRUE, message=FALSE, warning=FALSE}
GO <- read_csv('/Users/kylekovary/Documents/GitHub/Proteome-Expression-Covariation/analysisData/GO.csv') %>% mutate(GO = strsplit(GO, split = ','))

allGO <- read_csv('/Users/kylekovary/Documents/GitHub/Proteome-Expression-Covariation/analysisData/allGO.csv') %>% 
  mutate(Uniprot = strsplit(Uniprot, split = ','),
         GO = mapIds(GO.db, 
                     keys = GOterms, 
                     keytype = 'GOID', 
                     column = 'TERM') %>% as.character(),
         ontology = mapIds(GO.db, 
                     keys = GOterms, 
                     keytype = 'GOID', 
                     column = 'ONTOLOGY') %>% as.character()) %>%
  dplyr::select(GOterms, GO, ontology, everything())

corVector <- function(uniprotList, corMat){
  targets = uniprotList %>% unlist()
  temp = corMat[rownames(corMat) %in% targets,colnames(corMat) %in% targets]
  temp[lower.tri(temp, diag = F)]
}

allGO <- allGO %>% group_by(GOterms) %>% dplyr::mutate(cors = corVector(Uniprot, cor) %>% as_tibble() %>% as.list()) %>% ungroup()

x <- allGO %>% rowwise %>% mutate(size = length(Uniprot)) %>% dplyr::select(GOterms, size, cors) %>% unnest() %>% group_by(GOterms) %>% 
  summarise(median = median(cors, na.rm = T),
            sd = sd(cors, na.rm = T),
            size = mean(size)) %>% 
  mutate(GOterms = fct_reorder(GOterms, median))

ggplot(filter(x, size >= 2), aes(x = GOterms, y = median)) + geom_point(alpha = 0.25) +
  geom_hline(yintercept = 0.02154303, colour = 'red', linetype = 'dashed') + 
  theme_bw() + ylab('Mean Correlation') + ylim(-1,1)

```

## Plot highest correlated GO Terms with GO Term Names
```{r}
topGOnames <- x %>% filter(size > 2) %>% .[order(.$median, decreasing = T),] %>% 
  .[1:50,'GOterms'] %>% unlist()

topGO <- allGO %>% filter(GOterms %in% topGOnames) %>% 
  dplyr::select(GOterms, GO, cors) %>% unnest() %>%
  group_by(GO) %>% mutate(median = median(cors)) %>% ungroup() %>%
  mutate(GO = fct_reorder(GO, median))

ggplot(topGO, aes(x = GO, y = cors)) + geom_boxplot() + geom_jitter(alpha = 0.5, size = 0.5) + coord_flip() + scale_x_discrete(labels = function(x) str_wrap(x, width = 50)) + theme_bw() + ylim(-1,1)

mitoProts <- allGO %>% filter(GOterms == 'GO:0051321') %>% dplyr::select(Uniprot) %>% unnest() %>% unlist()
mitoDF <- protTidy %>% filter(condition != 'ctrl', Uniprot %in% mitoProts) %>% group_by(condition, Uniprot) %>% summarise(mean = mean(abundance), sd = sd(abundance) / (n()-1))

ggplot(mitoDF, aes(x = as.numeric(condition), y = mean, colour = Uniprot)) + geom_point() + geom_path() + geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0, alpha = 0.5) + theme_bw()

```

## Variance of GO Terms
```{r}

protVar <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
  separate(cell, into = c('run','label','condition'), remove = F) %>% 
  filter(condition != 'ctrl') %>% group_by(Uniprot) %>%
  summarise(cv = sd(abundance)/mean(abundance))

ggplot(protVar, aes(x = cv)) + geom_histogram(bins = 50) + theme_bw()

selectCV <- function(x){
  protVar %>% filter(Uniprot %in% unlist(x)) %>% 
    dplyr::select(cv) %>% as_tibble() %>% as.list()
}

allGO <- allGO %>% group_by(GOterms) %>% 
  dplyr::mutate(cv = selectCV(Uniprot)) %>% ungroup()

meanCV = allGO %>% dplyr::select(GOterms, cv) %>% unnest() %>% group_by(GOterms) %>% 
  summarise(mean = mean(cv), size = n()) %>% mutate(GOterms = fct_reorder(GOterms, mean))

ggplot(meanCV, aes(x = GOterms, y = mean, size = log2(size), alpha = 0.2)) + 
  geom_point() + theme_bw()

topGOnames <- meanCV %>% filter(size > 2) %>% .[order(.$mean, decreasing = F),] %>% 
  .[1:40,'GOterms'] %>% unlist()

topGO <- unnest(dplyr::select(filter(allGO, GOterms %in% topGOnames), GOterms, cv)) %>%
  mutate(GO = mapIds(GO.db, keys = GOterms, keytype = 'GOID', column = 'TERM')) %>%
  group_by(GO) %>% mutate(median = median(cv)) %>% ungroup() %>%
  mutate(GO = fct_reorder(GO, median))

ggplot(topGO, aes(x = GO, y = cv)) + geom_boxplot(outlier.size = 1) + geom_point() + coord_flip() + scale_x_discrete(labels = function(x) str_wrap(x, width = 50)) + theme_bw() + ylim(0,0.3)
```

```{r message=FALSE, warning=FALSE}
allGO[grep('ribosom',allGO$GO),]
glyc <- filter(allGO, GOterms == 'GO:0005840')$Uniprot %>% unlist()

glycMat <- cor[rownames(cor) %in% glyc,colnames(cor) %in% glyc]
rownames(glycMat) <- mapIds(org.Xl.eg.db, keys = rownames(glycMat), keytype = 'UNIPROT', column = 'SYMBOL')
pheatmap(glycMat, show_colnames = T, show_rownames = T, color = mypal, breaks = seq(-1,1, by = 0.01),
         cutree_cols = 3, cutree_rows = 3)

glyc <- protImpute %>% filter(Uniprot %in% glyc) %>% 
  mutate(symbol = mapIds(org.Xl.eg.db, keys = Uniprot, keytype = 'UNIPROT', column = 'SYMBOL')) %>%
  dplyr::select(Uniprot, symbol, everything())

glyc <- glyc %>% gather(key = sample, value = abundance, 3:ncol(glyc)) %>%
  dplyr::select(sample, symbol, abundance) %>% spread(key = symbol, value = abundance)

ggpairs(log2(glyc[,2:ncol(glyc),]))
```

## Cor vs CV
```{r}
test = allGO %>% rowwise() %>% mutate(n = length(Uniprot), medCors = mean(cors), medCV = mean(cv)) %>%
  filter(length(Uniprot) > 3, !is.na(medCors), !is.na(medCV))

ggplot(test, aes(x = medCors, y = medCV)) + geom_point(alpha = 0.5) + theme_bw()

ggplot(test, aes(x = medCors, y = medCV)) + 
  geom_point(alpha = 0.5) + 
  #stat_smooth(method="lm",formula=  y ~ I(1/x)) +
  geom_text_repel(data = filter(test, medCors > 0.15), aes(label = GO), colour = 'black', size = 3) + 
  theme_bw()

test2 = tibble(Cor = test$medCors, CV = test$medCV, mult = test$medCors * test$medCV) %>%
  gather(key = 'Key', value = 'Value')
ggplot(test2, aes(x = Value, colour = Key)) + geom_density() + theme_bw()
ggplot(test2, aes(x = Key, y = Value)) + geom_boxplot() + theme_bw()
# Variance vs sum of covariance
# Cor vs Var relationship for every protein pair... error propagation
```

## All combinations of 5 proteins -> total error vs correlation
```{r}
combMat <- combn(1:length(cor))
corVect <- tibble(protA = rep(rownames(cor), times = length(colnames(cor))), 
                  protB = rep(colnames(cor), each = length(rownames(cor))),
                  keep = as.vector(lower.tri(cor)),
                  cor = as.vector(cor)) %>% filter(keep) %>% dplyr::select(-keep)

corVect <- corVect %>% dplyr::rename(Uniprot = protA) %>% 
  left_join(., protVar, 'Uniprot') %>%
  dplyr::rename(protA = Uniprot, Uniprot = protB, cvA = cv) %>%
  left_join(., protVar, 'Uniprot') %>% dplyr::rename(protB = Uniprot, cvB = cv) %>%
  mutate(totVar = sqrt(cvA^2 + cvB^2))

ggplot(corVect, aes(x = cor, y = totVar)) + geom_hex(bins = 250) + theme_bw() +
  xlim(-1,1) + ylim(0,0.4) + 
  geom_vline(xintercept = 0.7, colour = 'red', linetype = 'dashed') +
  geom_hline(yintercept = 0.2, colour = 'red', linetype = 'dashed')

ggplot(filter(corVect, totVar > 0.2, cor > 0.7), aes(x = cor, y = totVar))  + 
  geom_point(alpha = 0.15, size = 0.1) + theme_bw() + xlim(-1,1) + ylim(0,0.4)
```

## Total Variance
```{r}
protImputeTidy <- protImpute %>% 
  gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
  separate(cell, into = c('run','label','condition'), remove = F) %>% 
  filter(condition != 'ctrl')

### Prototype function
## Combine the apply find correlation value and pulling out the CV value
geneList <- allGO$Uniprot[9] %>% unlist()
protCombs <- t(combn(geneList, 2))

groupVar <- function(x){
  geneList <- unlist(x)
  groupVars <- protVar %>% filter(Uniprot %in% geneList)
  return(sqrt(sum(groupVars$cv^2)))
}

# Fix by including new code from totVarGroup()
groupCor <- function(x){
  geneList <- unlist(x)

  if (length(geneList) > 1) {
    protCombs <- t(combn(geneList, 2))
    return(sqrt(sum(totVarPair(protCombs))))
  } else{
    return(as.numeric(protVar[protVar$Uniprot == geneList[1], 2]))
  }
}

totVarPair <- function(x){
  A <- as.character(x[1])
  B <- as.character(x[2])
  varA <- as.numeric(protVar[protVar$Uniprot == A, 2])
  varB <- as.numeric(protVar[protVar$Uniprot == B, 2])
  corAB <- cor[A,B]
  tot <- varA^2 + varB^2 + 2*corAB*varA*varB
  return(tot)
}

totVarGroup <- function(x){
  geneList <- unlist(x)

  if (length(geneList) > 1) {
    protCombs <- t(combn(geneList, 2))
    return(sqrt(sum(apply(protCombs, 1, function(x) totVarPair(x)))))
    #return(sqrt(sum(totVarPair(protCombs))))
  } else{
    return(as.numeric(protVar[protVar$Uniprot == geneList[1], 2]))
  }

}

allGO1 <- allGO %>% filter(!is.na(GOterms)) %>% rowwise() %>% 
  mutate(n = length(unlist(Uniprot)),
         totVar = totVarGroup(Uniprot),
         normVar = totVar / n) %>%
  filter(n > 3)

allGO1 %>% arrange(desc(normVar)) %>% print(n = 100)
allGO1 %>% arrange(normVar) %>% print(n = 100)

allGO1$GO <-factor(allGO1$GO, levels = allGO1$GO[order(allGO1$normVar)])

ggplot(allGO1, aes(x = normVar, y = totVar)) + geom_point()
ggplot(allGO1, aes(x = log2(n), y = log2(totVar))) + geom_point()
ggplot(allGO1, aes(x = GO, y = normVar)) + geom_point()

ggplot(allGO1, aes(x = normVar)) + geom_histogram(bins = 30) + 
  geom_vline(xintercept = c(0.15, 0.1), colour = 'red', linetype = 'dashed') +
  theme_bw()

```

### Abundance based on intensity
Does expression level explain lower variance?
```{r}

ggplot(protRawMean, aes(x = log2(mean), y = noise)) + geom_point() + theme_bw()

```

It doesn't appear to be the case. This could be because these are likely the top ~1000 highest expressed proteins that have a dynamic range of ~250X.

### Move up the GO branch to see how many high/low variance GO terms are from which parent branch
```{r}

```

### Is lower variance compensating for high correlations? How much of a role is negative correlation playing in reducing the total variance?
```{r}
dat <- left_join(ungroup(test), allGO1, by = "GOterms") %>%
  dplyr::select(GOterms, GO.x, ontology.x, Uniprot.x, cors.x, cv.x, n.x,medCors, medCV, normVar)

dat$score <- 'Med'
dat[dat$normVar > 0.145, 'score'] <- 'High'
dat[dat$normVar < 0.115, 'score'] <- 'Low'

ggplot(dat, aes(x = medCors, y = medCV, colour = score)) + geom_point() 
ggplot(dat, aes(x = medCors, y = medCV)) + geom_point() + geom_smooth(method = 'lm') + facet_grid(~score)

ggplot(filter(dat, n.x > 5), aes(x = medCors, y = medCV, colour = score)) + 
  geom_point(alpha = 0.75) + scale_color_manual(values=c("blue", "red", "grey"))
# + #scale_colour_gradient(low="red", high="blue") +
#  geom_text_repel(data = filter(dat, score != 'Med'), aes(label = GOterms), colour = 'black', size = 1)

ggplot(filter(dat, n.x > 5, score != 'Med', ontology.x != 'CC'), aes(x = medCors, y = medCV, colour = normVar)) + 
  geom_point() + scale_colour_gradient(low="red", high="blue") + 
  geom_smooth(method = 'lm') + facet_wrap(~score, scales = 'free') +
  geom_text_repel(data = filter(dat, n.x > 5, score != 'Med'), aes(label = GO.x), colour = 'black', size = 2)

plot_ly(filter(dat, n.x > 5), z = ~normVar, y = ~medCV, x = ~medCors, color = ~score, colors = c("red", "blue", "grey")) %>%
  add_markers() %>%
  layout(scene = list(zaxis = list(title = 'Normalized Variance'),
                     yaxis = list(title = 'Median CV'),
                     xaxis = list(title = 'Median Correlation')))

ggplot(filter(dat, n.x > 5), aes(y = normVar, x = medCV)) + geom_point(alpha = 0.75)
ggplot(filter(dat, n.x > 5), aes(y = normVar, x = medCors)) + geom_point(alpha = 0.75)

library(GGally)
p <- ggpairs(filter(dat, n.x > 5), columns = c('medCV', 'medCors','normVar'), 
        ggplot2::aes(colour = score, alpha = 0.5))

for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] + 
        scale_fill_manual(values=c("blue", "red", "grey")) +
        scale_color_manual(values=c("blue", "red", "grey"))  
  }
}
p

```

## Variance within highly correlated GO terms
```{r}
allGO[grep('large ribo',allGO$GO),]
hCor <- filter(allGO, GOterms == 'GO:0015934')$Uniprot %>% unlist()

hCor <- protImpute %>% filter(Uniprot %in% hCor) %>% 
  mutate(symbol = mapIds(org.Xl.eg.db, keys = Uniprot, keytype = 'UNIPROT', column = 'SYMBOL')) %>%
  dplyr::select(Uniprot, symbol, everything())

hCor <- hCor %>% gather(key = sample, value = abundance, 3:ncol(hCor)) %>%
  dplyr::select(sample, symbol, abundance) %>% spread(key = symbol, value = abundance) %>%
  gather(key = protein, value = abundance, 2:ncol(.))

hCorSD <- hCor %>% group_by(sample) %>% summarise(sd = sd(abundance))

ggplot(hCorSD, aes(x = sample, y = sd)) + geom_point() + geom_hline(yintercept = sd(hCor$abundance))
```

### 2019-08-21 Modeling.Rmd
Added lots

---
title: "Single Cell Xenopus TMT Analysis: Modeling"
author: "Kyle Kovary"
date: "5/15/2019"
output: html_document
---

## Complex Stoichiometry
Could be that when drawing from a correlated distribution, the variance between cells will be the same, but the variance within a cell will be much less... High correlation increases variance between cells but reduces variance within a cell for a given pathway. Need to test that high variance with high covariance matters for a single cell, could be that correlation is all that matters, which would invalidate the reasoning that for certain pathways there is a tradeoff between variance and correlation.
```{r}
correlatedValue = function(x, r){
  r2 = r**2
  ve = 1-r2
  SD = sqrt(ve)
  e  = rnorm(length(x), mean=0, sd=SD)
  y  = r*x + e
  return(y)
}

#set.seed(5)
n = 1000
x = rnorm(n)
y = correlatedValue(x=x, r=.6)

# multiply the variable by the SD you want and then add the mean you want

corMat <- function(nRow = 1000, nCol = 10, R = 0.5, mean = 1000, sd = 30){
  mat <- matrix(data = rnorm(nRow), nrow = nRow)

  for(i in 1:(nCol) - 1){
    mat <- cbind(mat, correlatedValue(x = mat[,i], r = R))
  }

  return((mat*sd) + mean)
}

```

```{r}
library(mvtnorm)

fun = function(cor_matrix, list_distributions, L)
{
    n = length(list_distributions)
    # Correlated Gaussian variables
    Gauss = rmvnorm(n=L, mean = rep(0,n), sig=cor_matrix)
    # convert them to uniform distribution.
    Unif = pnorm(Gauss) 
    # Convert them to whatever I want
    vars = sapply(1:n, FUN = function(i) list_distributions[[i]](Unif[,i]))
    return(vars)
}

# Create a matrix of values with a given R, mean, and standard deviation
corProts <- function(nProt, nCell, R, mean, sd) {

  cor_matrix =  matrix(data = R,
                       ncol = nProt,
                       nrow = nProt)
  for (i in 1:nProt) {
    cor_matrix[i, i] = 1
  }

  list_distributions <- list()
  for (i in 1:nProt) {
    list_distributions <-
      c(list_distributions, function(nCell)
        round(qnorm(nCell, mean, sd)))
  }

  mat = fun(cor_matrix, list_distributions, nCell)
  return(mat)
}

# Plot of total variance of 10 subunit complex

# Plot of max number of complete complexes
cors <- seq(0,1,0.02)
vars <- rev(seq(0,1,0.02))

corVarMat <- matrix(data = NA, ncol = length(cors), nrow = length(vars))
colnames(corVarMat) <- cors
rownames(corVarMat) <- vars

# Percent of potentially complete complexes
for(i in 1:length(cors)){
  for(j in 1:length(vars)){
    perc <- corProts(nProt = 10, nCell = 10000, R = cors[i], mean = 1000, sd = 1000*vars[j])
    perc <- mean(apply(perc, 1, function(x) min(abs(x)) / max(abs(x))), na.rm = T)
    corVarMat[j,i] <- perc
  }
}
pheatmap(corVarMat, cluster_cols = F, cluster_rows = F)
rotate <- function(x) t(apply(x, 2, rev))
pheatmap(rotate(corVarMat), cluster_cols = F, cluster_rows = F)
plot(diag(rotate(corVarMat)), rotate(corVarMat)[])

x <- tibble(index = 1:51,
            diagnal = diag(corVarMat), 
            zeroCor = corVarMat[,1],
            fullCor = corVarMat[1,])
x %>% gather(key = slice, value = percent, 2:4) %>% 
  ggplot(., aes(x = index, y = percent, colour = slice)) + 
  geom_point() + geom_vline(xintercept = 40, linetype = 'dashed')

x <- as_tibble(cbind(vars, corVarMat))
x <- x %>% gather(key = R, value = percent, 2:ncol(x)) %>% mutate(R = as.numeric(R)) %>%
  rename_(CV = vars)
ggplot(x, aes(x = CV, y = percent, colour = R, group = as.factor(R))) + geom_path() + scale_x_reverse() + theme_bw()

ggplot(x, aes(x = CV, y = percent, colour = as.factor(R))) + geom_path() + scale_x_reverse(limits = c(0.2,0)) + theme_bw()

# Within cell variance
for(i in 1:length(cors)){
  for(j in 1:length(vars)){
    stdev <- corProts(nProt = 10, nCell = 10000, R = cors[i], mean = 1000, sd = 1000*vars[j])
    stdev <- median(apply(stdev, 1, function(x) sd(x, na.rm = T) / mean(x, na.rm = T)), na.rm = T)
    corVarMat[j,i] <- stdev
  }
}

```

### 2019-08-21 Enriched GO Parents.R

@@ -1,0 +1,56 @@
+library(tidyverse)
+library(readxl)
+library(stringr)
+
+# Load in and combine data
+high <- read_xlsx(path = '/Users/kylekovary/Documents/GitHub/Proteome-Expression-Covariation/high_vs_low_var_GO.xlsx',
+                  sheet = 1) %>%
+  mutate(Var = 'high')
+low <- read_xlsx(path = '/Users/kylekovary/Documents/GitHub/Proteome-Expression-Covariation/high_vs_low_var_GO.xlsx',
+                 sheet = 2) %>%
+  mutate(Var = 'low')
+
+highLow <- full_join(high, low, by = c('Definitions')) %>% 
+  dplyr::select(Definitions, Counts.x, Counts.y) %>%
+  dplyr::rename(High_Var = Counts.x, Low_Var = Counts.y) %>%
+  gather(key = Var, value = Counts, High_Var:Low_Var)
+
+highLow[which(is.na(highLow$Counts)), 'Counts'] = 0
+
+# Tidy and calculate percentages
+highLow <- highLow %>% filter(!Definitions %in% c('biological_process',
+                                                 'molecular_function',
+                                                 'cellular_component',
+                                                 'cell',
+                                                 'cytoplasm',
+                                                 'intracellular')) %>% 
+  group_by(Var) %>% 
+  dplyr::mutate(per = Counts / sum(Counts)) %>%
+  ungroup() #%>% mutate(Definitions = fct_reorder(Definitions, per))
+
+# "Other" categories
+notOther <- highLow %>% group_by(Definitions) %>% summarise(sumPerc = sum(per)) %>%
+  filter(sumPerc > 0.04)
+notOther <- as.character(notOther$Definitions)
+
+other <- highLow %>% filter(!Definitions %in% notOther) %>% group_by(Var) %>%
+  summarise(Counts = sum(Counts), per = sum(per)) %>% mutate(Definitions = 'other') %>%
+  dplyr::select(Definitions, Var, Counts, per)
+
+highLow <- highLow %>% filter(Definitions %in% notOther)
+highLow <- rbind(highLow, other)
+highLow$Definitions = str_wrap(highLow$Definitions, width = 20)
+
+# Plots
+ggplot(highLow, aes(x = factor(1), y = per, fill = Definitions)) +
[…]

### 2019-08-21 peptideAnalysis.R

@@ -28,7 +28,8 @@ pepRaw = pepRaw %>% select(file, massID, everything()) %>% filter(!is.na(`126`),
 #################
 
 pepRaw = pepRaw %>% mutate(run = c('ctrl','a','b','c','d','e')[match(pepRaw$file, files)]) %>% 
-  select(file, run, everything())
+  group_by(file, `Annotated Sequence`) %>% mutate(pepID = letters[1:n()]) %>% ungroup() %>%
+  select(file, run, pepID, everything())
 
 pepTidy = pepRaw %>% gather(key = label, value = intensity, '126':'131') %>% 
   select(run, label, intensity, everything(), -file) %>% 
@@ -64,15 +65,81 @@ pepNorm = pepTidy %>% group_by(run, massID) %>%
   mutate(pepNorm = mean(pepNorm)) %>% filter(!duplicated(`Annotated Sequence`)) %>% ungroup()
 
 
+test = pepTidy %>% group_by(run, label) %>% 
+  ungroup() %>% group_by(`Annotated Sequence`, pepID, run) %>%
+  mutate(pepNorm = intensity / intensity[label == 126])
 
+test2 = filter(test, `Annotated Sequence` == 'aAAEGPmk', pepID == 'a') %>%
+  group_by(run) %>% mutate(norm = intensity / intensity[label == 126]) %>%
+  select(run, label, condition, pepID, intensity, norm)
 
-### Combine Peptides to Proteins ###
+model = lm(norm ~ run, data = test2)
+model = lm(intensity~run+label, data = test2)
+test2$resid = model$residuals
+ggplot(test2, aes(x = label, y = norm, colour = run)) + 
+  geom_point()
+ggplot(test2, aes(x = label, y = resid, colour = run)) + 
+  geom_point()
 
+testProt = test %>% group_by(run, condition, `Master Protein Accessions`) %>%
+  summarise(intensity = median(intensity, na.rm = T)) %>%
+  #group_by(run, `Master Protein Accessions`) %>% 
+  #mutate(intensity = intensity / intensity[condition == 'ctrl']) %>%
+  rename(Accession = `Master Protein Accessions`)
 
+#### Correlation between loading controls
[…]

### 2019-08-21 proteinAnalysis.R

@@ -1,0 +1,89 @@
+library(tidyverse)
+library(ggpubr)
+
+#############################
+#### Import Protein Data ####
+#############################
+
+files = list.files('TMTdata/', pattern = 'TargetProtein', full.names = T)
+
+for(i in seq_along(files)){
+  if(i == 1){
+    rm(protRaw)
+  }
+  if(!exists('protRaw')){
+    protRaw = read_tsv(file = files[i]) %>% mutate(file = files[i])
+  } else{
+    temp = read_tsv(file = files[i]) %>% mutate(file = files[i])
+    colnames(temp) = colnames(protRaw)
+    protRaw = rbind(protRaw, temp)
+    rm(temp)
+  }
+}
+protRaw = protRaw %>% select(file, everything())
+
+
+
+#########################
+### Tidy Protein Data ###
+#########################
+
+protRaw = protRaw %>% mutate(run = c('ctrl','a','b','c','d','e')[match(protRaw$file, files)]) %>% 
+  select(file, run, everything())
+
+protTidy = protRaw %>% gather(key = label, value = intensity, `Ratios: (127) / (126)` : `Ratios: (131) / (126)`) %>% 
+  select(run, label, intensity, everything(), -file) %>% 
+  mutate(condition = ifelse(run == 'ctrl', 
+                            paste0('ctrl',127:131)[match(as.character(label),paste0('Ratios: (',c(127:131), ') / (126)'))],
+                            c(0,20,40,60,80)[match(as.character(label),paste0('Ratios: (',c(127:131), ') / (126)'))]),
+         intensity = as.numeric(intensity)) %>%
+  select(run, label, condition, everything())
+
+
+###################################
+#### Protein Variance Analysis ####
+###################################
+
+runVar = protTidy %>% group_by(Accession, run) %>% summarise(cv = sd(intensity, na.rm = T) / mean(intensity, na.rm = T)) %>% mutate(group = 'run')
+condVar = protTidy %>% group_by(Accession, condition) %>% summarise(cv = sd(intensity, na.rm = T) / mean(intensity, na.rm = T)) %>% mutate(group = 'cond')
+
+allVar = full_join(runVar, condVar, by = 'Accession') %>% gather(key = group, value = A)
+
+ggplot(runVar, aes(cv, colour = run)) + geom_freqpoly(bins = 100) + theme_bw() + facet_grid(run~.)
[…]

### 2020-02-23 Analysis.Rmd
added RMarkdown manuscript

@@
 date: "5/15/2019"
-output: html_document
+output: pdf_document
 ---
@@
 library(ggrepel)
+library(plotly)
 
@@
 for(i in seq_along(files)){
-  if(i == 1){
+  if(i == 1 & exists('pepRaw')){
     rm(pepRaw)
-  }
-  if(!exists('pepRaw')){
+  } else if(!exists('pepRaw')){
     pepRaw = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
@@
 }
-pepRaw = pepRaw %>% dplyr::select(file, massID, everything()) %>% filter(!is.na(`126`), 
+pepRaw <- pepRaw %>% dplyr::select(file, massID, everything()) %>% filter(!is.na(`126`), 
                                                                   `Quan Usage` == 'Use',
@@
 mypal = rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201))
-k = NA
+k = 2
 map = pheatmap(cor, show_colnames = F, show_rownames = F, color = mypal, breaks = seq(-1,1, by = 0.01), cutree_cols = k, cutree_rows = k, legend = F)
 map
+
 tibble(Uniprot = protImpute$Uniprot, Cluster = cutree(map$tree_row, k = k)) %>% write_csv('/Users/kylekovary/Downloads/Map.csv')
@@
 columns(org.Xl.eg.db)
-mapIds(org.Xl.eg.db, keys = 'A0JPG5', keytype = 'UNIPROT', column = 'GOALL', multiVals = 'list')
+#mapIds(org.Xl.eg.db, keys = 'A0JPG5', keytype = 'UNIPROT', column = 'GOALL', multiVals = 'list')
 
@@
 
-ggpairs(log2(glyc[,2:ncol(glyc),]))
+#ggpairs(log2(glyc[,2:ncol(glyc),]))
 ```
@@
 ```{r}
-combMat <- combn(1:length(cor))
+combMat <- combn(1:length(colnames(cor)),2)
 corVect <- tibble(protA = rep(rownames(cor), times = length(colnames(cor))), 
@@
 ## Variance within highly correlated GO terms
-```{r}
+```{r message=FALSE, warning=FALSE}
 allGO[grep('large ribo',allGO$GO),]
@@
 ggplot(hCorSD, aes(x = sample, y = sd)) + geom_point() + geom_hline(yintercept = sd(hCor$abundance))
+
+BC_delta <- function(x) {
+  hCor <- filter(allGO, GOterms == x)$Uniprot %>% unlist()
+
+  hCor <- protImpute %>% filter(Uniprot %in% hCor) %>%
+    mutate(symbol = mapIds(
+      org.Xl.eg.db,
+      keys = Uniprot,
+      keytype = 'UNIPROT',
+      column = 'SYMBOL'
+    )) %>%
+    dplyr::select(Uniprot, symbol, everything())
+
+  hCor <- hCor %>% gather(key = sample, value = abundance, 3:ncol(hCor)) %>%
+    dplyr::select(sample, symbol, abundance)
+
+  hCorSD <-
+    hCor %>% group_by(sample) %>% summarise(sd = sd(abundance))
+
+  return(hCorSD %>% summarise(sum(sd < sd(hCor$abundance)) / n()) %>% as.numeric())
+}
+
+dat <- dat %>% rowwise() %>% mutate(bcDelta = BC_delta(GOterms))
+
+ggplot(dat, aes(x = medCors, y = bcDelta)) + geom_point(alpha = 0.5) + 
+  geom_text_repel(data = filter(test, medCors > 0.2), 
+                    aes(label = GO), colour = 'black', size = 2)
 ```

### 2020-02-23 Manuscript.Rmd
added RMarkdown manuscript

---
title: "Combined effects of protein expression variance and correlation on multicomponent systems"
author: "Kyle M. Kovary and Mary N. Teruel"
date: 'Updated: `r Sys.time()`'
output: 
  pdf_document:
    toc: true
    number_sections: false
    fig_width: 7
    fig_height: 6
    fig_caption: true
  geometry: margin=0.25in
header-includes:
  - \setlength{\parindent}{4em}
  - \setlength{\parskip}{0em}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Protein expression variation leads to phenotypic variance between cells. This has been demonstrated in cell signaling and differentiation decisions. Additionally coordinated expression of proteins between cells can tune signaling pathways either towards a more binary or analog modality. Though there is some evidence in bulk cell measurements and in bacteria that certain heteromeric subunits or metabolic pathways may be expressed in a coordinated fashion (i.e. operons in bacteria), there has not been a direct measurement of coordinated expression of proteins (independent of TFs) in metazoans (vertebrates?). Here, we measure cell-to-cell variability of relative protein abundance using quantitative proteomics of individual Xenopus laevis eggs and show that proteins involved in metabolic pathways or members of heteromeric complexes tend to have high correlations with other members of those pathways/complexes. Our previous work highlighted the fact that correlated expression increases the total variation of a pathway, so one would reason that certain pathways or complexes would need to compensate of this extra source of by reducing the variation of expression of these pathways or complexes. To test this we computed total variance score that took into account both the coefficient of variance and correlations between proteins in a pathway and found that the lower 10% of GO terms were highly enriched for metabolic pathways. When we looked at the relationship between CV and R between these GO terms we found a negative relationship between them, demonstrating that increased correlation needs to come at the expense of decreased variance. Simple molecular models of heteromeric complexes and metabolic pathways demonstrate that this tradeoff can result in higher efficiencies in both function and reduced energy waste. Together, our study argues for a control principle whereby coordinated expression of proteins in a pathway can require lower variance in order to reduce the overall pathway variance, enabling accurate control of active complexes or metabolic pathway activity.

# Introduction

The coordinated expression of proteins is vital across cellular function, from maintaining a dynamic steady state to differentiation of cells into specialized types in different tissues. The ability to regulate coordinated expression has been demonstrated to occur via transcription (transcription factors, chromatin regulation), translation (specialized ribosomes, mRNA structure/modifications), and degradation (E3 ligases). These processes, as well as the molecules they target, are all subject to noise, which has been shown to lead to differences in cell behavior and decisions in otherwise identical cells. Variability and coordinated expression (correlation) of proteins leads to increased population level control in binary decisions, and a decreased ability to execute analog signaling.

Despite these important insights, a global understanding of the variability of protein expression between single cells and the coordinated expression of groups of proteins is largely an nonexistent. To systematically assess these properties of single cells, we carried out a proteomics experiment on exceeding large single cells, Xenopus laevis eggs. An advantage of this approach is that we can get around much of the signal to noise issues that accompany studying single cells due to low sample amounts. Additionally, at this stage of development transcription is restricted so we are able to gain insights into non-transcriptional control of protein expression. We also utilized isobaric tagging in order to measure 25 individual eggs in 5 mass spectrometry runs. In this study, we were able to measure the relative abundance >1000 proteins across single cells in order to better understand the properties of stochastic and coordinated expression of proteins.

With this dataset we have been able to, for the first time, measured the relationship between protein expression variance and coordinated protein expression on a proteome scale in single cells. We have observed that certain classes of proteins, including protein complexes and metabolic pathways, are expressed in such a way that increased coordinated expression is balanced by decreased variation. By doing this, cells are able to decrease the total variance of a given complex or pathway, an elegant balancing act that allows for finer control of metabolic throughput and controlling the number of potentially formed complexes though stoichiometric control. Though this kind of coordinated expression leads to an increase in variance in a population of cells, it can reduce variation within a cell.

# Results

## Single cell proteomics reveals global protein expression variability and coordinated expression between protein pairs.

Activated Xenopus laevis eggs were collected at 5 time points across the first cell cycle (0, 20, 40, 60, and 80 minutes), with 5 eggs at each time point. Using TMT multiplexing and mass spectrometry, we were able to determine the relative abundance of more than 1300 proteins. Expression of these proteins were largely invariant across the cell cycle, revealing that these highly expressed genes are likely not regulated by cell cycle processes. Additionally, a PCA analysis of these eggs showed no discernible clustering on cell cycle time (Fig S1).

To determine the variance across these proteins, we calculated the coefficient of variation (CV) using all time points collected. This showed a wide range of variation containing multiple distributions (Fig 1B), many of which are consistent with our previous study of variation using targeted mass spectrometry (Fig S2). In order to see if variation was a regulated process, we grouped the protein variation by gene ontology terms (GO terms) and plotted them in ranked order. We saw that in general, processes xxx were…

Since we were able to measure all of these proteins within single cells, we are able to calculate coordinated expression of protein pairs at single cell resolution. Using Pearson correlation coefficient, we could determine coordinated expression of nearly 2 million protein pairs (Fig 1C) The distribution of correlation coefficients fit a normal distribution, with the majority of protein pairs appearing to not show significant coregulation. However, there appear to be a significant number of protein pairs containing high correlation coefficients, and the heatmap showed a lot of clustering of these highly coregulated pairs (Fig 1D).

### 2020-02-23 Manuscript/Manuscript.Rmd
added RMarkdown manuscript

---
title: 'Combined effects of protein expression variance and correlation on multicomponent systems'

author:

- name: Kyle M. Kovary
  affilnum: 1
  #email: kkovary@stanford.edu

- name: Mary N. Teruel
  affilnum: 1
  #email: mteruel@stanford.edu

affiliation:

- affilnum: 1
  affil: Department of Chemical and Systems Biology, Stanford University, Stanford, CA 94305, USA.

output:

  pdf_document:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    keep_tex: yes
    number_sections: no
    template: manuscript.latex
    includes:
      in_header: header.tex

  html_document: null

  word_document: null

documentclass: article
capsize: normalsize
geometry: margin=1in
fontsize: 11pt
spacing: doublespacing
bibliography: 
  - references.bib
  - knitcitations.bib
csl: cell.csl
abstract: Protein expression variation leads to phenotypic variance between cells. This has been demonstrated in cell signaling and differentiation decisions. Additionally coordinated expression of proteins between cells can tune signaling pathways either towards a more binary or analog modality. Though there is some evidence in bulk cell measurements and in bacteria that certain heteromeric subunits or metabolic pathways may be expressed in a coordinated fashion (i.e. operons in bacteria), there has not been a direct measurement of coordinated expression of proteins (independent of TFs) in metazoans (vertebrates?). Here, we measure cell-to-cell variability of relative protein abundance using quantitative proteomics of individual *Xenopus laevis* eggs and show that proteins involved in metabolic pathways or members of heteromeric complexes tend to have high correlations with other members of those pathways/complexes. Our previous work highlighted the fact that correlated expression increases the total variation of a pathway, so one would reason that certain pathways or complexes would need to compensate of this extra source of by reducing the variation of expression of these pathways or complexes. To test this we computed total variance score that took into account both the coefficient of variance and correlations between proteins in a pathway and found that the lower 10% of GO terms were highly enriched for metabolic pathways. When we looked at the relationship between CV and R between these GO terms we found a negative relationship between them, demonstrating that increased correlation needs to come at the expense of decreased variance. Simple molecular models of heteromeric complexes and metabolic pathways demonstrate that this tradeoff can result in higher efficiencies in both function and reduced energy waste. Together, our study argues for a control principle whereby coordinated expression of proteins in a pathway can require lower variance in order to reduce the overall pathway variance, enabling accurate control of active complexes or metabolic pathway activity.
---

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

#opts_knit$set(root.dir=normalizePath('../'))

### Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "output/figures/")

```

```{r knitcitations, echo=FALSE, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

*Keywords*: single cell, proteomics, stochasticity

\doublespace

\bleft

\justify

# Introduction

The coordinated expression of proteins is vital across cellular function, from maintaining a dynamic steady state to differentiation of cells into specialized types in different tissues. The ability to regulate coordinated expression has been demonstrated to occur via transcription (transcription factors, chromatin regulation), translation (specialized ribosomes, mRNA structure/modifications), and degradation (E3 ligases). These processes, as well as the molecules they target, are all subject to noise, which has been shown to lead to differences in cell behavior and decisions in otherwise identical cells. Variability and coordinated expression (correlation) of proteins leads to increased population level control in binary decisions, and a decreased ability to execute analog signaling.

Despite these important insights, a global understanding of the variability of protein expression between single cells and the coordinated expression of groups of proteins is largely an nonexistent. To systematically assess these properties of single cells, we carried out a proteomics experiment on exceeding large single cells, Xenopus laevis eggs. An advantage of this approach is that we can get around much of the signal to noise issues that accompany studying single cells due to low sample amounts. Additionally, at this stage of development transcription is restricted so we are able to gain insights into non-transcriptional control of protein expression. We also utilized isobaric tagging in order to measure 25 individual eggs in 5 mass spectrometry runs. In this study, we were able to measure the relative abundance >1000 proteins across single cells in order to better understand the properties of stochastic and coordinated expression of proteins.

With this dataset we have been able to, for the first time, measured the relationship between protein expression variance and coordinated protein expression on a proteome scale in single cells. We have observed that certain classes of proteins, including protein complexes and metabolic pathways, are expressed in such a way that increased coordinated expression is balanced by decreased variation. By doing this, cells are able to decrease the total variance of a given complex or pathway, an elegant balancing act that allows for finer control of metabolic throughput and controlling the number of potentially formed complexes though stoichiometric control. Though this kind of coordinated expression leads to an increase in variance in a population of cells, it can reduce variation within a cell.

To fetch bibliographic metadata automatically from the web. For example, citing a paper can be as easy as providing its DOI `r citep("10.1016/j.tree.2006.03.016")` or even just a few keywords `r citep(c("Heterogeneous ribosomes preferentially translate distinct subpools of mRNAs genome-wide","Controlling low rates of cell differentiation through noise and ultrahigh feedback"))`.

# Results

## Single cell proteomics reveals global protein expression variability and coordinated expression between protein pairs.

Activated Xenopus laevis eggs were collected at 5 time points across the first cell cycle (0, 20, 40, 60, and 80 minutes), with 5 eggs at each time point. Using TMT multiplexing and mass spectrometry, we were able to determine the relative abundance of more than 1300 proteins. Expression of these proteins were largely invariant across the cell cycle, revealing that these highly expressed genes are likely not regulated by cell cycle processes. Additionally, a PCA analysis of these eggs showed no discernible clustering on cell cycle time (Fig S1).

To determine the variance across these proteins, we calculated the coefficient of variation (CV) using all time points collected. This showed a wide range of variation containing multiple distributions (Fig 1B), many of which are consistent with our previous study of variation using targeted mass spectrometry (Fig S2). In order to see if variation was a regulated process, we grouped the protein variation by gene ontology terms (GO terms) and plotted them in ranked order. We saw that in general, processes xxx were…

Since we were able to measure all of these proteins within single cells, we are able to calculate coordinated expression of protein pairs at single cell resolution. Using Pearson correlation coefficient, we could determine coordinated expression of nearly 2 million protein pairs (Fig 1C) The distribution of correlation coefficients fit a normal distribution, with the majority of protein pairs appearing to not show significant coregulation. However, there appear to be a significant number of protein pairs containing high correlation coefficients, and the heatmap showed a lot of clustering of these highly coregulated pairs (Fig 1D).

## Protein modules display different levels of regulation of noise and co-expression of constituents

To investigate whether coordinated expression goes beyond just protein pairs, we wanted to see if pathways, complexes, and other modules had proteins that with coordinated expression. To answer this question, we grouped proteins by GO terms and calculated the median correlation coefficient across all pairs of proteins within the term. When the ranked media correlation is plotted it’s clear that there are a significant number of GO terms with higher than expected levels of coordinate expression.

## Total variance metric reveals certain modules enforce a tradeoff between protein co-expression and noise

Variance Sum Law — Independent Case

$$\sigma_{_{x+y}}^2 = \sigma_{_{x}}^2 + \sigma_{_{y}}^2$$

Variance Sum Law — Dependent Case

$$\sigma_{_{x+y}}^2 = \sigma_{_{x}}^2 + \sigma_{_{y}}^2 + 2Cov(x,y)$$
$$\rho = \dfrac{Cov(x,y)}{\sigma_{_{x}} \sigma_{_{y}}}$$
$$\sigma_{_{x+y}}^2 = \sigma_{_{x}}^2 + \sigma_{_{y}}^2 + 2\rho \sigma_{_{x}} \sigma_{_{y}}$$

$$\sigma_{{}_{total}} = \sqrt{\sum\limits_{i=1}^{n} \; \sigma_{i}^{2} \; \sum\limits_{j\neq i} \; 2\rho_{_{i,j}}\sigma_{_{i}}\sigma_{_{j}}}$$

$$\varsigma_{{}_{norm}} = \dfrac{\sigma_{{}_{total}}}{n}$$

When calculating the total variance of a pathway or general group of proteins, the total variance is the square root of the sums of the squares of the coefficient of variance. However, if any of the proteins are correlated in expression, an extra term must be added to account for this correlation. This means that if a group of proteins have a positive correlation, then the total variance will be higher than if there was no correlation. In order to study how variance and coordinated expression are related between single cells, we wanted to calculate a normalized total variance metric. Since we were not interested in the increase of variance due to the number of proteins in a module, we calculated the total variance using the coefficient of variation and correlation coefficients and normalize to the total number of proteins in the module. When total variance is plotted against the number of proteins, there is a clear linear relationship between the two. Once we normalize to the total number of proteins this relationship is lost and we are left with the effects of coefficient of variation and correlation.

## Low total variance modules are enriched for metabolic pathways and heteromeric protein complexes

## Enforcement of a tradeoff between protein co-expression and noise increases efficiency of metabolic pathways and heteromeric protein complexes

# Methods

## Collection and activation of Xenopus laevis eggs:

Xenopus egg extracts were prepared based on modifications of a previous protocol (Tsai et al, 2014). All of the animal protocols used in this manuscript were approved by the Stanford University Administrative Panel on Laboratory Animal Care. To induce egg laying, female Xenopus laevis were injected with human chorionic gonadotropin injection the night before each experiment. To collect the eggs, the frogs were subjected to pelvic massage, and the eggs were collected in 1X Marc’s Modified Ringer’s (MMR) buffer (0.1 M NaCl, 2 mM KCl, 1 mM MgCl2, 2 mM CaCl2, 5 mM HEPES, pH 7.8). To remove the jelly coat from the eggs, they were placed in a solution of 2% cysteine in 1× MMR buffer for 4 min and gently agitated, after which they were washed four times with 1× MMR buffer. To activate the cell cycle, eggs were placed in a solution of 0.5 $\mu$g/ml of calcium ionophore A23187 (Sigma) and 1X MMR buffer for 3 min, after which they were washed four times with 1× MMR buffer. Single eggs were collected at their respective timepoints and placed into 600uL tubes and snap frozen in liquid nitrogen before being stored at 80°C.

## Sample preparation for mass spectrometry:

Single eggs were lysed mechanically by pipetting the egg in 100$\mu$L of lysis buffer (100 mM NaCl, 25 mM Tris pH 8.2, Complete EDTA- free protease inhibitor cocktail (Sigma). The lysate was then placed in a 400 uLnatural polyethylene microcentrifuge tube (E&K
Scientific #485050) and spun at 15,000 g in a right angle centrifuge (Beckman Microfuge E) at 4°C for 5 min. The lipid layer was removed by using a razor blade to cut the tube off just beneath it, and the cytoplasmic fraction was pipetted into a 1.5-ml protein LoBind tube (Fisher Scientific #13-698-794), being careful to leave the yolk behind. To precipitate the proteins from the cytoplasmic fraction, 1 ml of ice cold acetone was added to each sample and placed at 20°C overnight. To collect precipitated proteins, the samples were centrifuged at
18,000 g for 20 min at 4°C. Acetone was decanted, and the protein pellets were resolubilized in 25$\mu$L of 8 M urea. To fully solubilize the protein pellet, the samples were placed in a shaker for 1 h at room temperature. The samples were then diluted to 2 M urea with 50 mM ammonium bicarbonate to a 100$\mu$L volume, after which protein concentration was measured in duplicate with a BCA assay by taking two 10$\mu$L aliquots of each sample. The proteins in the remaining 80$\mu$L of sample volume were reduced with 10 mM TCEP and incubated for 30 min at 37°C, then alkylated with 15 mM iodoacetamide and incubated in the dark at room temperature.

Next, the samples were diluted to 1 M urea with 50 mM ammonium bicarbonate. Trypsin (Promega #V5113) was then added at a ratio of 10 ng trypsin per 1ug protein (no < 500 ng was added to a sample). The trypsin digestion was carried out at 37°C for 12–16 h. To stop the trypsin, formic acid (Fisher #A117-50) was added at a ratio of 3$\mu$l per 100$\mu$l of sample to bring the pH down to < 3.

Peptides were cleaned up using an Oasis HLB uElution plate (Waters), equilibrated, and washed with 0.04% trifluoroacetic acid in water, and eluted in 80% acetonitrile with 0.2% formic acid. All solutions used are HPLC grade. Samples were then lyophilized. To remove any variance produced by phosphorylated peptides, the samples were phosphatase-treated. Peptides were resolubilized in 50$\mu$L of 1X NEBuffer 3 (no BSA), and calf intestinal alkaline phosphatase (NEB #M0290S) was added at a ratio 0.25 units per lg of peptide and incubated for 1 h at 37°C. The peptides were cleaned up again according to steps described above. Peptides were resolubilized in 2% acetonitrile and 0.1% formic acid before MS analysis.

## Mass spectrometry data collection:

Ask SUMS for details

## Mass spectrometry data analysis:

Ask SUMS for details

## Data processing:

To minimize the effects of non-biological variance, a correction factor was used to correct for these biases. First, each peptide was normalized by the median across all of the samples. Second, the vector of all peptides for each cell was divided by the median value across all peptides. Before peptides were used to estimate protein abundance, highly variable peptides needed to be filtered out. Peptides were grouped by their master UniProt accession number, and all peptides whose log transformed values were more than 2$\sigma$ away from the mean were discarded. Next, protein abundances were estimated by taking the median normalized value for each peptide for a protein in each sample. Missing protein levels were imputed using the k-nearest neighbors algorithm, with k being set to 1 and the similarity measure for distance the Gower’s distance between the proteome vectors.

\clearpage

# References

```{r write_citations, cache=FALSE, include=FALSE}
write.bibtex(file="knitcitations.bib")
```

<div id = "refs"></div>

\eleft

\clearpage

```{r sessioninfo, echo = FALSE, eval = TRUE}
# set eval = FALSE if you don't want this info (useful for reproducibility) to appear 
sessionInfo()
```

### 2020-02-23 Manuscript/Manuscript.tex
added RMarkdown manuscript

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Original default rstudio/pandoc latex file
%% upated by @jhollist 09/15/2014
%% inspired by @cboetting https://github.com/cboettig/template and
%% @rmflight blog posts:
%% http://rmflight.github.io/posts/2014/07/analyses_as_packages.html 
%% http://rmflight.github.io/posts/2014/07/vignetteAnalysis.html).  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt,]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Combined effects of protein expression variance and correlation on multicomponent systems},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Changes borrowed from @cboettig, added by @jhollist 
% A modified page layout 
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
\usepackage{lineno} % add 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%Packages and layout changes by @jhollist 09/15/2014
\usepackage{ragged2e}
\usepackage[font=normalsize]{caption}
  \usepackage[doublespacing]{setspace}
\usepackage{parskip}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\lfoot{\thepage}
%%Changed default abstract width and added lines
\renewenvironment{abstract}{
  \hfill\begin{minipage}{1\textwidth}
  \rule{\textwidth}{1pt}\vspace{5pt}
  \normalsize
  \begin{justify}
  \bfseries\abstractname\vspace{5pt}
  \end{justify}}
  {\par\noindent\rule{\textwidth}{1pt}\end{minipage}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Combined effects of protein expression variance and correlation on
multicomponent systems}
\author{
Kyle M. Kovary
Mary N. Teruel
}
\date{}
% Allowing for landscape pages
\usepackage{lscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

% Left justification of the text: see https://www.sharelatex.com/learn/Text_alignment
% \usepackage[document]{ragged2e} % already in the latex template
\newcommand{\bleft}{\begin{flushleft}}
\newcommand{\eleft}{\end{flushleft}}

%%Fix tightlist error: https://stackoverflow.com/questions/40438037/tightlist-error-using-pandoc-with-markdown
%%Added 2018-03-26 
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
%%%

\begin{document}
%%Edited by @jhollist 09/15/2014
%%Adds title from YAML
\begin{singlespace}
\begin{center}
\huge Combined effects of protein expression variance and correlation on
multicomponent systems
\end{center}
%%Adds Author, correspond email asterisk, and affilnum from YAML
\begin{center}
\large
Kyle M. Kovary \textsuperscript{1} 
Mary N. Teruel \textsuperscript{1} 
\end{center}
%%Adds affiliations from YAML
\begin{justify}
\footnotesize \emph{ 
\\*
\textsuperscript{1}Department of Chemical and Systems Biology, Stanford University,
Stanford, CA 94305, USA.\\*
}
%%Adds corresponding author email(s) from YAML
\newcounter{num}
\setcounter{num}{1}
\\[0.1cm]
\footnotesize \emph{ 
}
\end{justify}
%%Adds date from YAML
\normalsize

\begin{abstract}
Protein expression variation leads to phenotypic variance between cells.
This has been demonstrated in cell signaling and differentiation
decisions. Additionally coordinated expression of proteins between cells
can tune signaling pathways either towards a more binary or analog
modality. Though there is some evidence in bulk cell measurements and in
bacteria that certain heteromeric subunits or metabolic pathways may be
expressed in a coordinated fashion (i.e.~operons in bacteria), there has
not been a direct measurement of coordinated expression of proteins
(independent of TFs) in metazoans (vertebrates?). Here, we measure
cell-to-cell variability of relative protein abundance using
quantitative proteomics of individual \emph{Xenopus laevis} eggs and
show that proteins involved in metabolic pathways or members of
heteromeric complexes tend to have high correlations with other members
of those pathways/complexes. Our previous work highlighted the fact that
correlated expression increases the total variation of a pathway, so one
would reason that certain pathways or complexes would need to compensate
of this extra source of by reducing the variation of expression of these
pathways or complexes. To test this we computed total variance score
that took into account both the coefficient of variance and correlations
between proteins in a pathway and found that the lower 10\% of GO terms
were highly enriched for metabolic pathways. When we looked at the
relationship between CV and R between these GO terms we found a negative
relationship between them, demonstrating that increased correlation
needs to come at the expense of decreased variance. Simple molecular
models of heteromeric complexes and metabolic pathways demonstrate that
this tradeoff can result in higher efficiencies in both function and
reduced energy waste. Together, our study argues for a control principle
whereby coordinated expression of proteins in a pathway can require
lower variance in order to reduce the overall pathway variance, enabling
accurate control of active complexes or metabolic pathway activity.
\end{abstract}
\end{singlespace}

\emph{Keywords}: single cell, proteomics, stochasticity

\doublespace

\bleft

\justify

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The coordinated expression of proteins is vital across cellular
function, from maintaining a dynamic steady state to differentiation of
cells into specialized types in different tissues. The ability to
regulate coordinated expression has been demonstrated to occur via
transcription (transcription factors, chromatin regulation), translation
(specialized ribosomes, mRNA structure/modifications), and degradation
(E3 ligases). These processes, as well as the molecules they target, are
all subject to noise, which has been shown to lead to differences in
cell behavior and decisions in otherwise identical cells. Variability
and coordinated expression (correlation) of proteins leads to increased
population level control in binary decisions, and a decreased ability to
execute analog signaling.

Despite these important insights, a global understanding of the
variability of protein expression between single cells and the
coordinated expression of groups of proteins is largely an nonexistent.
To systematically assess these properties of single cells, we carried
out a proteomics experiment on exceeding large single cells, Xenopus
laevis eggs. An advantage of this approach is that we can get around
much of the signal to noise issues that accompany studying single cells
due to low sample amounts. Additionally, at this stage of development
transcription is restricted so we are able to gain insights into
non-transcriptional control of protein expression. We also utilized
isobaric tagging in order to measure 25 individual eggs in 5 mass
spectrometry runs. In this study, we were able to measure the relative
abundance \textgreater{}1000 proteins across single cells in order to
better understand the properties of stochastic and coordinated
expression of proteins.

With this dataset we have been able to, for the first time, measured the
relationship between protein expression variance and coordinated protein
expression on a proteome scale in single cells. We have observed that
certain classes of proteins, including protein complexes and metabolic
pathways, are expressed in such a way that increased coordinated
expression is balanced by decreased variation. By doing this, cells are
able to decrease the total variance of a given complex or pathway, an
elegant balancing act that allows for finer control of metabolic
throughput and controlling the number of potentially formed complexes
though stoichiometric control. Though this kind of coordinated
expression leads to an increase in variance in a population of cells, it
can reduce variation within a cell.

To fetch bibliographic metadata automatically from the web. For example,
citing a paper can be as easy as providing its DOI (Clark and Gelfand,
2006) or even just a few keywords (Ahrends et al., 2014; Shi et al.,
2017).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{single-cell-proteomics-reveals-global-protein-expression-variability-and-coordinated-expression-between-protein-pairs.}{%
\subsection{Single cell proteomics reveals global protein expression
variability and coordinated expression between protein
pairs.}\label{single-cell-proteomics-reveals-global-protein-expression-variability-and-coordinated-expression-between-protein-pairs.}}

Activated Xenopus laevis eggs were collected at 5 time points across the
first cell cycle (0, 20, 40, 60, and 80 minutes), with 5 eggs at each
time point. Using TMT multiplexing and mass spectrometry, we were able
to determine the relative abundance of more than 1300 proteins.
Expression of these proteins were largely invariant across the cell
cycle, revealing that these highly expressed genes are likely not
regulated by cell cycle processes. Additionally, a PCA analysis of these
eggs showed no discernible clustering on cell cycle time (Fig S1).

To determine the variance across these proteins, we calculated the
coefficient of variation (CV) using all time points collected. This
showed a wide range of variation containing multiple distributions (Fig
1B), many of which are consistent with our previous study of variation
using targeted mass spectrometry (Fig S2). In order to see if variation
was a regulated process, we grouped the protein variation by gene
ontology terms (GO terms) and plotted them in ranked order. We saw that
in general, processes xxx were\ldots{}

Since we were able to measure all of these proteins within single cells,
we are able to calculate coordinated expression of protein pairs at
single cell resolution. Using Pearson correlation coefficient, we could
determine coordinated expression of nearly 2 million protein pairs (Fig
1C) The distribution of correlation coefficients fit a normal
distribution, with the majority of protein pairs appearing to not show
significant coregulation. However, there appear to be a significant
number of protein pairs containing high correlation coefficients, and
the heatmap showed a lot of clustering of these highly coregulated pairs
(Fig 1D).

\hypertarget{protein-modules-display-different-levels-of-regulation-of-noise-and-co-expression-of-constituents}{%
\subsection{Protein modules display different levels of regulation of
noise and co-expression of
constituents}\label{protein-modules-display-different-levels-of-regulation-of-noise-and-co-expression-of-constituents}}

To investigate whether coordinated expression goes beyond just protein
pairs, we wanted to see if pathways, complexes, and other modules had
proteins that with coordinated expression. To answer this question, we
grouped proteins by GO terms and calculated the median correlation
coefficient across all pairs of proteins within the term. When the
ranked media correlation is plotted it's clear that there are a
significant number of GO terms with higher than expected levels of
coordinate expression.

\hypertarget{total-variance-metric-reveals-certain-modules-enforce-a-tradeoff-between-protein-co-expression-and-noise}{%
\subsection{Total variance metric reveals certain modules enforce a
tradeoff between protein co-expression and
noise}\label{total-variance-metric-reveals-certain-modules-enforce-a-tradeoff-between-protein-co-expression-and-noise}}

Variance Sum Law --- Independent Case

\[\sigma_{_{x+y}}^2 = \sigma_{_{x}}^2 + \sigma_{_{y}}^2\]

Variance Sum Law --- Dependent Case

\[\sigma_{_{x+y}}^2 = \sigma_{_{x}}^2 + \sigma_{_{y}}^2 + 2Cov(x,y)\]
\[\rho = \dfrac{Cov(x,y)}{\sigma_{_{x}} \sigma_{_{y}}}\]
\[\sigma_{_{x+y}}^2 = \sigma_{_{x}}^2 + \sigma_{_{y}}^2 + 2\rho \sigma_{_{x}} \sigma_{_{y}}\]

\[\sigma_{{}_{total}} = \sqrt{\sum\limits_{i=1}^{n} \; \sigma_{i}^{2} \; \sum\limits_{j\neq i} \; 2\rho_{_{i,j}}\sigma_{_{i}}\sigma_{_{j}}}\]

\[\varsigma_{{}_{norm}} = \dfrac{\sigma_{{}_{total}}}{n}\]

When calculating the total variance of a pathway or general group of
proteins, the total variance is the square root of the sums of the
squares of the coefficient of variance. However, if any of the proteins
are correlated in expression, an extra term must be added to account for
this correlation. This means that if a group of proteins have a positive
correlation, then the total variance will be higher than if there was no
correlation. In order to study how variance and coordinated expression
are related between single cells, we wanted to calculate a normalized
total variance metric. Since we were not interested in the increase of
variance due to the number of proteins in a module, we calculated the
total variance using the coefficient of variation and correlation
coefficients and normalize to the total number of proteins in the
module. When total variance is plotted against the number of proteins,
there is a clear linear relationship between the two. Once we normalize
to the total number of proteins this relationship is lost and we are
left with the effects of coefficient of variation and correlation.

\hypertarget{low-total-variance-modules-are-enriched-for-metabolic-pathways-and-heteromeric-protein-complexes}{%
\subsection{Low total variance modules are enriched for metabolic
pathways and heteromeric protein
complexes}\label{low-total-variance-modules-are-enriched-for-metabolic-pathways-and-heteromeric-protein-complexes}}

\hypertarget{enforcement-of-a-tradeoff-between-protein-co-expression-and-noise-increases-efficiency-of-metabolic-pathways-and-heteromeric-protein-complexes}{%
\subsection{Enforcement of a tradeoff between protein co-expression and
noise increases efficiency of metabolic pathways and heteromeric protein
complexes}\label{enforcement-of-a-tradeoff-between-protein-co-expression-and-noise-increases-efficiency-of-metabolic-pathways-and-heteromeric-protein-complexes}}

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{collection-and-activation-of-xenopus-laevis-eggs}{%
\subsection{Collection and activation of Xenopus laevis
eggs:}\label{collection-and-activation-of-xenopus-laevis-eggs}}

Xenopus egg extracts were prepared based on modifications of a previous
protocol (Tsai et al, 2014). All of the animal protocols used in this
manuscript were approved by the Stanford University Administrative Panel
on Laboratory Animal Care. To induce egg laying, female Xenopus laevis
were injected with human chorionic gonadotropin injection the night
before each experiment. To collect the eggs, the frogs were subjected to
pelvic massage, and the eggs were collected in 1X Marc's Modified
Ringer's (MMR) buffer (0.1 M NaCl, 2 mM KCl, 1 mM MgCl2, 2 mM CaCl2, 5
mM HEPES, pH 7.8). To remove the jelly coat from the eggs, they were
placed in a solution of 2\% cysteine in 1× MMR buffer for 4 min and
gently agitated, after which they were washed four times with 1× MMR
buffer. To activate the cell cycle, eggs were placed in a solution of
0.5 \(\mu\)g/ml of calcium ionophore A23187 (Sigma) and 1X MMR buffer
for 3 min, after which they were washed four times with 1× MMR buffer.
Single eggs were collected at their respective timepoints and placed
into 600uL tubes and snap frozen in liquid nitrogen before being stored
at 80°C.

\hypertarget{sample-preparation-for-mass-spectrometry}{%
\subsection{Sample preparation for mass
spectrometry:}\label{sample-preparation-for-mass-spectrometry}}

Single eggs were lysed mechanically by pipetting the egg in 100\(\mu\)L
of lysis buffer (100 mM NaCl, 25 mM Tris pH 8.2, Complete EDTA- free
protease inhibitor cocktail (Sigma). The lysate was then placed in a 400
uLnatural polyethylene microcentrifuge tube (E\&K Scientific \#485050)
and spun at 15,000 g in a right angle centrifuge (Beckman Microfuge E)
at 4°C for 5 min. The lipid layer was removed by using a razor blade to
cut the tube off just beneath it, and the cytoplasmic fraction was
pipetted into a 1.5-ml protein LoBind tube (Fisher Scientific
\#13-698-794), being careful to leave the yolk behind. To precipitate
the proteins from the cytoplasmic fraction, 1 ml of ice cold acetone was
added to each sample and placed at 20°C overnight. To collect
precipitated proteins, the samples were centrifuged at 18,000 g for 20
min at 4°C. Acetone was decanted, and the protein pellets were
resolubilized in 25\(\mu\)L of 8 M urea. To fully solubilize the protein
pellet, the samples were placed in a shaker for 1 h at room temperature.
The samples were then diluted to 2 M urea with 50 mM ammonium
bicarbonate to a 100\(\mu\)L volume, after which protein concentration
was measured in duplicate with a BCA assay by taking two 10\(\mu\)L
aliquots of each sample. The proteins in the remaining 80\(\mu\)L of
sample volume were reduced with 10 mM TCEP and incubated for 30 min at
37°C, then alkylated with 15 mM iodoacetamide and incubated in the dark
at room temperature.

Next, the samples were diluted to 1 M urea with 50 mM ammonium
bicarbonate. Trypsin (Promega \#V5113) was then added at a ratio of 10
ng trypsin per 1ug protein (no \textless{} 500 ng was added to a
sample). The trypsin digestion was carried out at 37°C for 12--16 h. To
stop the trypsin, formic acid (Fisher \#A117-50) was added at a ratio of
3\(\mu\)l per 100\(\mu\)l of sample to bring the pH down to \textless{}
3.

Peptides were cleaned up using an Oasis HLB uElution plate (Waters),
equilibrated, and washed with 0.04\% trifluoroacetic acid in water, and
eluted in 80\% acetonitrile with 0.2\% formic acid. All solutions used
are HPLC grade. Samples were then lyophilized. To remove any variance
produced by phosphorylated peptides, the samples were
phosphatase-treated. Peptides were resolubilized in 50\(\mu\)L of 1X
NEBuffer 3 (no BSA), and calf intestinal alkaline phosphatase (NEB
\#M0290S) was added at a ratio 0.25 units per lg of peptide and
incubated for 1 h at 37°C. The peptides were cleaned up again according
to steps described above. Peptides were resolubilized in 2\%
acetonitrile and 0.1\% formic acid before MS analysis.

\hypertarget{mass-spectrometry-data-collection}{%
\subsection{Mass spectrometry data
collection:}\label{mass-spectrometry-data-collection}}

Ask SUMS for details

\hypertarget{mass-spectrometry-data-analysis}{%
\subsection{Mass spectrometry data
analysis:}\label{mass-spectrometry-data-analysis}}

Ask SUMS for details

\hypertarget{data-processing}{%
\subsection{Data processing:}\label{data-processing}}

To minimize the effects of non-biological variance, a correction factor
was used to correct for these biases. First, each peptide was normalized
by the median across all of the samples. Second, the vector of all
peptides for each cell was divided by the median value across all
peptides. Before peptides were used to estimate protein abundance,
highly variable peptides needed to be filtered out. Peptides were
grouped by their master UniProt accession number, and all peptides whose
log transformed values were more than 2\(\sigma\) away from the mean
were discarded. Next, protein abundances were estimated by taking the
median normalized value for each peptide for a protein in each sample.
Missing protein levels were imputed using the k-nearest neighbors
algorithm, with k being set to 1 and the similarity measure for distance
the Gower's distance between the proteome vectors.

\clearpage

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-ahrends2014controlling}{}%
Ahrends, R., Ota, A., Kovary, K.M., Kudo, T., Park, B.O., and Teruel,
M.N. (2014). Controlling low rates of cell differentiation through noise
and ultrahigh feedback. Science \emph{344}, 1384--1389.

\leavevmode\hypertarget{ref-Clark_2006}{}%
Clark, J.S., and Gelfand, A.E. (2006). A future for models and data in
environmental science. Trends in Ecology \& Evolution \emph{21},
375--380.

\leavevmode\hypertarget{ref-shi2017heterogeneous}{}%
Shi, Z., Fujii, K., Kovary, K.M., Genuth, N.R., Röst, H.L., Teruel,
M.N., and Barna, M. (2017). Heterogeneous ribosomes preferentially
translate distinct subpools of mRNAs genome-wide. Molecular Cell
\emph{67}, 71--83.e7.

\eleft

\clearpage

\begin{verbatim}
R version 3.6.1 (2019-07-05)
Platform: x86_64-apple-darwin15.6.0 (64-bit)
Running under: macOS Catalina 10.15.3

Matrix products: default
BLAS:   /Library/Frameworks/R.framework/Versions/3.6/Resources/lib/libRblas.0.dylib
LAPACK: /Library/Frameworks/R.framework/Versions/3.6/Resources/lib/libRlapack.dylib

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base

other attached packages:
[1] knitcitations_1.0.10 knitr_1.28

loaded via a namespace (and not attached):
 [1] Rcpp_1.0.3        lubridate_1.7.4   digest_0.6.24     plyr_1.8.5       
 [5] R6_2.4.1          jsonlite_1.6.1    formatR_1.7       magrittr_1.5     
 [9] evaluate_0.14     bibtex_0.4.2.2    httr_1.4.1        rlang_0.4.4      
[13] stringi_1.4.6     curl_4.3          xml2_1.2.2        rmarkdown_2.1    
[17] tools_3.6.1       stringr_1.4.0     RefManageR_1.2.12 xfun_0.12        
[21] yaml_2.2.1        compiler_3.6.1    htmltools_0.4.0  
\end{verbatim}

\end{document}

### 2020-02-25 Analysis.Rmd
updated

@@
 ```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
-ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-20,20)
+# ggplot(pepTidy, aes(log2(intensity), colour = cell)) + 
+#   geom_density() + theme_bw() + xlim(-20,20)
 ```
@@
 
-ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
+# ggpairs(log2(pairs[,2:ncol(pairs)]), 
+#         lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
 ```
@@
 
-ggplot(meanSD, aes(x = log10(mean), y = sd)) + geom_point() + geom_smooth(method = lm, formula = y ~ poly(x, 2))
+# ggplot(meanSD, aes(x = log10(mean), y = sd)) + geom_point() + geom_smooth(method = lm, formula = y ~ poly(x, 2))
 ```
@@
 
-ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-8,8)
+# ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-8,8)
 ```
@@
 
-ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
+# ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
 ```
@@
 map = pheatmap(cor, show_colnames = F, show_rownames = F, color = mypal, breaks = seq(-1,1, by = 0.01), cutree_cols = k, cutree_rows = k, legend = F)
-map
+#map
 
@@
 GO <- protImpute %>% dplyr::select(Uniprot) %>% filter(Uniprot %in% keys) %>% 
-  rowwise() %>% 
-  mutate(GO = mapIds(org.Xl.eg.db, 
+  mutate(GOterms = mapIds(org.Xl.eg.db, 
                      keys = Uniprot, 
@@
                      column = 'GOALL', 
-                     multiVals = 'list')) %>% 
-  ungroup()
+                     multiVals = 'list'))
 
-allGO <- tibble(GOterms = unlist(GO$GO) %>% unique())
-allGO <- allGO %>% group_by(GOterms) %>% mutate(Uniprot = GO[grep(GOterms, GO$GO),1] %>% as.list()) %>% ungroup()
-
-write_csv(mutate(GO, GO = paste(unlist(GO), sep = '', collapse = ',')), '/Users/kylekovary/Documents/GitHub/Proteome-Expression-Covariation/analysisData/GO.csv')
-
-write_csv(mutate(allGO, Uniprot = paste(unlist(Uniprot), sep = '', collapse = ',')), '/Users/kylekovary/Documents/GitHub/Proteome-Expression-Covariation/analysisData/allGO.csv')
-
-```
-
-## GO Term Analysis
-```{r echo=TRUE, message=FALSE, warning=FALSE}
-GO <- read_csv('/Users/kylekovary/Documents/GitHub/Proteome-Expression-Covariation/analysisData/GO.csv') %>% mutate(GO = strsplit(GO, split = ','))
-
-allGO <- read_csv('/Users/kylekovary/Documents/GitHub/Proteome-Expression-Covariation/analysisData/allGO.csv') %>% 
-  mutate(Uniprot = strsplit(Uniprot, split = ','),
-         GO = mapIds(GO.db, 
+allGO <- unnest(GO, cols = GOterms) %>% nest(Uniprot = Uniprot) %>%
+  rowwise() %>% mutate(n = length(unlist(Uniprot))) %>% ungroup() %>%
+  mutate(GO = mapIds(GO.db, 
                      keys = GOterms, 
@@
 
+# allGO <- tibble(GOterms = unlist(GO$GO) %>% unique())
+# allGO <- allGO %>% group_by(GOterms) %>% mutate(Uniprot = GO[grep(GOterms, GO$GO),1] %>% as.list()) %>% ungroup() %>%
+#   mutate(
+#          GO = mapIds(GO.db, 
+#                      keys = GOterms, 
+#                      keytype = 'GOID', 
+#                      column = 'TERM') %>% as.character(),
+#          ontology = mapIds(GO.db, 
+#                      keys = GOterms, 
+#                      keytype = 'GOID', 
+#                      column = 'ONTOLOGY') %>% as.character()) %>%
+#   dplyr::select(GOterms, GO, ontology, everything())
+
 corVector <- function(uniprotList, corMat){
@@
 
-allGO <- allGO %>% group_by(GOterms) %>% dplyr::mutate(cors = corVector(Uniprot, cor) %>% as_tibble() %>% as.list()) %>% ungroup()
+selectCV <- function(x){
+  protVar %>% filter(Uniprot %in% unlist(x)) %>% 
+    dplyr::select(cv) %>% as_tibble() %>% as.list()
+}
 
-x <- allGO %>% rowwise %>% mutate(size = length(Uniprot)) %>% dplyr::select(GOterms, size, cors) %>% unnest() %>% group_by(GOterms) %>% 
+allGO <- allGO %>% group_by(GOterms) %>% 
+  dplyr::mutate(cors = corVector(Uniprot, cor) %>% as_tibble() %>% as.list(),
+                cv = selectCV(Uniprot)) %>% ungroup() %>%
+  rowwise() %>% mutate(medCors = mean(cors, na.rm = TRUE),
+                       medCV = mean(cv, na.rm = TRUE)) %>%
+  ungroup() %>%
+filter(length(Uniprot) > 3,
+       !is.na(medCors),
+       !is.na(medCV))
+
+#saveRDS(GO, 'analysisData/GO.RDS')
+#saveRDS(allGO, 'analysisData/allGO.RDS')
+```
+
+## GO Term Analysis
+```{r echo=TRUE, message=FALSE, warning=FALSE}
+#GO <- readRDS('analysisData/GO.RDS')
+#allGO <- readRDS('analysisData/allGO.RDS')
+
+x <- allGO %>% rowwise %>% mutate(size = nrow(Uniprot)) %>% dplyr::select(GOterms, size, cors) %>% unnest() %>% group_by(GOterms) %>% 
   summarise(median = median(cors, na.rm = T),
@@
 
-selectCV <- function(x){
-  protVar %>% filter(Uniprot %in% unlist(x)) %>% 
-    dplyr::select(cv) %>% as_tibble() %>% as.list()
-}
-
-allGO <- allGO %>% group_by(GOterms) %>% 
-  dplyr::mutate(cv = selectCV(Uniprot)) %>% ungroup()
-
 meanCV = allGO %>% dplyr::select(GOterms, cv) %>% unnest() %>% group_by(GOterms) %>% 
@@
 ```{r}
-test = allGO %>% rowwise() %>% mutate(n = length(Uniprot), medCors = mean(cors), medCV = mean(cv)) %>%
-  filter(length(Uniprot) > 3, !is.na(medCors), !is.na(medCV))
+test = allGO %>% filter(length(Uniprot) > 3, !is.na(medCors), !is.na(medCV))
 

### 2020-03-02 Untitled.Rmd
improved analysis

---
title: "Single Cell Xenopus TMT Analysis"
author: "Kyle Kovary"
date: "5/15/2019"
output: pdf_document
---

## Setup
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RColorBrewer)
library(pheatmap)
library(GGally)
library(VIM)
library(org.Xl.eg.db)
library(GO.db)
library(ggrepel)
library(plotly)
```

## Import Data
```{r setup, include=FALSE}

# Import Data
files = list.files('TMTdata/', pattern = 'TargetPeptideSpectrumMatch', full.names = T)

for(i in seq_along(files)){
  if(i == 1 & exists('pepRaw')){
    rm(pepRaw)
  } 
  if(!exists('pepRaw')){
    pepRaw = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
  } else{
    temp = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
    colnames(temp) = colnames(pepRaw)
    pepRaw = rbind(pepRaw, temp)
    rm(temp)
  }
}
pepRaw <- pepRaw %>% dplyr::select(file, massID, everything()) %>% filter(!is.na(`126`), 
                                                                  `Quan Usage` == 'Use',
                                                                  Confidence != 'Low')

```

## Data Organization
```{r }
pepTidy <- pepRaw %>% 
  mutate(run = c('ctrl','a','b','c','d','e')[match(pepRaw$file, files)]) %>% 
  group_by(file, `Annotated Sequence`) %>% 
  mutate(pepID = letters[1:n()]) %>% ungroup() %>%
  dplyr::select(file, run, pepID, everything()) %>%
  gather(key = label, value = intensity, '126':'131') %>% 
  dplyr::select(run, label, intensity, everything(), -file) %>% 
  mutate(condition = ifelse(run == 'ctrl', 'ctrl',c('ctrl',0,20,40,60,80)[match(label,c(126:131))]),
         intensity = as.numeric(intensity)) %>% unite(peptide, c(`Annotated Sequence`,'pepID')) %>%
  dplyr::rename(Uniprot = `Master Protein Accessions`) %>% 
  unite(cell, c('run','label','condition'), remove = F) %>%
  dplyr::select(cell, run, label, condition, peptide, Uniprot, intensity)
```

## Normalization
```{r}
protImpute <- pepTidy %>% 
  group_by(peptide,run) %>% 
  mutate(intensity = intensity / intensity[label == "126"]) %>%
  ungroup() %>%
  filter(run != "ctrl", condition != "ctrl") %>%
  #dplyr::filter(abs((intensity)) < 2*sd((pepTidy$intensity), na.rm = T)) %>%
  group_by(cell, run, label, condition, Uniprot) %>% 
  summarise(abundance = mean(intensity, na.rm = T)) %>% ungroup() %>%
  group_by(Uniprot) %>%
  mutate(abundance = abundance / mean(abundance, na.rm = T)) %>% 
  group_by(cell) %>% 
  mutate(abundance = abundance / mean(abundance, na.rm = T)) %>% 
  ungroup() %>% 
  dplyr::select(cell, Uniprot, abundance) %>% 
  spread(cell, abundance) %>%
  kNN(data = ., k = 1, impNA = TRUE) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)

```

## Summary Stats
```{r}
protVar <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
  separate(cell, into = c('run','label','condition'), remove = F) %>% 
  filter(condition != 'ctrl') %>% group_by(Uniprot) %>%
  summarise(var = var(abundance),
            sd = sd(abundance),
            mean = mean(abundance),
            cv = sd / mean)

cor = cor(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
colnames(cor) = protImpute$Uniprot
rownames(cor) = protImpute$Uniprot

cov = cov(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
colnames(cov) = protImpute$Uniprot
rownames(cov) = protImpute$Uniprot
```

## Map GO Terms to Proteins

```{r eval=TRUE, message=FALSE, warning=FALSE}
corVector <- function(uniprotList, corMat){
  targets = uniprotList %>% unlist()
  temp = corMat[rownames(corMat) %in% targets,colnames(corMat) %in% targets]
  temp[lower.tri(temp, diag = F)]
}

selectCV <- function(x){
  protVar %>% filter(Uniprot %in% unlist(x)) %>% 
    dplyr::select(cv) %>% as_tibble() %>% as.list()
}

keys <- keys(org.Xl.eg.db, keytype = 'UNIPROT')

allGO <- protImpute %>% dplyr::select(Uniprot) %>% filter(Uniprot %in% keys) %>% 
  mutate(GOterms = mapIds(org.Xl.eg.db, 
                     keys = Uniprot, 
                     keytype = 'UNIPROT', 
                     column = 'GOALL', 
                     multiVals = 'list')) %>%
  unnest(cols = GOterms) %>% nest(Uniprot = Uniprot) %>%
  rowwise() %>% mutate(n = length(unlist(Uniprot))) %>% ungroup() %>%
  mutate(GO = mapIds(GO.db, 
                     keys = GOterms, 
                     keytype = 'GOID', 
                     column = 'TERM') %>% as.character(),
         ontology = mapIds(GO.db, 
                     keys = GOterms, 
                     keytype = 'GOID', 
                     column = 'ONTOLOGY') %>% as.character()) %>%
  dplyr::select(GOterms, GO, ontology, everything())%>% 
  group_by(GOterms) %>% 
  dplyr::mutate(cors = corVector(Uniprot, cor) %>% as_tibble() %>% as.list(),
                covs = corVector(Uniprot, cov) %>% as_tibble() %>% as.list(),
                cv = selectCV(Uniprot)) %>% ungroup() %>%
  rowwise() %>% mutate(medCors = mean(cors, na.rm = TRUE),
                       medCV = mean(cv, na.rm = TRUE),
                       totVar = sqrt(sum(cv) + 2*sum(covs)),
                       normVar = totVar / n) %>%
  filter(n > 3) %>%
  ungroup()

```

## Plots
```{r fig.height=8, fig.width=5}
allGO1 <- allGO1 %>% mutate(interval = cut_number(normVar, n = 10))

ggplot(allGO1, aes(x = normVar)) + geom_histogram(bins = 50)

ggplot(allGO1, 
       aes(x = medCors, y = medCV, color = interval == "[0.00652,0.0333]")) + 
  geom_point(alpha = 0.5) + theme_bw() +
  #facet_wrap(~interval) +
  theme(legend.position = "bottom")

ggplot(allGO1, 
       aes(x = medCors, y = medCV)) + 
  geom_point(alpha = 0.5) + theme_bw() +
  geom_label_repel(data = filter(allGO1, 
                                 interval == "[0.00652,0.0333]",
                                 medCors > 0.15), 
                  aes(label = GO), 
                  colour = 'black', 
                  size = 2,
                  force = 10) + 
  theme_bw()
```

## Cor histogram
```{r}
goTerms <- c("GO:0006739","GO:0006096","GO:0015934")

allGO[grep('large ribosomal subunit',allGO$GO),]
glyc <- filter(allGO, GOterms == 'GO:0006096')$Uniprot %>% unlist()

cor_rug_plot <- function(x){
  temp <- filter(allGO, GOterms %in% x) %>%
    dplyr::select(GO, cors) %>% unnest(cols = cors)

  temp <- allGO %>% select(GO, cors) %>%
    unnest(cols = cors) %>% mutate(GO = "all") %>%
    rbind(temp,.)

  ggplot(temp, aes(x = cors, fill = GO, color = GO)) +
    geom_density(alpha = 0.25) + 
    geom_rug(data = filter(temp, GO != "all")) + 
    theme_bw()
}

```

## KEGG
```{r}
library(KEGG.db)
kegg <- plyr::ldply(as.list(org.Xl.egPATH2EG),data.frame) %>% as_tibble() %>%
  rename(kegg_id = ".id", gene_id = "X..i..")

kegg_names <- plyr::ldply(as.list(KEGGPATHNAME2ID),data.frame) %>%
  rename(kegg_name = ".id",
         kegg_id = "X..i..")

kegg <- protImpute %>% dplyr::select(Uniprot) %>% filter(Uniprot %in% keys) %>% 
  mutate(KEGG = mapIds(org.Xl.eg.db, 
                     keys = Uniprot, 
                     keytype = 'UNIPROT', 
                     column = "PATH", 
                     multiVals = 'list'))

allKEGG <- unnest(kegg, cols = KEGG) %>% nest(Uniprot = Uniprot) %>%
  rowwise() %>% mutate(n = length(unlist(Uniprot))) %>% ungroup() %>%
  mutate(KEGG_name = kegg_names$kegg_name[match(KEGG,kegg_names$kegg_id)])

corVector <- function(uniprotList, corMat){
  targets = uniprotList %>% unlist()
  temp = corMat[rownames(corMat) %in% targets,colnames(corMat) %in% targets]
  temp[lower.tri(temp, diag = F)]
}

selectCV <- function(x){
  protVar %>% filter(Uniprot %in% unlist(x)) %>% 
    dplyr::select(cv) %>% as_tibble() %>% as.list()
}

allKEGG <- allKEGG %>% group_by(KEGG) %>% 
  dplyr::mutate(cors = corVector(Uniprot, cor) %>% as_tibble() %>% as.list(),
                covs = corVector(Uniprot, cov) %>% as_tibble() %>% as.list(),
                cv = selectCV(Uniprot)) %>% ungroup() %>%
  rowwise() %>% mutate(medCors = mean(cors, na.rm = TRUE),
                       medCV = mean(cv, na.rm = TRUE)) %>%
  ungroup() %>%
filter(n > 3,
       !is.na(medCors),
       !is.na(medCV))

allKEGG <- allKEGG %>% 
  rowwise() %>% mutate(totVar = sqrt(sum(cv, na.rm = T) + sum(covs, na.rm = T)),
                             normVar = totVar / n) %>% ungroup() %>%
  mutate(interval = cut_number(normVar, n = 4))

ggplot(allKEGG, aes(x = medCors, y = medCV)) + geom_point(aes(color = interval)) +
  geom_label_repel(data = allKEGG, 
                  aes(label = KEGG_name), 
                  colour = 'black', 
                  size = 2,
                  force = 10)

glyc <- filter(allKEGG, KEGG == '00010')$Uniprot %>% unlist()

glycMat <- cor[rownames(cor) %in% glyc,colnames(cor) %in% glyc]
rownames(glycMat) <- mapIds(org.Xl.eg.db, keys = rownames(glycMat), keytype = 'UNIPROT', column = 'SYMBOL')
pheatmap(glycMat, show_colnames = T, show_rownames = T, color = mypal, breaks = seq(-1,1, by = 0.01),
         cutree_cols = 3, cutree_rows = 3)

# pathview(genes, 
#          pathway.id = pathway, 
#          species = 'mmu', 
#          same.layer = F,
#          low = list(gene = 'blue'),
#          mid = list(gene = 'transparent'),
#          high = list(gene = 'red'),
#          limit = list(gene = c(-2,2))
# )

```

### 2020-03-02 Untitled.Rmd
added glycolysis cluster analysis

@@
   filter(run != "ctrl", condition != "ctrl") %>%
-  #dplyr::filter(abs((intensity)) < 2*sd((pepTidy$intensity), na.rm = T)) %>%
+  dplyr::filter(abs((intensity)) < 2*sd((pepTidy$intensity), na.rm = T)) %>%
   group_by(cell, run, label, condition, Uniprot) %>% 
@@
 
-ggplot(allKEGG, aes(x = medCors, y = medCV)) + geom_point(aes(color = interval)) +
+ggplot(allKEGG, aes(x = corSkew, y = medCV)) + geom_point(aes(color = interval)) +
   geom_label_repel(data = allKEGG, 
@@
 rownames(glycMat) <- mapIds(org.Xl.eg.db, keys = rownames(glycMat), keytype = 'UNIPROT', column = 'SYMBOL')
-pheatmap(glycMat, show_colnames = T, show_rownames = T, color = mypal, breaks = seq(-1,1, by = 0.01),
-         cutree_cols = 3, cutree_rows = 3)
+
+row_anno <- data.frame(cv = unlist(allKEGG$cv[1]))
+rownames(row_anno) <- rownames(glycMat)
+
+map <- pheatmap(glycMat, show_colnames = T, show_rownames = T, color = mypal, breaks = seq(-1,1, by = 0.01),
+         cutree_cols = 3, cutree_rows = 3, annotation_row = row_anno)
+
+map.clust <- cbind(t(glycMat), cluster = cutree(map$tree_row, k = 3))
+map.clust[,ncol(map.clust) == 2]
+
+glyc2 <- names(which(map.clust[,ncol(map.clust)] == 2))
+totVarGroup(glyc2) / length(glyc2)
+protVar %>% filter(Uniprot %in% glyc2) %>%
+  summarise(meanCV = mean(cv))
+
+glyc2 <- glycMat[which(map.clust[,ncol(map.clust)] == 2),which(map.clust[,ncol(map.clust)] == 2)]
+pheatmap(glyc2,
+         color = mypal, breaks = seq(-1,1, by = 0.01))
+
+glyc2 <- tibble(GO = "Glycolysis",
+       cor = glyc2[lower.tri(glyc2)])
+
+rbind(glyc2,tibble(GO = "All",
+      cor = cor[lower.tri(cor)])) %>%
+  ggplot(., aes(x = cor, fill = GO, alpha = 0.75)) + geom_density() +
+  geom_rug(data = glyc2)
 

### 2020-03-03 Analysis.Rmd
update

@@
     rm(pepRaw)
-  } else if(!exists('pepRaw')){
+  } 
+  if(!exists('pepRaw')){
     pepRaw = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
@@
   mutate(condition = ifelse(run == 'ctrl', 'ctrl',c('ctrl',0,20,40,60,80)[match(label,c(126:131))]),
-         intensity = as.numeric(intensity)) %>% unite(peptide, c(`Annotated Sequence`,'pepID')) %>%
+         intensity = as.numeric(intensity)) %>% 
+  unite(peptide, c(`Annotated Sequence`,'pepID')) %>%
   dplyr::rename(Uniprot = `Master Protein Accessions`) %>% 
@@
 
-```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
-meanSD = pepTidy %>% filter(condition != 'ctrl') %>% 
-  group_by(peptide) %>% 
-  summarise(mean = mean(intensity, na.rm = T), 
-            sd = sd(intensity, na.rm = T) / mean)
-
-# ggplot(meanSD, aes(x = log10(mean), y = sd)) + geom_point() + geom_smooth(method = lm, formula = y ~ poly(x, 2))
-```
-
 ## Variance and Co-variance Between Cells After Cell Normalization
@@
 ```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
-pepTidy = pepTidy %>% group_by(cell) %>% mutate(intensity = intensity / median(intensity, na.rm = T))
+pepTidy = pepTidy %>% group_by(cell) %>% mutate(intensity = intensity / mean(intensity, na.rm = T))
 
@@
 pepTidy = pepTidy %>% group_by(run, peptide) %>% 
-  mutate(intensity = intensity / median(intensity, na.rm = T)) %>% ungroup()
+  mutate(intensity = intensity / mean(intensity, na.rm = T)) %>% ungroup()
 
@@
 
-There are
-
 ## Variance and Co-variance Between Cells After Removing Peptides Outside of 2*sd
@@
 ```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
-pepTidy %>% dplyr::filter(abs(log2(intensity)) < 2*sd(log2(pepTidy$intensity), na.rm = T))
+pepTidy <- pepTidy %>% dplyr::filter(abs(log2(intensity)) < 2*sd(log2(pepTidy$intensity), na.rm = T))
 
-ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-4,4)
+#ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-4,4)
 ```
@@
 
+protVar <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
+  separate(cell, into = c('run','label','condition'), remove = F) %>% 
+  filter(condition != 'ctrl') %>% group_by(Uniprot) %>%
+  summarise(var = var(abundance),
+            sd = sd(abundance),
+            mean = mean(abundance),
+            cv = sd / mean)
+
+ggplot(protVar, aes(x = sd)) + geom_histogram(bins = 50) + theme_bw()
+
 cor = cor(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
@@
 mypal = rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201))
-k = 2
+k = 1
 map = pheatmap(cor, show_colnames = F, show_rownames = F, color = mypal, breaks = seq(-1,1, by = 0.01), cutree_cols = k, cutree_rows = k, legend = F)
-#map
+map
 
-tibble(Uniprot = protImpute$Uniprot, Cluster = cutree(map$tree_row, k = k)) %>% write_csv('/Users/kylekovary/Downloads/Map.csv')
+cov = cov(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
+colnames(cov) = protImpute$Uniprot
+rownames(cov) = protImpute$Uniprot
+
+#tibble(Uniprot = protImpute$Uniprot, Cluster = cutree(map$tree_row, k = k)) %>% write_csv('/Users/kylekovary/Downloads/Map.csv')
 
@@
 pcaDF = pcaDF %>% separate(cell, into = c('run','label','condition'), remove = F) %>% print()
-ggplot(pcaDF, aes(x = PC1, y = PC2, colour = condition)) + geom_point(size = 5)
+ggplot(pcaDF, aes(x = PC1, y = PC2, colour = condition)) + stat_ellipse() + 
+  geom_point(size = 5)
 
@@
 
-```{r eval=F}
-columns(org.Xl.eg.db)
+```{r eval=TRUE}
+#columns(org.Xl.eg.db)
 #mapIds(org.Xl.eg.db, keys = 'A0JPG5', keytype = 'UNIPROT', column = 'GOALL', multiVals = 'list')
@@
   dplyr::mutate(cors = corVector(Uniprot, cor) %>% as_tibble() %>% as.list(),
+                covs = corVector(Uniprot, cov) %>% as_tibble() %>% as.list(),
                 cv = selectCV(Uniprot)) %>% ungroup() %>%
@@
   ungroup() %>%
-filter(length(Uniprot) > 3,
+filter(n > 3,
        !is.na(medCors),
        !is.na(medCV))
+
+allGO <- allGO %>% rowwise %>% mutate(totVar = sqrt(sum(cv) + sum(covs)),
+                             normVar = totVar / n)
 
@@
 
-protVar <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
-  separate(cell, into = c('run','label','condition'), remove = F) %>% 
-  filter(condition != 'ctrl') %>% group_by(Uniprot) %>%
-  summarise(cv = sd(abundance)/mean(abundance))
-
-ggplot(protVar, aes(x = cv)) + geom_histogram(bins = 50) + theme_bw()
-
 meanCV = allGO %>% dplyr::select(GOterms, cv) %>% unnest() %>% group_by(GOterms) %>% 
@@
 ```{r message=FALSE, warning=FALSE}
-allGO[grep('ribosom',allGO$GO),]
-glyc <- filter(allGO, GOterms == 'GO:0005840')$Uniprot %>% unlist()
+allGO[grep('glycoly',allGO$GO),]
+glyc <- filter(allGO, GOterms == 'GO:0006096')$Uniprot %>% unlist()
 
@@
 
+```{r}
+x <- tibble(cor = c(corVect$cor,glycMat[lower.tri(glycMat)]),
+       label = c(rep('All', nrow(corVect)),
+                 rep('Gly',length(glycMat[lower.tri(glycMat)]))))
+
+ggplot(x, aes(x = cor, fill = label, color = label)) + geom_density(alpha = 0.5) + geom_rug(data = filter(x, label == 'Gly'))
+
+```
+
 ## Cor vs CV
 ```{r}
-test = allGO %>% filter(length(Uniprot) > 3, !is.na(medCors), !is.na(medCV))
+test <- allGO %>% filter(n > 3, !is.na(medCors), !is.na(medCV))
 
@@
 allGO1 <- allGO %>% filter(!is.na(GOterms)) %>% rowwise() %>%
-  mutate(n = length(unlist(Uniprot)),
-         totVar = totVarGroup(Uniprot),
+  mutate(totVar = totVarGroup(Uniprot),
          normVar = totVar / n) %>%
-  filter(n > 3)
+  filter(n > 3) %>% ungroup()
 
@@
 
-ggplot(allGO1, aes(x = normVar)) + geom_histogram(bins = 30) + 
+ggplot(allGO, aes(x = normVar)) + geom_histogram(bins = 30) + 
   geom_vline(xintercept = c(0.15, 0.1), colour = 'red', linetype = 'dashed') +
@@
 dat <- left_join(ungroup(test), allGO1, by = "GOterms") %>%
-  dplyr::select(GOterms, GO.x, ontology.x, Uniprot.x, cors.x, cv.x, n.x,medCors, medCV, normVar)
+  dplyr::select(GOterms, GO.x, ontology.x, Uniprot.x, cors.x, cv.x, n.x,medCors.x, medCV.x, normVar) %>%
+  dplyr::rename(GO = GO.x, ontology = ontology.x, Uniprot = Uniprot.x, cors = cors.x, cv = cv.x,
+                n = n.x, medCors = medCors.x, medCV = medCV.x) %>%
+  mutate(score = ifelse(normVar > 0.145,
+                              'High',
+                              ifelse(normVar < 0.115,
+                                     'Low',
+                                     'Med')))
 
-dat$score <- 'Med'
-dat[dat$normVar > 0.145, 'score'] <- 'High'
-dat[dat$normVar < 0.115, 'score'] <- 'Low'
+# dat$score <- 'Med'
+# dat[dat$normVar > 0.145, 'score'] <- 'High'
+# dat[dat$normVar < 0.115, 'score'] <- 'Low'
 
@@
 
-ggplot(filter(dat, n.x > 5), aes(x = medCors, y = medCV, colour = score)) + 
+ggplot(filter(dat, n > 5), aes(x = medCors, y = medCV, colour = score)) + 
   geom_point(alpha = 0.75) + scale_color_manual(values=c("blue", "red", "grey"))
@@
 
-ggplot(filter(dat, n.x > 5, score != 'Med', ontology.x != 'CC'), aes(x = medCors, y = medCV, colour = normVar)) + 
+ggplot(filter(dat, n > 5, score != 'Med', ontology != 'CC'), aes(x = medCors, y = medCV, colour = normVar)) + 
   geom_point() + scale_colour_gradient(low="red", high="blue") + 
   geom_smooth(method = 'lm') + facet_wrap(~score, scales = 'free') +
-  geom_text_repel(data = filter(dat, n.x > 5, score != 'Med'), aes(label = GO.x), colour = 'black', size = 2)
+  geom_text_repel(data = filter(dat, n > 5, score != 'Med'), aes(label = GO), colour = 'black', size = 2)
 
-plot_ly(filter(dat, n.x > 5), z = ~normVar, y = ~medCV, x = ~medCors, color = ~score, colors = c("red", "blue", "grey")) %>%
+plot_ly(filter(dat, n > 5), z = ~normVar, y = ~medCV, x = ~medCors, color = ~score, colors = c("red", "blue", "grey")) %>%
   add_markers() %>%
@@
 
-ggplot(filter(dat, n.x > 5), aes(y = normVar, x = medCV)) + geom_point(alpha = 0.75)
-ggplot(filter(dat, n.x > 5), aes(y = normVar, x = medCors)) + geom_point(alpha = 0.75)
+ggplot(filter(dat, n > 5), aes(y = normVar, x = medCV)) + geom_point(alpha = 0.75)
+ggplot(filter(dat, n > 5), aes(y = normVar, x = medCors)) + geom_point(alpha = 0.75)
 
 library(GGally)
-p <- ggpairs(filter(dat, n.x > 5), columns = c('medCV', 'medCors','normVar'), 
+p <- ggpairs(filter(dat, n > 5), columns = c('medCV', 'medCors','normVar'), 
         ggplot2::aes(colour = score, alpha = 0.5))

### 2020-03-03 Manuscript/Manuscript.Rmd
updated
update

@@
   word_document: null
-
+z
 documentclass: article
@@
 
-Variance Sum Law — Independent Case
-
-$$\sigma_{_{x+y}}^2 = \sigma_{_{x}}^2 + \sigma_{_{y}}^2$$
-
 Variance Sum Law — Dependent Case
 
-$$\sigma_{_{x+y}}^2 = \sigma_{_{x}}^2 + \sigma_{_{y}}^2 + 2Cov(x,y)$$
-$$\rho = \dfrac{Cov(x,y)}{\sigma_{_{x}} \sigma_{_{y}}}$$
-$$\sigma_{_{x+y}}^2 = \sigma_{_{x}}^2 + \sigma_{_{y}}^2 + 2\rho \sigma_{_{x}} \sigma_{_{y}}$$
+The sum of variance for a group of proteins $X_{1},\dotsc,X_{n}$ with dependent relationships:
 
-$$\sigma_{{}_{total}} = \sqrt{\sum\limits_{i=1}^{n} \; \sigma_{i}^{2} \; \sum\limits_{j\neq i} \; 2\rho_{_{i,j}}\sigma_{_{i}}\sigma_{_{j}}}$$
+$$var\left(\sum_{i=1}^n X_{i}\right) = \sum_{i=1}^{n} var(X_{i}) + 2\sum_{i,j:i<j}^{n} cov(X_{i},X_{j})$$
 
-$$\varsigma_{{}_{norm}} = \dfrac{\sigma_{{}_{total}}}{n}$$
+$$\sigma_{{}_{total}} = \sqrt{var\left(\sum_{i=1}^n X_{i}\right)}$$
+
+$$\sigma_{{}_{norm}} = \dfrac{\sigma_{{}_{total}}}{n}$$
 

### 2020-03-03 Peptide Analysis.Rmd
update

--- 2020-02-25 Analysis.Rmd
@@
 ---
-title: "Single Cell Xenopus TMT Analysis"
+title: "Single Cell Xenopus TMT Analysis2"
 author: "Kyle Kovary"
@@
 ```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
-pairs = pepTidy %>% filter(condition == 'ctrl') %>% 
-  dplyr::select(cell, peptide, intensity) %>% spread(cell, intensity)
+pepTidy %>% filter(condition == 'ctrl') %>% 
+  dplyr::select(cell, peptide, intensity) %>% spread(cell, intensity) %>%
 
@@
 
-```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
-meanSD = pepTidy %>% filter(condition != 'ctrl') %>% 
+```{r echo=TRUE, fig.height=4, fig.width=4,message=FALSE, warning=FALSE}
+pepTidy %>% filter(condition != 'ctrl') %>% 
+  mutate(intensity = log2(intensity)) %>%
   group_by(peptide) %>% 
   summarise(mean = mean(intensity, na.rm = T), 
-            sd = sd(intensity, na.rm = T) / mean)
+            median = median(intensity, na.rm = T), 
+            sd = sd(intensity, na.rm = T),
+            cv = sd / mean) %>%
+ggplot(., aes(x = median, y = cv)) + 
+  geom_point(size = 0.25, alpha = 0.1) + 
+  geom_smooth(method = lm, formula = y ~ poly(x, 2)) +
+  ylim(0,0.2)
+```
 
-# ggplot(meanSD, aes(x = log10(mean), y = sd)) + geom_point() + geom_smooth(method = lm, formula = y ~ poly(x, 2))
+## Variance and Co-variance Between Cells After Removing Peptides Outside of 2*sd
+
+```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
+pepTidy <- pepTidy %>% dplyr::filter(abs(log2(intensity)) < 2*sd(log2(pepTidy$intensity), na.rm = T))
+
+#ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-4,4)
+```
+
+The standard deviation of all peptides across eggs was calculated, and every peptide that had a value outside of `r 2*sd(log2(pepTidy$intensity), na.rm = T)`
+
+## Group into proteins based on median
+
+```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
+protTidy = pepTidy %>% group_by(cell, run, label, condition, Uniprot) %>% 
+  summarise(abundance = median(intensity, na.rm = T)) %>% ungroup()
+
+pairs = protTidy %>% dplyr::filter(condition == 'ctrl') %>% 
+  dplyr::select(cell, Uniprot, abundance) %>% 
+  spread(cell, abundance)
+
+ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
+
+pairs = protTidy %>% dplyr::select(cell, Uniprot, abundance) %>% spread(cell, abundance)
+
+```
+
+## Impute data
+```{r}
+protImpute = kNN(data = pairs, k = 1) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)
+
+protVar <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
+  separate(cell, into = c('run','label','condition'), remove = F) %>% 
+  filter(condition != 'ctrl') %>% group_by(Uniprot) %>%
+  summarise(var = var(abundance),
+            sd = sd(abundance),
+            mean = mean(abundance),
+            cv = sd / mean)
+
+ggplot(protVar, aes(x = sd)) + geom_histogram(bins = 50) + theme_bw()
+```
+
+## Variance vs Mean
+```{r}
+ggplot(protVar, aes(x = log2(mean), y = var/mean^2)) + geom_point() + geom_smooth(method = lm, formula = y ~ poly(x, 2))
+```
+
+## Calculate variance and covariance
+```{r}
+cor = cor(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
+colnames(cor) = protImpute$Uniprot
+rownames(cor) = protImpute$Uniprot
+
+cov = cov(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
+colnames(cov) = protImpute$Uniprot
+rownames(cov) = protImpute$Uniprot
+
+mypal = rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201))
+k = 2
+# map = pheatmap(cor, show_colnames = F, show_rownames = F, color = mypal, breaks = seq(-1,1, by = 0.01), cutree_cols = k, cutree_rows = k, legend = F)
+#map
+
+tibble(Uniprot = protImpute$Uniprot, Cluster = cutree(map$tree_row, k = k)) %>% write_csv('/Users/kylekovary/Downloads/Map.csv')
+```
+
+## PCA Analysis
+```{r}
+remove = which(is.na(pairs), arr.ind = T)[,1] %>% unique()
+pca = prcomp(protImpute[,2:ncol(protImpute)])
+
+pcaDF = tibble(cell = rownames(pca$rotation), PC1 = pca$rotation[,1], PC2 = pca$rotation[,2])
+pcaDF = pcaDF %>% separate(cell, into = c('run','label','condition'), remove = F) %>% print()
+ggplot(pcaDF, aes(x = PC1, y = PC2, colour = condition)) + geom_point(size = 5)
 ```
@@
 ```
-
-There are
-
-## Variance and Co-variance Between Cells After Removing Peptides Outside of 2*sd
-
-```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
-pepTidy %>% dplyr::filter(abs(log2(intensity)) < 2*sd(log2(pepTidy$intensity), na.rm = T))
-
-ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-4,4)
-```
-
-The standard deviation of all peptides across eggs was calculated, and every peptide that had a value outside of `r 2*sd(log2(pepTidy$intensity), na.rm = T)`
-
-## Group into proteins based on median
-
-```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
-protTidy = pepTidy %>% filter(abs(log2(intensity)) < 2*sd(log2(pepTidy$intensity), na.rm = T)) %>% group_by(cell, run, label, condition, Uniprot) %>% 
-  summarise(abundance = median(intensity, na.rm = T)) %>% ungroup()
-
-pairs = protTidy %>% dplyr::filter(condition == 'ctrl') %>% 
-  dplyr::select(cell, Uniprot, abundance) %>% 
-  spread(cell, abundance)
-
-ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
-
-pairs = protTidy %>% dplyr::select(cell, Uniprot, abundance) %>% spread(cell, abundance)
-
-protImpute = kNN(data = pairs, k = 1) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)
-
-cor = cor(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
-colnames(cor) = protImpute$Uniprot
-rownames(cor) = protImpute$Uniprot
-
-mypal = rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201))
-k = 2
-map = pheatmap(cor, show_colnames = F, show_rownames = F, color = mypal, breaks = seq(-1,1, by = 0.01), cutree_cols = k, cutree_rows = k, legend = F)
-#map
-
-tibble(Uniprot = protImpute$Uniprot, Cluster = cutree(map$tree_row, k = k)) %>% write_csv('/Users/kylekovary/Downloads/Map.csv')
-
-remove = which(is.na(pairs), arr.ind = T)[,1] %>% unique()
-pca = prcomp(protImpute[,2:ncol(protImpute)])
-
-pcaDF = tibble(cell = rownames(pca$rotation), PC1 = pca$rotation[,1], PC2 = pca$rotation[,2])
-pcaDF = pcaDF %>% separate(cell, into = c('run','label','condition'), remove = F) %>% print()
-ggplot(pcaDF, aes(x = PC1, y = PC2, colour = condition)) + geom_point(size = 5)
-
-```
-
-## Map GO Terms to Proteins
-
-```{r eval=F}
-columns(org.Xl.eg.db)
-#mapIds(org.Xl.eg.db, keys = 'A0JPG5', keytype = 'UNIPROT', column = 'GOALL', multiVals = 'list')
-
-keys <- keys(org.Xl.eg.db, keytype = 'UNIPROT')
-
-GO <- protImpute %>% dplyr::select(Uniprot) %>% filter(Uniprot %in% keys) %>% 
-  mutate(GOterms = mapIds(org.Xl.eg.db, 
-                     keys = Uniprot, 
-                     keytype = 'UNIPROT', 
-                     column = 'GOALL', 
-                     multiVals = 'list'))
-
-allGO <- unnest(GO, cols = GOterms) %>% nest(Uniprot = Uniprot) %>%
-  rowwise() %>% mutate(n = length(unlist(Uniprot))) %>% ungroup() %>%
-  mutate(GO = mapIds(GO.db, 
-                     keys = GOterms, 
-                     keytype = 'GOID', 
-                     column = 'TERM') %>% as.character(),
-         ontology = mapIds(GO.db, 
-                     keys = GOterms, 
-                     keytype = 'GOID', 
-                     column = 'ONTOLOGY') %>% as.character()) %>%
-  dplyr::select(GOterms, GO, ontology, everything())
-
-# allGO <- tibble(GOterms = unlist(GO$GO) %>% unique())
-# allGO <- allGO %>% group_by(GOterms) %>% mutate(Uniprot = GO[grep(GOterms, GO$GO),1] %>% as.list()) %>% ungroup() %>%
-#   mutate(
-#          GO = mapIds(GO.db, 
-#                      keys = GOterms, 
-#                      keytype = 'GOID', 
-#                      column = 'TERM') %>% as.character(),
-#          ontology = mapIds(GO.db, 
-#                      keys = GOterms, 
-#                      keytype = 'GOID', 
-#                      column = 'ONTOLOGY') %>% as.character()) %>%
-#   dplyr::select(GOterms, GO, ontology, everything())
-
-corVector <- function(uniprotList, corMat){
-  targets = uniprotList %>% unlist()
-  temp = corMat[rownames(corMat) %in% targets,colnames(corMat) %in% targets]
-  temp[lower.tri(temp, diag = F)]
-}
-
-selectCV <- function(x){
-  protVar %>% filter(Uniprot %in% unlist(x)) %>% 
-    dplyr::select(cv) %>% as_tibble() %>% as.list()
-}
-
-allGO <- allGO %>% group_by(GOterms) %>% 
-  dplyr::mutate(cors = corVector(Uniprot, cor) %>% as_tibble() %>% as.list(),
-                cv = selectCV(Uniprot)) %>% ungroup() %>%
-  rowwise() %>% mutate(medCors = mean(cors, na.rm = TRUE),
-                       medCV = mean(cv, na.rm = TRUE)) %>%
-  ungroup() %>%
-filter(length(Uniprot) > 3,
-       !is.na(medCors),
-       !is.na(medCV))
-
-#saveRDS(GO, 'analysisData/GO.RDS')
-#saveRDS(allGO, 'analysisData/allGO.RDS')
-```
-
-## GO Term Analysis
-```{r echo=TRUE, message=FALSE, warning=FALSE}
-#GO <- readRDS('analysisData/GO.RDS')
-#allGO <- readRDS('analysisData/allGO.RDS')
-
-x <- allGO %>% rowwise %>% mutate(size = nrow(Uniprot)) %>% dplyr::select(GOterms, size, cors) %>% unnest() %>% group_by(GOterms) %>% 
-  summarise(median = median(cors, na.rm = T),
-            sd = sd(cors, na.rm = T),
-            size = mean(size)) %>% 
-  mutate(GOterms = fct_reorder(GOterms, median))
-
-ggplot(filter(x, size >= 2), aes(x = GOterms, y = median)) + geom_point(alpha = 0.25) +
-  geom_hline(yintercept = 0.02154303, colour = 'red', linetype = 'dashed') + 
-  theme_bw() + ylab('Mean Correlation') + ylim(-1,1)
-
-```
-
-## Plot highest correlated GO Terms with GO Term Names
-```{r}
-topGOnames <- x %>% filter(size > 2) %>% .[order(.$median, decreasing = T),] %>% 
-  .[1:50,'GOterms'] %>% unlist()
-
-topGO <- allGO %>% filter(GOterms %in% topGOnames) %>% 
-  dplyr::select(GOterms, GO, cors) %>% unnest() %>%
-  group_by(GO) %>% mutate(median = median(cors)) %>% ungroup() %>%
-  mutate(GO = fct_reorder(GO, median))
-
-ggplot(topGO, aes(x = GO, y = cors)) + geom_boxplot() + geom_jitter(alpha = 0.5, size = 0.5) + coord_flip() + scale_x_discrete(labels = function(x) str_wrap(x, width = 50)) + theme_bw() + ylim(-1,1)
-
-mitoProts <- allGO %>% filter(GOterms == 'GO:0051321') %>% dplyr::select(Uniprot) %>% unnest() %>% unlist()
-mitoDF <- protTidy %>% filter(condition != 'ctrl', Uniprot %in% mitoProts) %>% group_by(condition, Uniprot) %>% summarise(mean = mean(abundance), sd = sd(abundance) / (n()-1))
-
-ggplot(mitoDF, aes(x = as.numeric(condition), y = mean, colour = Uniprot)) + geom_point() + geom_path() + geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0, alpha = 0.5) + theme_bw()
-
-```
-
-## Variance of GO Terms
-```{r}
-
-protVar <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
-  separate(cell, into = c('run','label','condition'), remove = F) %>% 
-  filter(condition != 'ctrl') %>% group_by(Uniprot) %>%
-  summarise(cv = sd(abundance)/mean(abundance))
-
-ggplot(protVar, aes(x = cv)) + geom_histogram(bins = 50) + theme_bw()
-
-meanCV = allGO %>% dplyr::select(GOterms, cv) %>% unnest() %>% group_by(GOterms) %>% 
-  summarise(mean = mean(cv), size = n()) %>% mutate(GOterms = fct_reorder(GOterms, mean))
-
-ggplot(meanCV, aes(x = GOterms, y = mean, size = log2(size), alpha = 0.2)) + 
-  geom_point() + theme_bw()
-
-topGOnames <- meanCV %>% filter(size > 2) %>% .[order(.$mean, decreasing = F),] %>% 
-  .[1:40,'GOterms'] %>% unlist()
-
-topGO <- unnest(dplyr::select(filter(allGO, GOterms %in% topGOnames), GOterms, cv)) %>%
-  mutate(GO = mapIds(GO.db, keys = GOterms, keytype = 'GOID', column = 'TERM')) %>%
-  group_by(GO) %>% mutate(median = median(cv)) %>% ungroup() %>%
-  mutate(GO = fct_reorder(GO, median))
-
-ggplot(topGO, aes(x = GO, y = cv)) + geom_boxplot(outlier.size = 1) + geom_point() + coord_flip() + scale_x_discrete(labels = function(x) str_wrap(x, width = 50)) + theme_bw() + ylim(0,0.3)
-```
-
-```{r message=FALSE, warning=FALSE}
-allGO[grep('ribosom',allGO$GO),]
-glyc <- filter(allGO, GOterms == 'GO:0005840')$Uniprot %>% unlist()
-
-glycMat <- cor[rownames(cor) %in% glyc,colnames(cor) %in% glyc]
-rownames(glycMat) <- mapIds(org.Xl.eg.db, keys = rownames(glycMat), keytype = 'UNIPROT', column = 'SYMBOL')
-pheatmap(glycMat, show_colnames = T, show_rownames = T, color = mypal, breaks = seq(-1,1, by = 0.01),
-         cutree_cols = 3, cutree_rows = 3)
-
-glyc <- protImpute %>% filter(Uniprot %in% glyc) %>% 
-  mutate(symbol = mapIds(org.Xl.eg.db, keys = Uniprot, keytype = 'UNIPROT', column = 'SYMBOL')) %>%
-  dplyr::select(Uniprot, symbol, everything())
-
-glyc <- glyc %>% gather(key = sample, value = abundance, 3:ncol(glyc)) %>%
-  dplyr::select(sample, symbol, abundance) %>% spread(key = symbol, value = abundance)
-
-#ggpairs(log2(glyc[,2:ncol(glyc),]))
-```
-
-## Cor vs CV
-```{r}
-test = allGO %>% filter(length(Uniprot) > 3, !is.na(medCors), !is.na(medCV))
-
-ggplot(test, aes(x = medCors, y = medCV)) + geom_point(alpha = 0.5) + theme_bw()
-
-ggplot(test, aes(x = medCors, y = medCV)) + 
-  geom_point(alpha = 0.5) + 
-  #stat_smooth(method="lm",formula=  y ~ I(1/x)) +
-  geom_text_repel(data = filter(test, medCors > 0.15), aes(label = GO), colour = 'black', size = 3) + 
-  theme_bw()
-
-test2 = tibble(Cor = test$medCors, CV = test$medCV, mult = test$medCors * test$medCV) %>%
-  gather(key = 'Key', value = 'Value')
-ggplot(test2, aes(x = Value, colour = Key)) + geom_density() + theme_bw()
-ggplot(test2, aes(x = Key, y = Value)) + geom_boxplot() + theme_bw()
-# Variance vs sum of covariance
-# Cor vs Var relationship for every protein pair... error propagation
-```
-
-## All combinations of 5 proteins -> total error vs correlation
-```{r}
-combMat <- combn(1:length(colnames(cor)),2)
-corVect <- tibble(protA = rep(rownames(cor), times = length(colnames(cor))), 
-                  protB = rep(colnames(cor), each = length(rownames(cor))),
-                  keep = as.vector(lower.tri(cor)),
-                  cor = as.vector(cor)) %>% filter(keep) %>% dplyr::select(-keep)
-
-corVect <- corVect %>% dplyr::rename(Uniprot = protA) %>% 
-  left_join(., protVar, 'Uniprot') %>%
-  dplyr::rename(protA = Uniprot, Uniprot = protB, cvA = cv) %>%
-  left_join(., protVar, 'Uniprot') %>% dplyr::rename(protB = Uniprot, cvB = cv) %>%
-  mutate(totVar = sqrt(cvA^2 + cvB^2))
-
-ggplot(corVect, aes(x = cor, y = totVar)) + geom_hex(bins = 250) + theme_bw() +
-  xlim(-1,1) + ylim(0,0.4) + 
-  geom_vline(xintercept = 0.7, colour = 'red', linetype = 'dashed') +
-  geom_hline(yintercept = 0.2, colour = 'red', linetype = 'dashed')
-
-ggplot(filter(corVect, totVar > 0.2, cor > 0.7), aes(x = cor, y = totVar))  + 
-  geom_point(alpha = 0.15, size = 0.1) + theme_bw() + xlim(-1,1) + ylim(0,0.4)
-```
-
-## Total Variance
-```{r}
-protImputeTidy <- protImpute %>% 
-  gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
-  separate(cell, into = c('run','label','condition'), remove = F) %>% 
-  filter(condition != 'ctrl')
-
-### Prototype function
-## Combine the apply find correlation value and pulling out the CV value
-geneList <- allGO$Uniprot[9] %>% unlist()
-protCombs <- t(combn(geneList, 2))
-
-groupVar <- function(x){
-  geneList <- unlist(x)
-  groupVars <- protVar %>% filter(Uniprot %in% geneList)
-  return(sqrt(sum(groupVars$cv^2)))
-}
-
-# Fix by including new code from totVarGroup()
-groupCor <- function(x){
-  geneList <- unlist(x)
-
-  if (length(geneList) > 1) {
-    protCombs <- t(combn(geneList, 2))
-    return(sqrt(sum(totVarPair(protCombs))))
-  } else{
-    return(as.numeric(protVar[protVar$Uniprot == geneList[1], 2]))
-  }
-}
-
-totVarPair <- function(x){
-  A <- as.character(x[1])
-  B <- as.character(x[2])
-  varA <- as.numeric(protVar[protVar$Uniprot == A, 2])
-  varB <- as.numeric(protVar[protVar$Uniprot == B, 2])
-  corAB <- cor[A,B]
-  tot <- varA^2 + varB^2 + 2*corAB*varA*varB
-  return(tot)
-}
-
-totVarGroup <- function(x){
-  geneList <- unlist(x)
-
-  if (length(geneList) > 1) {
-    protCombs <- t(combn(geneList, 2))
-    return(sqrt(sum(apply(protCombs, 1, function(x) totVarPair(x)))))
-    #return(sqrt(sum(totVarPair(protCombs))))
-  } else{
-    return(as.numeric(protVar[protVar$Uniprot == geneList[1], 2]))
-  }
-
-}
-
-allGO1 <- allGO %>% filter(!is.na(GOterms)) %>% rowwise() %>% 
-  mutate(n = length(unlist(Uniprot)),
-         totVar = totVarGroup(Uniprot),
-         normVar = totVar / n) %>%
-  filter(n > 3)
-
-allGO1 %>% arrange(desc(normVar)) %>% print(n = 100)
-allGO1 %>% arrange(normVar) %>% print(n = 100)
-
-allGO1$GO <-factor(allGO1$GO, levels = allGO1$GO[order(allGO1$normVar)])
-
-ggplot(allGO1, aes(x = normVar, y = totVar)) + geom_point()
-ggplot(allGO1, aes(x = log2(n), y = log2(totVar))) + geom_point()
-ggplot(allGO1, aes(x = GO, y = normVar)) + geom_point()
-
-ggplot(allGO1, aes(x = normVar)) + geom_histogram(bins = 30) + 
-  geom_vline(xintercept = c(0.15, 0.1), colour = 'red', linetype = 'dashed') +
-  theme_bw()
-
-```
-
-### Abundance based on intensity
-Does expression level explain lower variance?
-```{r}
-
-ggplot(protRawMean, aes(x = log2(mean), y = noise)) + geom_point() + theme_bw()
-
-```
-
-It doesn't appear to be the case. This could be because these are likely the top ~1000 highest expressed proteins that have a dynamic range of ~250X.
-
-### Move up the GO branch to see how many high/low variance GO terms are from which parent branch
-```{r}
-
-```
-
-### Is lower variance compensating for high correlations? How much of a role is negative correlation playing in reducing the total variance?
-```{r}
-dat <- left_join(ungroup(test), allGO1, by = "GOterms") %>%
-  dplyr::select(GOterms, GO.x, ontology.x, Uniprot.x, cors.x, cv.x, n.x,medCors, medCV, normVar)
-
-dat$score <- 'Med'
-dat[dat$normVar > 0.145, 'score'] <- 'High'
-dat[dat$normVar < 0.115, 'score'] <- 'Low'
-
-ggplot(dat, aes(x = medCors, y = medCV, colour = score)) + geom_point() 
-ggplot(dat, aes(x = medCors, y = medCV)) + geom_point() + geom_smooth(method = 'lm') + facet_grid(~score)
-
-ggplot(filter(dat, n.x > 5), aes(x = medCors, y = medCV, colour = score)) + 
-  geom_point(alpha = 0.75) + scale_color_manual(values=c("blue", "red", "grey"))
-# + #scale_colour_gradient(low="red", high="blue") +
-#  geom_text_repel(data = filter(dat, score != 'Med'), aes(label = GOterms), colour = 'black', size = 1)
-
-ggplot(filter(dat, n.x > 5, score != 'Med', ontology.x != 'CC'), aes(x = medCors, y = medCV, colour = normVar)) + 
-  geom_point() + scale_colour_gradient(low="red", high="blue") + 
-  geom_smooth(method = 'lm') + facet_wrap(~score, scales = 'free') +
-  geom_text_repel(data = filter(dat, n.x > 5, score != 'Med'), aes(label = GO.x), colour = 'black', size = 2)
-
-plot_ly(filter(dat, n.x > 5), z = ~normVar, y = ~medCV, x = ~medCors, color = ~score, colors = c("red", "blue", "grey")) %>%
-  add_markers() %>%
-  layout(scene = list(zaxis = list(title = 'Normalized Variance'),
-                     yaxis = list(title = 'Median CV'),
-                     xaxis = list(title = 'Median Correlation')))
-
-ggplot(filter(dat, n.x > 5), aes(y = normVar, x = medCV)) + geom_point(alpha = 0.75)
-ggplot(filter(dat, n.x > 5), aes(y = normVar, x = medCors)) + geom_point(alpha = 0.75)
-
-library(GGally)
-p <- ggpairs(filter(dat, n.x > 5), columns = c('medCV', 'medCors','normVar'), 
-        ggplot2::aes(colour = score, alpha = 0.5))
-
-for(i in 1:p$nrow) {
-  for(j in 1:p$ncol){
-    p[i,j] <- p[i,j] + 
-        scale_fill_manual(values=c("blue", "red", "grey")) +
-        scale_color_manual(values=c("blue", "red", "grey"))  
-  }
-}
-p
-
-```
-
-## Variance within highly correlated GO terms
-```{r message=FALSE, warning=FALSE}
-allGO[grep('large ribo',allGO$GO),]
-hCor <- filter(allGO, GOterms == 'GO:0015934')$Uniprot %>% unlist()
-
-hCor <- protImpute %>% filter(Uniprot %in% hCor) %>% 
-  mutate(symbol = mapIds(org.Xl.eg.db, keys = Uniprot, keytype = 'UNIPROT', column = 'SYMBOL')) %>%
-  dplyr::select(Uniprot, symbol, everything())
-
-hCor <- hCor %>% gather(key = sample, value = abundance, 3:ncol(hCor)) %>%
-  dplyr::select(sample, symbol, abundance) %>% spread(key = symbol, value = abundance) %>%
-  gather(key = protein, value = abundance, 2:ncol(.))
-
-hCorSD <- hCor %>% group_by(sample) %>% summarise(sd = sd(abundance))
-
-ggplot(hCorSD, aes(x = sample, y = sd)) + geom_point() + geom_hline(yintercept = sd(hCor$abundance))
-
-BC_delta <- function(x) {
-  hCor <- filter(allGO, GOterms == x)$Uniprot %>% unlist()
-
-  hCor <- protImpute %>% filter(Uniprot %in% hCor) %>%
-    mutate(symbol = mapIds(
-      org.Xl.eg.db,
-      keys = Uniprot,
-      keytype = 'UNIPROT',
-      column = 'SYMBOL'
-    )) %>%
-    dplyr::select(Uniprot, symbol, everything())
-
-  hCor <- hCor %>% gather(key = sample, value = abundance, 3:ncol(hCor)) %>%
-    dplyr::select(sample, symbol, abundance)
-
-  hCorSD <-
-    hCor %>% group_by(sample) %>% summarise(sd = sd(abundance))
-
-  return(hCorSD %>% summarise(sum(sd < sd(hCor$abundance)) / n()) %>% as.numeric())
-}
-
-dat <- dat %>% rowwise() %>% mutate(bcDelta = BC_delta(GOterms))
-
-ggplot(dat, aes(x = medCors, y = bcDelta)) + geom_point(alpha = 0.5) + 
-  geom_text_repel(data = filter(test, medCors > 0.2), 
-                    aes(label = GO), colour = 'black', size = 2)
-```

### 2020-03-03 Protein Analysis.Rmd
update

---
title: "Single Cell Xenopus TMT Analysis2"
author: "Kyle Kovary"
date: "5/15/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RColorBrewer)
library(pheatmap)
library(GGally)
library(VIM)
library(org.Xl.eg.db)
library(GO.db)
library(ggrepel)
library(plotly)
library(ComplexHeatmap)

# Import Data
files = list.files('TMTdata/', pattern = 'TargetProtein', full.names = T)

for(i in seq_along(files)){
  if(i == 1 & exists('protRaw')){
    rm(protRaw)
  } 
  if(!exists('protRaw')){
    protRaw = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
  } else{
    temp = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
    colnames(temp) = colnames(protRaw)
    protRaw = rbind(protRaw, temp)
    rm(temp)
  }
}
protRaw <- protRaw %>% dplyr::select(file, massID, everything()) %>% filter(`Confidence A2` != 'Red',
                                                                            Master == 'IsMasterProtein')

```

## Data Organization
```{r }
protRaw = protRaw %>% 
  mutate(run = c('ctrl','a','b','c','d','e')[match(protRaw$file, files)]) %>% 
  dplyr::select(file, run, everything())

protTidy = protRaw %>% 
  gather(key = label, 
         value = intensity, 
         `Ratios: (127) / (126)`:`Ratios: (131) / (126)`) %>% 
  dplyr::select(run, label, intensity, everything(), -file) %>% 
  mutate(condition = c(0,20,40,60,80)[match(label,names(protRaw)[16:20])],
         intensity = as.numeric(intensity)) %>% 
  dplyr::rename(Uniprot = Accession) %>% 
  unite(cell, c('run','label','condition'), remove = F) %>%
  dplyr::select(cell, run, label, condition, Uniprot, intensity)

protTidy <- protTidy[-grep("_",protTidy$Uniprot),]
```

## Data Normalization
```{r }
protTidy <- protTidy %>% group_by(Uniprot) %>% mutate(intensity = intensity / median(intensity, na.rm = TRUE)) %>%
  group_by(cell) %>% mutate(intensity = intensity / median(intensity, na.rm = TRUE)) %>% ungroup()
```

## Impute data
```{r}
protImpute <- protTidy %>% dplyr::select(run,condition,Uniprot,intensity) %>%
  unite(col = "cell", run:condition) %>%
  pivot_wider(names_from = cell, values_from = intensity) %>%
  kNN(data = ., k = 1) %>% as_tibble() %>% dplyr::select(Uniprot:e_80)

protVar <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
  separate(cell, into = c('run','label'), remove = F) %>% 
  filter(run != 'ctrl') %>% group_by(Uniprot) %>%
  summarise(var = var(abundance),
            sd = sd(abundance),
            mean = mean(abundance),
            cv = sd / mean)

ggplot(protVar, aes(x = sd)) + geom_histogram(bins = 50) + theme_bw()
ggplot(protVar, aes(x = mean, y = cv)) + geom_point()
```

## OPTIONAL - Filter out proteins with variance less than control
```{r}
bio_vs_tech_var <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
  separate(cell, into = c('run','label'), remove = F) %>% 
  mutate(bio = ifelse(run == "ctrl",
                      "tech_var",
                      "bio_var")) %>%
  group_by(Uniprot, bio) %>%
  summarise(cv = sd(abundance) / mean(abundance)) %>%
  pivot_wider(names_from = bio, values_from = cv) %>%
  mutate(keep = bio_var > tech_var)

ggplot(bio_vs_tech_var, aes(x = log2(bio_var / tech_var))) + geom_histogram(bins = 50) +
  geom_vline(xintercept = 0)

keep <- bio_vs_tech_var %>% filter(keep) %>% pull(Uniprot)
```

## Variance vs Mean
```{r}
ggplot(protVar, aes(x = log2(mean), y = var/mean^2)) + geom_point() + geom_smooth(method = lm, formula = y ~ poly(x, 2))
```

## Calculate variance and covariance
```{r}
cor = cor(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
colnames(cor) = protImpute$Uniprot
rownames(cor) = protImpute$Uniprot

cov = cov(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
colnames(cov) = protImpute$Uniprot
rownames(cov) = protImpute$Uniprot

mypal = rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201))
k = 2
# map = pheatmap(cor, show_colnames = F, show_rownames = F, color = mypal, breaks = seq(-1,1, by = 0.01), cutree_cols = k, cutree_rows = k, legend = F)
#map

tibble(Uniprot = protImpute$Uniprot, Cluster = cutree(map$tree_row, k = k)) %>% write_csv('/Users/kylekovary/Downloads/Map.csv')
```

## PCA Analysis
```{r}
remove = which(is.na(pairs), arr.ind = T)[,1] %>% unique()
pca = prcomp(protImpute[,2:ncol(protImpute)])

pcaDF = tibble(cell = rownames(pca$rotation), PC1 = pca$rotation[,1], PC2 = pca$rotation[,2])
pcaDF = pcaDF %>% separate(cell, into = c('run','label'), remove = F) %>% print()
ggplot(pcaDF, aes(x = PC1, y = PC2, colour = run)) + geom_point(size = 5)
```

## Variance and Co-variance Between Cells After Cell Normalization

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
protTidy = protTidy %>% group_by(cell) %>% mutate(intensity = intensity / median(intensity, na.rm = T))

# ggplot(protTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-8,8)
```

```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
pairs = protTidy %>% filter(condition == 'ctrl') %>% 
  dplyr::select(cell, peptide, intensity) %>% spread(cell, intensity)

# ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
```

## Variance and Co-variance Between Cells After Peptide Batch Normalization

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
protTidy = protTidy %>% group_by(run, peptide) %>% 
  mutate(intensity = intensity / median(intensity, na.rm = T)) %>% ungroup()

# ggplot(protTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-8,8)
```

```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
pairs = protTidy %>% dplyr::filter(condition == 'ctrl') %>% 
  dplyr::select(cell, peptide, intensity) %>% spread(cell, intensity)

# ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
```

### 2020-03-03 Untitled.Rmd
update

@@
 library(plotly)
+library(cowplot)
+library(patchwork)
 ```
@@
 
-rbind(glyc2,tibble(GO = "All",
+p1 <- rbind(glyc2,tibble(GO = "All",
       cor = cor[lower.tri(cor)])) %>%
-  ggplot(., aes(x = cor, fill = GO, alpha = 0.75)) + geom_density() +
-  geom_rug(data = glyc2)
+  ggplot(.,aes(x = cor, fill = GO)) + geom_density(alpha = 0.75, color = "transparent") +
+  scale_fill_manual(values = c('#bababa','#e08214')) +
+  geom_rug(data = glyc2, color = '#e08214') +
+  theme_bw() +
+  #theme(legend.position = c(.2,.8)) +
+  theme(legend.position = "none") +
+  xlim(-1,1)
+
+p2 <- rbind(glyc2,tibble(GO = "All",
+      cor = cor[lower.tri(cor)])) %>%
+  ggplot(., aes(x = GO, y = cor, fill = GO)) + 
+  geom_boxplot(notch = TRUE, alpha = 0.75) +
+  scale_fill_manual(values = c('#bababa','#e08214')) +
+  theme_bw() +
+  theme(legend.position = "bottom") +
+  ylim(-1,1) +
+  coord_flip()
+
+plot_grid(p1,p2, ncol = 1, 
+          rel_heights = c(2,1),
+          align = 'v')
+
+p1/p2 + 
+  plot_annotation(title = "Glycolysis",
+                  tag_levels = "A") +
+  plot_layout(heights = c(2,1))
 
@@
 # )
+  p1
+```
+
+## Figure 1
+```{r fig.height=6.5, fig.width=6.5}
+set.seed(41)
+
+low_cor <- corProts(2, 100, 0, 100, 30)
+low_cor <- low_cor %>% as_tibble() %>% mutate(cor = "low_cor", time = 1:nrow(low_cor)) %>%
+  rename(ProtA = "V1", ProtB = "V2")
+
+high_cor <- corProts(2, 100, 0.8, 100, 30)
+high_cor <- high_cor %>% as_tibble() %>% mutate(cor = "high_cor", time = 1:nrow(high_cor)) %>%
+  rename(ProtA = "V1", ProtB = "V2")
+
+p1 <- rbind(low_cor, high_cor) %>% pivot_longer(ProtA:ProtB, names_to = "protein", values_to = "abundance") %>%
+  ggplot(., aes(x = time, y = abundance, color = protein)) + geom_path() +
+  facet_wrap(~cor, ncol = 1) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+p2 <- rbind(low_cor, high_cor) %>% pivot_longer(ProtA:ProtB, names_to = "protein", values_to = "abundance") %>%
+  ggplot(., aes(x = protein, y = abundance, color = protein)) + 
+  geom_boxplot() + facet_wrap(~cor, ncol = 1) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+p3 <- rbind(low_cor, high_cor) %>%
+  ggplot(., aes(x = ProtA, y = ProtB, color = cor)) + geom_point() +
+  stat_ellipse() +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+p4 <- tibble(cor = c("High Cor","Low Cor"),
+       var_sum = c(sqrt(30^2 + 30^2 + 2*0.8*30^2),
+                   sqrt(30^2 + 30^2 + 2*0*30^2))) %>%
+  ggplot(., aes(x = cor, y = var_sum, fill = cor)) + geom_bar(stat = "identity") +
+  theme_minimal()
+
+# pp1 <- p1 + p2 + plot_layout(widths = c(3,1))
+# pp2 <- p3 + p4
+# 
+# pp1 / pp2 + plot_annotation(tag_levels = "A") & theme(legend.position = "none")
+
+p1 / (p3 + p4) + plot_annotation(tag_levels = "A") & theme(legend.position = "none")
 

### 2020-03-04 Analysis.Rmd
updated

@@
   B <- as.character(x[2])
-  varA <- as.numeric(protVar[protVar$Uniprot == A, 2])
-  varB <- as.numeric(protVar[protVar$Uniprot == B, 2])
+  varA <- as.numeric(protVar[protVar$Uniprot == A, 3])
+  varB <- as.numeric(protVar[protVar$Uniprot == B, 3])
   corAB <- cor[A,B]

### 2020-03-04 Manuscript/Manuscript.Rmd
updated

@@
   word_document: null
-z
 documentclass: article
@@
 
-library("knitr")
+library(knitr)
+library(tidyverse)
+library(RColorBrewer)
+library(pheatmap)
+library(GGally)
+library(VIM)
+library(org.Xl.eg.db)
+library(GO.db)
+library(ggrepel)
+library(plotly)
+library(cowplot)
+library(patchwork)
+source("../functions.R")
+library(knitcitations)
+
+cleanbib()   
+cite_options(citation_format = "pandoc")
 
@@
 
-```{r knitcitations, echo=FALSE, cache = FALSE}
-library(knitcitations)
-cleanbib()   
-cite_options(citation_format = "pandoc")
+```{r data_import, include=FALSE}
+
+# Import Data
+files = list.files('../TMTdata/', pattern = 'TargetPeptideSpectrumMatch', full.names = T)
+
+for(i in seq_along(files)){
+  if(i == 1 & exists('pepRaw')){
+    rm(pepRaw)
+  } 
+  if(!exists('pepRaw')){
+    pepRaw = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
+  } else{
+    temp = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
+    colnames(temp) = colnames(pepRaw)
+    pepRaw = rbind(pepRaw, temp)
+    rm(temp)
+  }
+}
+pepRaw <- pepRaw %>% dplyr::select(file, massID, everything()) %>% filter(!is.na(`126`), 
+                                                                  `Quan Usage` == 'Use',
+                                                                  Confidence != 'Low')
+
+```
+
+```{r data_organization, include=FALSE}
+pepTidy <- pepRaw %>% 
+  mutate(run = c('ctrl','a','b','c','d','e')[match(pepRaw$file, files)]) %>% 
+  group_by(file, `Annotated Sequence`) %>% 
+  mutate(pepID = letters[1:n()]) %>% ungroup() %>%
+  dplyr::select(file, run, pepID, everything()) %>%
+  gather(key = label, value = intensity, '126':'131') %>% 
+  dplyr::select(run, label, intensity, everything(), -file) %>% 
+  mutate(condition = ifelse(run == 'ctrl', 'ctrl',c('ctrl',0,20,40,60,80)[match(label,c(126:131))]),
+         intensity = as.numeric(intensity)) %>% unite(peptide, c(`Annotated Sequence`,'pepID')) %>%
+  dplyr::rename(Uniprot = `Master Protein Accessions`) %>% 
+  unite(cell, c('run','label','condition'), remove = F) %>%
+  dplyr::select(cell, run, label, condition, peptide, Uniprot, intensity)
+```
+
+```{r normalization, include=FALSE}
+protImpute <- pepTidy %>% 
+  group_by(peptide,run) %>% 
+  mutate(intensity = intensity / intensity[label == "126"]) %>%
+  ungroup() %>%
+  filter(run != "ctrl", condition != "ctrl") %>%
+  dplyr::filter(abs((intensity)) < 2*sd((pepTidy$intensity), na.rm = T)) %>%
+  group_by(cell, run, label, condition, Uniprot) %>% 
+  summarise(abundance = mean(intensity, na.rm = T)) %>% ungroup() %>%
+  group_by(Uniprot) %>%
+  mutate(abundance = abundance / mean(abundance, na.rm = T)) %>% 
+  group_by(cell) %>% 
+  mutate(abundance = abundance / mean(abundance, na.rm = T)) %>% 
+  ungroup() %>% 
+  dplyr::select(cell, Uniprot, abundance) %>% 
+  spread(cell, abundance) %>%
+  kNN(data = ., k = 1, impNA = TRUE) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)
+
+```
+
+```{r summary_stats, include=FALSE}
+protVar <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
+  separate(cell, into = c('run','label','condition'), remove = F) %>% 
+  filter(condition != 'ctrl') %>% group_by(Uniprot) %>%
+  summarise(var = var(abundance),
+            sd = sd(abundance),
+            mean = mean(abundance),
+            cv = sd / mean)
+
+cor = cor(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
+colnames(cor) = protImpute$Uniprot
+rownames(cor) = protImpute$Uniprot
+
+cov = cov(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
+colnames(cov) = protImpute$Uniprot
+rownames(cov) = protImpute$Uniprot
 ```
@@
 
-The coordinated expression of proteins is vital across cellular function, from maintaining a dynamic steady state to differentiation of cells into specialized types in different tissues.
+The coordinated expression of proteins is vital for many aspects of cellular function, from maintaining a dynamic steady state to differentiation of cells into specialized types in different tissues.
 The ability to regulate coordinated expression has been demonstrated to occur via transcription (transcription factors, chromatin regulation), translation (specialized ribosomes, mRNA structure/modifications), and degradation (E3 ligases).
@@
 
-To fetch bibliographic metadata automatically from the web.
-For example, citing a paper can be as easy as providing its DOI `r citep("10.1016/j.tree.2006.03.016")` or even just a few keywords `r citep(c("Heterogeneous ribosomes preferentially translate distinct subpools of mRNAs genome-wide","Controlling low rates of cell differentiation through noise and ultrahigh feedback"))`.
-
 # Results
@@
 
-Activated Xenopus laevis eggs were collected at 5 time points across the first cell cycle (0, 20, 40, 60, and 80 minutes), with 5 eggs at each time point.
+The specific requirments for coordinated protein expression of pathways or complexes can vary, with some requiring strict stoichiometric regulation with others having much more relaxed requirements.
+Strict stoichiometry scenerios require coordinated expression of proteins (Figure 1A, top), and others can have uncoordinated expression (Figure 1A bottom).
+At the single cell level, coordinated expression will result in a high correlation coefficient, whereas un-coordinated expression will result in a low correlation coefficient (Figure 1B).
+Even when the expression variation (standard deviation or coefficient of variation) and abundance is identical between these two scenerios, the total variance of a pathway or complex can be significantly different due to the variance sum law: $var\left(\sum_{i=1}^n X_{i}\right) = \sum_{i=1}^{n} var(X_{i}) + 2\sum_{i,j:i<j}^{n} cov(X_{i},X_{j})$ (Figure 1C).
+Our previous work has shown that this effect is important in single cell pathway activation dynamics `r citep("doi.org/10.15252/msb.20177997")`.
+
+In order to study the relationship between protein expression variance and coexpression (correlated expression) of cellular pathways and complexes at the proteome level, we utilized *Xenopus laevis* eggs as a single cell model.
+Activated *Xenopus laevis* eggs were collected at 5 time points across the first cell cycle (0, 20, 40, 60, and 80 minutes), with 5 eggs at each time point.
 Using TMT multiplexing and mass spectrometry, we were able to determine the relative abundance of more than 1300 proteins.
@@
 
-To determine the variance across these proteins, we calculated the coefficient of variation (CV) using all time points collected.
-This showed a wide range of variation containing multiple distributions (Fig 1B), many of which are consistent with our previous study of variation using targeted mass spectrometry (Fig S2).
+To determine the variance across these proteins, we calculated the variance and coefficient of variation (CV) using all time points collected.
+This showed a wide range of variation containing multiple distributions (Fig 1D), many of which are consistent with our previous study of variation using targeted mass spectrometry (Fig S2).
 In order to see if variation was a regulated process, we grouped the protein variation by gene ontology terms (GO terms) and plotted them in ranked order.
@@
 However, there appear to be a significant number of protein pairs containing high correlation coefficients, and the heatmap showed a lot of clustering of these highly coregulated pairs (Fig 1D).
+
+```{r figure_1, echo=FALSE, fig.height=6.5, fig.width=6.5, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}
+
+set.seed(41)
+
+low_cor <- corProts(2, 100, 0, 100, 30)
+low_cor <- low_cor %>% as_tibble() %>% mutate(cor = "low_cor", time = 1:nrow(low_cor)) %>%
+  rename(ProtA = "V1", ProtB = "V2")
+
+high_cor <- corProts(2, 100, 0.8, 100, 30)
+high_cor <- high_cor %>% as_tibble() %>% mutate(cor = "high_cor", time = 1:nrow(high_cor)) %>%
+  rename(ProtA = "V1", ProtB = "V2")
+
+p1 <- rbind(low_cor, high_cor) %>% pivot_longer(ProtA:ProtB, names_to = "protein", values_to = "abundance") %>%
+  ggplot(., aes(x = time, y = abundance, color = protein)) + geom_path() +
+  facet_wrap(~cor, ncol = 1) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+p2 <- rbind(low_cor, high_cor) %>% pivot_longer(ProtA:ProtB, names_to = "protein", values_to = "abundance") %>%
+  ggplot(., aes(x = protein, y = abundance, color = protein)) + 
+  geom_boxplot() + facet_wrap(~cor, ncol = 1) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+p3 <- rbind(low_cor, high_cor) %>%
+  ggplot(., aes(x = ProtA, y = ProtB, color = cor)) + geom_point() +
+  stat_ellipse() +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+p4 <- tibble(cor = c("High Cor","Low Cor"),
+       var_sum = c(sqrt(30^2 + 30^2 + 2*0.8*30^2),
+                   sqrt(30^2 + 30^2 + 2*0*30^2))) %>%
+  ggplot(., aes(x = cor, y = var_sum, fill = cor)) + geom_bar(stat = "identity") +
+  theme_minimal()
+
+# pp1 <- p1 + p2 + plot_layout(widths = c(3,1))
+# pp2 <- p3 + p4
+# 
+# pp1 / pp2 + plot_annotation(tag_levels = "A") & theme(legend.position = "none")
+
+p1 / (p3 + p4) + plot_annotation(tag_levels = "A") & theme(legend.position = "none")
+
+fig_legend_1 <- c("A) Simulated expression of two proteins over time in a single cell with a mean of 100 and a standard deviation of 30 with expression correlations of 0 and 0.8.
+B) Pairwise plots of the two proteins in the single cells between uncorrelated and correlated scenerios.
+C) The total variance between the uncorrelated and correlated scenerios from A and B.")
+
+```
 
@@
 The sum of variance for a group of proteins $X_{1},\dotsc,X_{n}$ with dependent relationships:
-
-$$var\left(\sum_{i=1}^n X_{i}\right) = \sum_{i=1}^{n} var(X_{i}) + 2\sum_{i,j:i<j}^{n} cov(X_{i},X_{j})$$
 

### 2020-03-04 Manuscript/Manuscript.tex
updated
update

@@
 \usepackage[margin=1in]{geometry}
+\usepackage{graphicx}
+% Redefine \includegraphics so that, unless explicit options are
+% given, the image width will not exceed the width of the page.
+% Images get their normal width if they fit onto the page, but
+% are scaled down if they would overflow the margins.
+\makeatletter
+\def\ScaleIfNeeded{%
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+\let\Oldincludegraphics\includegraphics
+{%
+ \catcode`\@=11\relax%
+ \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
+}%
 \ifxetex
@@
 
-The coordinated expression of proteins is vital across cellular
-function, from maintaining a dynamic steady state to differentiation of
-cells into specialized types in different tissues. The ability to
-regulate coordinated expression has been demonstrated to occur via
-transcription (transcription factors, chromatin regulation), translation
-(specialized ribosomes, mRNA structure/modifications), and degradation
-(E3 ligases). These processes, as well as the molecules they target, are
-all subject to noise, which has been shown to lead to differences in
-cell behavior and decisions in otherwise identical cells. Variability
-and coordinated expression (correlation) of proteins leads to increased
-population level control in binary decisions, and a decreased ability to
-execute analog signaling.
+The coordinated expression of proteins is vital for many aspects of
+cellular function, from maintaining a dynamic steady state to
+differentiation of cells into specialized types in different tissues.
+The ability to regulate coordinated expression has been demonstrated to
+occur via transcription (transcription factors, chromatin regulation),
+translation (specialized ribosomes, mRNA structure/modifications), and
+degradation (E3 ligases). These processes, as well as the molecules they
+target, are all subject to noise, which has been shown to lead to
+differences in cell behavior and decisions in otherwise identical cells.
+Variability and coordinated expression (correlation) of proteins leads
+to increased population level control in binary decisions, and a
+decreased ability to execute analog signaling.
 
@@
 
-To fetch bibliographic metadata automatically from the web. For example,
-citing a paper can be as easy as providing its DOI (Clark and Gelfand,
-2006) or even just a few keywords (Ahrends et al., 2014; Shi et al.,
-2017).
-
 \hypertarget{results}{%
@@
 
-Activated Xenopus laevis eggs were collected at 5 time points across the
-first cell cycle (0, 20, 40, 60, and 80 minutes), with 5 eggs at each
-time point. Using TMT multiplexing and mass spectrometry, we were able
-to determine the relative abundance of more than 1300 proteins.
-Expression of these proteins were largely invariant across the cell
-cycle, revealing that these highly expressed genes are likely not
-regulated by cell cycle processes. Additionally, a PCA analysis of these
-eggs showed no discernible clustering on cell cycle time (Fig S1).
+The specific requirments for coordinated protein expression of pathways
+or complexes can vary, with some requiring strict stoichiometric
+regulation with others having much more relaxed requirements. Strict
+stoichiometry scenerios require coordinated expression of proteins
+(Figure 1A, top), and others can have uncoordinated expression (Figure
+1A bottom). At the single cell level, coordinated expression will result
+in a high correlation coefficient, whereas un-coordinated expression
+will result in a low correlation coefficient (Figure 1B). Even when the
+expression variation (standard deviation or coefficient of variation)
+and abundance is identical between these two scenerios, the total
+variance of a pathway or complex can be significantly different due to
+the variance sum law:
+\(var\left(\sum_{i=1}^n X_{i}\right) = \sum_{i=1}^{n} var(X_{i}) + 2\sum_{i,j:i<j}^{n} cov(X_{i},X_{j})\)
+(Figure 1C). Our previous work has shown that this effect is important
+in single cell pathway activation dynamics (Kovary et al., 2018).
+
+In order to study the relationship between protein expression variance
+and coexpression (correlated expression) of cellular pathways and
+complexes at the proteome level, we utilized \emph{Xenopus laevis} eggs
+as a single cell model. Activated \emph{Xenopus laevis} eggs were
+collected at 5 time points across the first cell cycle (0, 20, 40, 60,
+and 80 minutes), with 5 eggs at each time point. Using TMT multiplexing
+and mass spectrometry, we were able to determine the relative abundance
+of more than 1300 proteins. Expression of these proteins were largely
+invariant across the cell cycle, revealing that these highly expressed
+genes are likely not regulated by cell cycle processes. Additionally, a
+PCA analysis of these eggs showed no discernible clustering on cell
+cycle time (Fig S1).
 
 To determine the variance across these proteins, we calculated the
-coefficient of variation (CV) using all time points collected. This
-showed a wide range of variation containing multiple distributions (Fig
-1B), many of which are consistent with our previous study of variation
-using targeted mass spectrometry (Fig S2). In order to see if variation
-was a regulated process, we grouped the protein variation by gene
-ontology terms (GO terms) and plotted them in ranked order. We saw that
-in general, processes xxx were\ldots{}
+variance and coefficient of variation (CV) using all time points
+collected. This showed a wide range of variation containing multiple
+distributions (Fig 1D), many of which are consistent with our previous
+study of variation using targeted mass spectrometry (Fig S2). In order
+to see if variation was a regulated process, we grouped the protein
+variation by gene ontology terms (GO terms) and plotted them in ranked
+order. We saw that in general, processes xxx were\ldots{}
 
@@
 (Fig 1D).
+
+\begin{figure}
+\centering
+\includegraphics{output/figures/figure_1-1.pdf}
+\caption{A) Simulated expression of two proteins over time in a single
+cell with a mean of 100 and a standard deviation of 30 with expression
+correlations of 0 and 0.8. B) Pairwise plots of the two proteins in the
+single cells between uncorrelated and correlated scenerios. C) The total
+variance between the uncorrelated and correlated scenerios from A and
+B.}
+\end{figure}
 
@@
 
-Variance Sum Law --- Independent Case
-
-\[\sigma_{_{x+y}}^2 = \sigma_{_{x}}^2 + \sigma_{_{y}}^2\]
-
 Variance Sum Law --- Dependent Case
 
-\[\sigma_{_{x+y}}^2 = \sigma_{_{x}}^2 + \sigma_{_{y}}^2 + 2Cov(x,y)\]
-\[\rho = \dfrac{Cov(x,y)}{\sigma_{_{x}} \sigma_{_{y}}}\]
-\[\sigma_{_{x+y}}^2 = \sigma_{_{x}}^2 + \sigma_{_{y}}^2 + 2\rho \sigma_{_{x}} \sigma_{_{y}}\]
+The sum of variance for a group of proteins \(X_{1},\dotsc,X_{n}\) with
+dependent relationships:
 
-\[\sigma_{{}_{total}} = \sqrt{\sum\limits_{i=1}^{n} \; \sigma_{i}^{2} \; \sum\limits_{j\neq i} \; 2\rho_{_{i,j}}\sigma_{_{i}}\sigma_{_{j}}}\]
+\[\sigma_{{}_{total}} = \sqrt{var\left(\sum_{i=1}^n X_{i}\right)}\]
 
-\[\varsigma_{{}_{norm}} = \dfrac{\sigma_{{}_{total}}}{n}\]
+\[\sigma_{{}_{norm}} = \dfrac{\sigma_{{}_{total}}}{n}\]
 
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-ahrends2014controlling}{}%
-Ahrends, R., Ota, A., Kovary, K.M., Kudo, T., Park, B.O., and Teruel,
-M.N. (2014). Controlling low rates of cell differentiation through noise
-and ultrahigh feedback. Science \emph{344}, 1384--1389.
-
-\leavevmode\hypertarget{ref-Clark_2006}{}%
-Clark, J.S., and Gelfand, A.E. (2006). A future for models and data in
-environmental science. Trends in Ecology \& Evolution \emph{21},
-375--380.
-
-\leavevmode\hypertarget{ref-shi2017heterogeneous}{}%
-Shi, Z., Fujii, K., Kovary, K.M., Genuth, N.R., Röst, H.L., Teruel,
-M.N., and Barna, M. (2017). Heterogeneous ribosomes preferentially
-translate distinct subpools of mRNAs genome-wide. Molecular Cell
-\emph{67}, 71--83.e7.
+\leavevmode\hypertarget{ref-Kovary_2018}{}%
+Kovary, K.M., Taylor, B., Zhao, M.L., and Teruel, M.N. (2018).
+Expression variation and covariation impair analog and enable binary
+signaling control. Molecular Systems Biology \emph{14}.
 

### 2020-03-04 functions.R

@@ -1,0 +1,34 @@
+library(mvtnorm)
+
+fun = function(cor_matrix, list_distributions, L)
+{
+  n = length(list_distributions)
+  # Correlated Gaussian variables
+  Gauss = rmvnorm(n=L, mean = rep(0,n), sig=cor_matrix)
+  # convert them to uniform distribution.
+  Unif = pnorm(Gauss) 
+  # Convert them to whatever I want
+  vars = sapply(1:n, FUN = function(i) list_distributions[[i]](Unif[,i]))
+  return(vars)
+}
+
+# Create a matrix of values with a given R, mean, and standard deviation
+corProts <- function(nProt, nCell, R, mean, sd) {
+  
+  cor_matrix =  matrix(data = R,
+                       ncol = nProt,
+                       nrow = nProt)
+  for (i in 1:nProt) {
+    cor_matrix[i, i] = 1
+  }
+  
+  list_distributions <- list()
+  for (i in 1:nProt) {
+    list_distributions <-
+      c(list_distributions, function(nCell)
+        round(qnorm(nCell, mean, sd)))
+  }
+  
+  mat = fun(cor_matrix, list_distributions, nCell)
+  return(mat)
+}

### 2020-03-05 Manuscript/Manuscript.Rmd
updated

@@
 library(RColorBrewer)
-library(pheatmap)
+#library(pheatmap)
+library(ComplexHeatmap)
 library(GGally)
@@
 library(knitcitations)
+library(ggplotify)
 
@@
 
-```{r figure_1, echo=FALSE, fig.height=6.5, fig.width=6.5, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}
+```{r figure_1, echo=FALSE, fig.height=8, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}
 
@@
   theme_minimal() +
-  theme(legend.position = "bottom")
+  theme(legend.position = "none")
 
@@
   ggplot(., aes(x = cor, y = var_sum, fill = cor)) + geom_bar(stat = "identity") +
-  theme_minimal()
+  theme_minimal() +
+  theme(legend.position = "none")
+
+# Variation histogram
+p5 <- ggplot(protVar, aes(x = cv)) + 
+  geom_histogram(bins = 200) + xlim(0,.8) + theme_minimal() +
+  theme(legend.position = "none")
+
+# Correlation matrix
+cor = cor(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
+colnames(cor) = protImpute$Uniprot
+rownames(cor) = protImpute$Uniprot
+
+mypal = rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201))
+
+p6 <- as.ggplot(
+  Heatmap(cor, 
+        col = mypal,
+        heatmap_legend_param = list(title = "R"),
+        show_column_dend = FALSE,
+        show_row_dend = FALSE,
+        show_column_names = FALSE,
+        show_row_names = FALSE,
+        use_raster = TRUE,
+        raster_quality = 1)
+  )
 
@@
 
-p1 / (p3 + p4) + plot_annotation(tag_levels = "A") & theme(legend.position = "none")
+(p1 / (p3 + p4)) / (p5 + p6) + 
+  plot_annotation(tag_levels = "A")
 
@@
 B) Pairwise plots of the two proteins in the single cells between uncorrelated and correlated scenerios.
-C) The total variance between the uncorrelated and correlated scenerios from A and B.")
+C) The total variance between the uncorrelated and correlated scenerios from A and B.
+D) Histogram of the coefficient of variation of all measured proteins.
+E) Heatmap of the pairwise correlation coefficients between all measured protein pairs.")
 

### 2020-03-05 Untitled.Rmd
updated

@@
 ## Plots
-```{r fig.height=8, fig.width=5}
+```{r}
 allGO1 <- allGO1 %>% mutate(interval = cut_number(normVar, n = 10))
 
-ggplot(allGO1, aes(x = normVar)) + geom_histogram(bins = 50)
+ggplot(allGO1, aes(x = normVar)) + geom_histogram(bins = 100, aes(fill = interval))
 
 ggplot(allGO1, 
-       aes(x = medCors, y = medCV, color = interval == "[0.00652,0.0333]")) + 
+       aes(x = medCors, y = medCV, color = interval)) + 
   geom_point(alpha = 0.5) + theme_bw() +
@@
   geom_label_repel(data = filter(allGO1, 
-                                 interval == "[0.00652,0.0333]",
+                                 interval == "[0.142,0.38]",
                                  medCors > 0.15), 
@@
 
+n <- 2
+R <- 0.8
+
+model_tot_var <- function(n,R,v){
+  sqrt(n*v^2 + n*2*R*v^2)
+}
+
 ```

### 2020-03-05 Manuscript/Manuscript.Rmd
updated

@@
 
-mypal = rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201))
+p6 <- tibble(R = as_vector(cor[lower.tri(cor)])) %>%
+  ggplot(., aes(x = R)) + geom_histogram(bins = 1000) + xlim(-1,1) +
+  theme_minimal()
 
-p6 <- as.ggplot(
+p7 <- as.ggplot(
   Heatmap(cor, 
-        col = mypal,
+        col = rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201)),
         heatmap_legend_param = list(title = "R"),
@@
 
-(p1 / (p3 + p4)) / (p5 + p6) + 
+(p1 / (p3 + p4)) / (p5 + p6 | p7) + 
   plot_annotation(tag_levels = "A")

### 2020-03-06 Manuscript/Manuscript.Rmd
updated

@@
 library(ggplotify)
+library(latex2exp)
 
@@
 
+$$var\left(\sum_{i=1}^n X_{i}\right) = \sum_{i=1}^{n} var(X_{i}) + 2\sum_{i,j:i<j}^{n} cov(X_{i},X_{j})$$
+
+$$\sigma_{{}_{total}} = \sum_{i=1}^{n} \sigma_{X_{i}}^2 + 2\sum_{i,j:i<j}^{n} \rho(X_{i},X_{j})\sigma_{X_i}\sigma_{X_j}$$
+
 $$\sigma_{{}_{total}} = \sqrt{var\left(\sum_{i=1}^n X_{i}\right)}$$
@@
 $$\sigma_{{}_{norm}} = \dfrac{\sigma_{{}_{total}}}{n}$$
+
+$$
+R =
+\begin{bmatrix} 
+\rho_{11} & \rho_{12} & \cdots & \rho_{1p} \\
+\rho_{21} & \rho_{22} & \cdots & \rho_{2p} \\
+\vdots & \vdots & \ddots & \vdots \\
+\rho_{p1} & \rho_{p2} & \cdots & \rho_{pp}
+\end{bmatrix}
+$$
+
+```{r}
+library(magick)
+
+cowplot::ggdraw() + cowplot::draw_image(magick::image_read_pdf("../images/preview.pdf", density = 600))
+
+```
+
+$$
+\begin{bmatrix}
+\sigma(X_{1}) \\
+\sigma(X_{1}) \\
+\sigma(X_{1}) \\
+\vdots \\
+\sigma(X_{i}) \\
+\end{bmatrix}^2
++
+\begin{bmatrix} 
+\rho(X_1,X_2) \\
+\rho(X_1,X_3) \\
+\rho(X_1,X_4) \\
+\vdots \\
+\rho(X_i,X_j)
+\end{bmatrix}
+
+\begin{bmatrix}
+\sigma(X_{1}) \\
+\sigma(X_{1}) \\
+\sigma(X_{1}) \\
+\vdots \\
+\sigma(X_{i}) \\
+\end{bmatrix}
+
+\begin{bmatrix}
+\sigma(X_{2}) \\
+\sigma(X_{3}) \\
+\sigma(X_{4}) \\
+\vdots \\
+\sigma(X_{j}) \\
+\end{bmatrix}
+ $$
 

### 2020-03-06 Untitled.Rmd
updated

@@
 
-ggplot(allKEGG, aes(x = corSkew, y = medCV)) + geom_point(aes(color = interval)) +
+ggplot(allKEGG, aes(x = medCors, y = medCV)) + geom_point(aes(color = interval)) +
   geom_label_repel(data = allKEGG, 
@@
 p1/p2 + 
-  plot_annotation(title = "Glycolysis",
-                  tag_levels = "A") +
+  plot_annotation(tag_levels = "A") +
   plot_layout(heights = c(2,1))
@@
 # )
-  p1
 ```

### 2020-03-06 scratch.R

@@ -1,0 +1,51 @@
+
+
+hist(rnorm(10000, 100, 10) / 100)
+hist(log2(rnorm(10000)))
+
+
+a <- c("linear","log")
+
+x <- rnorm(1000, mean = 1, sd = 0.1)
+sd(x) / mean(x)
+sd(x*100) / mean(x*100)
+
+x = rnorm(100,100,10)
+y = x * rnorm(100,100,10)
+
+
+
+#mat <- cor(matrix(rnorm(50), nrow = 5, ncol = 10))
+
+prot_names <- filter(allGO, GOterms == 'GO:0005840')$Uniprot %>% unlist()
+mat <- cor[rownames(cor) %in% prot_names,colnames(cor) %in% prot_names]
+
+colnames(mat) <- NULL
+rownames(mat) <- NULL
+hc = hclust(as.dist(1 - mat))
+mat = mat[hc$order, hc$order]
+mat[lower.tri(mat)] = NA
+diag(mat) <- NA
+
+pheatmap(mat, 
+         cluster_col = F, 
+         cluster_row = F, 
+         color = rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201)), 
+         border_color = NA,
+         na_col = "white",
+         legend = FALSE,
+         breaks = seq(-1,1, by = 0.01))
+
+protImputeTidy %>% filter(Uniprot %in% prot_names) %>%
+  ggplot(., aes(x = abundance, group = Uniprot, alpha = 0.05)) + geom_density(color = "transparent", fill = "grey")
+
+protVar %>% filter(Uniprot %in% prot_names) %>%
+  ggplot(., aes(x = 1, y = Uniprot, fill = sd)) + geom_tile() +
+  coord_equal()
+
+
+pca = prcomp(as.matrix(glyc[,2:ncol(glyc)]))
+
+pcaDF = tibble(PC1 = pca$rotation[,1], PC2 = pca$rotation[,2])
+ggplot(pcaDF, aes(x = PC1, y = PC2)) + geom_point()
+

### 2020-03-10 Manuscript/Manuscript.Rmd
update

@@
 library(latex2exp)
+library(KEGG.db)
 
@@
 rownames(cov) = protImpute$Uniprot
+```
+
+```{r group_prots, include=FALSE}
+# GO Terms
+GO <- protImpute %>% dplyr::select(Uniprot) %>% filter(Uniprot %in% keys) %>% 
+  mutate(GOterms = mapIds(org.Xl.eg.db, 
+                     keys = Uniprot, 
+                     keytype = 'UNIPROT', 
+                     column = 'GOALL', 
+                     multiVals = 'list'))
+
+allGO <- unnest(GO, cols = GOterms) %>% nest(Uniprot = Uniprot) %>%
+  rowwise() %>% mutate(n = length(unlist(Uniprot))) %>% ungroup() %>%
+  mutate(name = mapIds(GO.db, 
+                     keys = GOterms, 
+                     keytype = 'GOID', 
+                     column = 'TERM') %>% as.character(),
+         ontology = mapIds(GO.db, 
+                     keys = GOterms, 
+                     keytype = 'GOID', 
+                     column = 'ONTOLOGY') %>% as.character(),
+         type = "GO") %>%
+  rename(accession = GOterms) %>%
+  dplyr::select(type, accession, name, ontology, everything())
+
+# KEGG Pathways
+kegg <- plyr::ldply(as.list(org.Xl.egPATH2EG),data.frame) %>% as_tibble() %>%
+  rename(kegg_id = ".id", gene_id = "X..i..")
+
+kegg_names <- plyr::ldply(as.list(KEGGPATHNAME2ID),data.frame) %>%
+  rename(kegg_name = ".id",
+         kegg_id = "X..i..")
+
+kegg <- protImpute %>% dplyr::select(Uniprot) %>% filter(Uniprot %in% keys) %>% 
+  mutate(KEGG = mapIds(org.Xl.eg.db, 
+                     keys = Uniprot, 
+                     keytype = 'UNIPROT', 
+                     column = "PATH", 
+                     multiVals = 'list'))
+
+allKEGG <- unnest(kegg, cols = KEGG) %>% nest(Uniprot = Uniprot) %>%
+  rowwise() %>% mutate(n = length(unlist(Uniprot))) %>% ungroup() %>%
+  mutate(name = kegg_names$kegg_name[match(KEGG,kegg_names$kegg_id)],
+         type = "KEGG",
+         ontology = NA) %>% rename(accession = KEGG) %>%
+  dplyr::select(type, accession, name, ontology, Uniprot, n)
+
+allGroups <- rbind(allGO, allKEGG)
+
+# Calculate summary variables
+corVector <- function(uniprotList, corMat){
+  targets = uniprotList %>% unlist()
+  temp = corMat[rownames(corMat) %in% targets,colnames(corMat) %in% targets]
+  temp[lower.tri(temp, diag = F)]
+}
+
+selectCV <- function(x){
+  protVar %>% filter(Uniprot %in% unlist(x)) %>% 
+    dplyr::select(cv) %>% as_tibble() %>% as.list()
+}
+
+allGroups <- allGroups %>% group_by(accession) %>% 
+  dplyr::mutate(cors = corVector(Uniprot, cor) %>% as_tibble() %>% as.list(),
+                covs = corVector(Uniprot, cov) %>% as_tibble() %>% as.list(),
+                cv = selectCV(Uniprot)) %>% ungroup() %>%
+  rowwise() %>% mutate(medCors = mean(cors, na.rm = TRUE),
+                       medCV = mean(cv, na.rm = TRUE)) %>%
+  ungroup() %>%
+filter(n > 3,
+       !is.na(medCors),
+       !is.na(medCV))
+
+```
+
+```{r total_var, include=FALSE}
+groupCor <- function(x){
+  geneList <- unlist(x)
+
+  if (length(geneList) > 1) {
+    protCombs <- t(combn(geneList, 2))
+    return(sqrt(sum(totVarPair(protCombs))))
+  } else{
+    return(as.numeric(protVar[protVar$Uniprot == geneList[1], 2]))
+  }
+}
+
+totVarPair <- function(x){
+  A <- as.character(x[1])
+  B <- as.character(x[2])
+  varA <- as.numeric(protVar[protVar$Uniprot == A, 3])
+  varB <- as.numeric(protVar[protVar$Uniprot == B, 3])
+  corAB <- cor[A,B]
+  tot <- varA^2 + varB^2 + 2*corAB*varA*varB
+  return(tot)
+}
+
+totVarGroup <- function(x){
+  geneList <- unlist(x)
+
+  if (length(geneList) > 1) {
+    protCombs <- t(combn(geneList, 2))
+    return(sqrt(sum(apply(protCombs, 1, function(x) totVarPair(x)))))
+    #return(sqrt(sum(totVarPair(protCombs))))
+  } else{
+    return(as.numeric(protVar[protVar$Uniprot == geneList[1], 2]))
+  }
+
+}
+
+allGroups <- allGroups %>% filter(!is.na(accession)) %>% rowwise() %>%
+  mutate(totVar = totVarGroup(Uniprot),
+         normVar = totVar / n) %>%
+  filter(n > 3) %>% ungroup()
 ```
@@
 When the ranked media correlation is plotted it’s clear that there are a significant number of GO terms with higher than expected levels of coordinate expression.
+
+```{r figure_2,, echo=FALSE, fig.height=8, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_2}
+allGroups <- allGroups %>% mutate(interval = cut_number(normVar, n = 10)) %>%
+  mutate(interval = ifelse(interval == levels(interval)[1],
+                           "low",
+                           ifelse(interval == levels(interval)[10],
+                                  "high",
+                                  "mid")))
+
+ggplot(allGroups, aes(x = normVar)) + geom_histogram(bins = 100, aes(fill = interval)) + theme_minimal() +
+  #scale_fill_viridis_d()
+  scale_fill_brewer(palette = "PiYG")
+
+ggplot(allGroups, aes(x = medCors, y = medCV, color = interval)) + 
+  geom_point() + 
+  theme_minimal()
+
+#fig_legend_2 <- c("A) ")
+```
 

### 2020-03-10 Manuscript/Manuscript.Rmd
.

@@
 ```{r figure_2,, echo=FALSE, fig.height=8, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_2}
-allGroups <- allGroups %>% mutate(interval = cut_number(normVar, n = 10)) %>%
+allGroups <- allGroups %>% mutate(interval = cut_number(normVar, n = 7)) %>%
   mutate(interval = ifelse(interval == levels(interval)[1],
                            "low",
-                           ifelse(interval == levels(interval)[10],
+                           ifelse(interval == levels(interval)[7],
                                   "high",
@@
 
-ggplot(allGroups, aes(x = normVar)) + geom_histogram(bins = 100, aes(fill = interval)) + theme_minimal() +
+p1 <- ggplot(allGroups, aes(x = normVar, fill = interval)) + 
+  geom_histogram(bins = 50) + theme_minimal() +
   #scale_fill_viridis_d()
-  scale_fill_brewer(palette = "PiYG")
+  scale_fill_manual(values = c("#d53e4f","#3288bd","grey")) +
+  xlab("Normalized Variation") + ylab("Number")
 
-ggplot(allGroups, aes(x = medCors, y = medCV, color = interval)) + 
-  geom_point() + 
-  theme_minimal()
+interesting_groups_low <- c("03010","00030","GO:0046496","00071","GO:0006096","GO:0015934","GO:0019843","00051")
+interesting_groups_high <- c("GO:0042995","03008","GO:0007264","GO:0005885","GO:0007166","04150","GO:0035556","GO:0016791","GO:0007165")
+interesting_groups <- c(interesting_groups_low, interesting_groups_high)
 
-#fig_legend_2 <- c("A) ")
+p2 <- ggplot(filter(allGroups, n > 5), aes(x = medCors, y = medCV, color = interval)) + 
+  geom_point(data = filter(allGroups, interval == "mid", n > 5),
+             size = 2, alpha = 0.5, stroke = 0) +
+  geom_point(data = filter(allGroups, interval != "mid", n > 5),
+             size = 3, alpha = 0.75, stroke = 0) +
+  geom_label_repel(data = filter(allGroups, accession %in% interesting_groups), 
+                  aes(label = name), 
+                  colour = 'black', 
+                  size = 2,
+                  force = 100,
+                  box.padding   = 1.5) +
+  theme_minimal() +
+  scale_color_manual(values = c("#d53e4f","#3288bd","grey")) +
+  xlab("Mean Correlation") + ylab("Mean Coefficient of Variation")
+
+p3 <- ggplot(allGroups, aes(x = medCors, fill = interval)) + 
+  annotate("rect", xmin = 0.15, xmax = Inf, ymin = 0, ymax = Inf, fill = "orange", alpha = 0.5) + 
+  geom_density(data = filter(allGroups, interval != "mid"), alpha = 0.75, color = NA) +
+  geom_density(data = filter(allGroups, interval == "mid"), alpha = 0.5, color = NA) +
+  theme_minimal() +
+  scale_fill_manual(values = c("#d53e4f","#3288bd","grey"))
+
+p4 <- ggplot(allGroups, aes(x = medCV, fill = interval)) + 
+  geom_density(alpha = 0.75, color = NA) +
+  theme_minimal() +
+  scale_fill_manual(values = c("#d53e4f","#3288bd","grey"))
+
+(p1 + p3 + p4) / p2 + plot_layout(heights = c(1,5)) +
+  plot_annotation(tag_levels = "A") & 
+  theme(legend.position = "none")
+
+fig_legend_2 <- c("A) Histogram of the normalized variation values across all of the GO terms and KEGG pathways.
+The bottom and top 10% are colored blue and red respectively.
+B) Pairwise plot between the mean correlation and mean coefficient of variaiton within each GO term and KEGG pathway with color coding from A.
+Interestingly, both the top and bottom 10% of normalized variation have groups of proteins with high levels of correlation, despite high correlation adding to total variance.
+In the lowest 10% of normalized variation groups there appears to be a negative correlation between mean correlation and variation.")
+```
+
+```{r figure_3, echo=FALSE, fig.height=8, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_3}
+glyc <- filter(allGroups, accession == '00010')$Uniprot %>% unlist()
+
+glycMat <- cor[rownames(cor) %in% glyc,colnames(cor) %in% glyc]
+rownames(glycMat) <- mapIds(org.Xl.eg.db, keys = rownames(glycMat), keytype = 'UNIPROT', column = 'SYMBOL')
+
+row_anno <- data.frame(cv = unlist(filter(allGroups, accession == "00010") %>% pull(cv)))
+rownames(row_anno) <- rownames(glycMat)
+
+map <- pheatmap(glycMat, show_colnames = T, show_rownames = T, color = mypal, breaks = seq(-1,1, by = 0.01),
+         cutree_cols = 3, cutree_rows = 3, annotation_row = row_anno)
+
+map.clust <- cbind(t(glycMat), cluster = cutree(map$tree_row, k = 3))
+map.clust[,ncol(map.clust) == 2]
+
+glyc2 <- names(which(map.clust[,ncol(map.clust)] == 2))
+totVarGroup(glyc2) / length(glyc2)
+protVar %>% filter(Uniprot %in% glyc2) %>%
+  summarise(meanCV = mean(cv))
+
+glyc2 <- glycMat[which(map.clust[,ncol(map.clust)] == 2),which(map.clust[,ncol(map.clust)] == 2)]
+
+# p7 <- as.ggplot(
+#   Heatmap(cor, 
+#         col = rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201)),
+#         heatmap_legend_param = list(title = "R"),
+#         show_column_dend = FALSE,
+#         show_row_dend = FALSE,
+#         show_column_names = FALSE,
+#         show_row_names = FALSE,
+#         use_raster = TRUE,
+#         raster_quality = 1)
+#   )
+
+p3 <- as.ggplot(
+  pheatmap(glyc2, 
+           color = mypal, 
+           breaks = seq(-1,1, by = 0.01), 
+           border_color = NA, 
+           treeheight_col = 0,
+           treeheight_row = 0,
+           cellheight = 10, 
+           cellwidth = 10,
+           legend = FALSE)
+)
+
+glyc2 <- tibble(GO = "Glycolysis",
+       cor = glyc2[lower.tri(glyc2)])
+
+p1 <- rbind(glyc2,tibble(GO = "All",
+      cor = cor[lower.tri(cor)])) %>%
+  ggplot(.,aes(x = cor, fill = GO)) + geom_density(alpha = 0.75, color = "transparent") +
+  scale_fill_manual(values = c('#bababa','#e08214')) +
+  geom_rug(data = glyc2, color = '#e08214') +
+  theme_bw() +
+  #theme(legend.position = c(.2,.8)) +
+  theme(legend.position = "none") +
+  xlim(-1,1)
+
+p2 <- rbind(glyc2,tibble(GO = "All",
+      cor = cor[lower.tri(cor)])) %>%
+  ggplot(., aes(x = GO, y = cor, fill = GO)) + 
+  geom_boxplot(notch = TRUE, alpha = 0.75) +
+  scale_fill_manual(values = c('#bababa','#e08214')) +
+  theme_bw() +
+  theme(legend.position = "bottom") +
+  ylim(-1,1) +
+  coord_flip()
+
+p4 <- p1/p2 + plot_layout(heights = c(2,1))
+
+fig_legend_3 <- c("A) Histogram of the normalized variation values across all of the GO terms and KEGG pathways.
+The bottom and top 10% are colored blue and red respectively.
+B) Pairwise plot between the mean correlation and mean coefficient of variaiton within each GO term and KEGG pathway with color coding from A.
+Interestingly, both the top and bottom 10% of normalized variation have groups of proteins with high levels of correlation, despite high correlation adding to total variance.
+In the lowest 10% of normalized variation groups there appears to be a negative correlation between mean correlation and variation.")
 ```

### 2020-03-10 Manuscript/Manuscript.Rmd
.

@@
 p1 <- rbind(low_cor, high_cor) %>% pivot_longer(ProtA:ProtB, names_to = "protein", values_to = "abundance") %>%
-  ggplot(., aes(x = time, y = abundance, color = protein)) + geom_path() +
+  ggplot(., aes(x = time, y = abundance, color = protein)) + 
+  #annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, aes(fill = cor)) +
+  #geom_rect(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, aes(fill = cor, color = NA), alpha = 0.002) +
+  geom_path() +
   facet_wrap(~cor, ncol = 1) +
+  scale_color_manual(values = c("#d53e4f","#3288bd")) +
   theme_minimal() +
@@
   ggplot(., aes(x = ProtA, y = ProtB, color = cor)) + geom_point() +
+  scale_color_manual(values = c("#d53e4f","#3288bd")) +
   stat_ellipse() +
@@
   ggplot(., aes(x = cor, y = var_sum, fill = cor)) + geom_bar(stat = "identity") +
+  scale_fill_manual(values = c("#d53e4f","#3288bd")) +
   theme_minimal() +

### 2020-03-11 Manuscript/Manuscript.Rmd
.

--- 2020-03-10 Manuscript/Manuscript.Rmd
@@
 
+selectSD <- function(x){
+  protVar %>% filter(Uniprot %in% unlist(x)) %>% 
+    dplyr::select(sd) %>% as_tibble() %>% as.list()
+}
+
 allGroups <- allGroups %>% group_by(accession) %>% 
@@
                 covs = corVector(Uniprot, cov) %>% as_tibble() %>% as.list(),
-                cv = selectCV(Uniprot)) %>% ungroup() %>%
+                cv = selectCV(Uniprot),
+                sd = selectSD(Uniprot)) %>% ungroup() %>%
   rowwise() %>% mutate(medCors = mean(cors, na.rm = TRUE),
-                       medCV = mean(cv, na.rm = TRUE)) %>%
+                       medCV = mean(cv, na.rm = TRUE),
+                       medSD = mean(sd, na.rm = TRUE)) %>%
   ungroup() %>%
@@
 
-```{r figure_1, echo=FALSE, fig.height=8, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}
+```{r figure_1, echo=FALSE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}
 
@@
 
-p4 <- tibble(cor = c("High Cor","Low Cor"),
+p4 <- tibble(cor = c("High","Low"),
        var_sum = c(sqrt(30^2 + 30^2 + 2*0.8*30^2),
@@
 
+p8 <- cowplot::ggdraw() + cowplot::draw_image(magick::image_read_svg("../images/egg_cartoon.svg"))
+
 # pp1 <- p1 + p2 + plot_layout(widths = c(3,1))
@@
 
-(p1 / (p3 + p4)) / (p5 + p6 | p7) + 
+top <- p1 + p3 + p4 + plot_layout(nrow = 1,
+                                  widths = c(1/3,1/3,1/6),
+                                  heights = c(1,1,1))
+bottom <- p8 + p5 + p6 + p7 + plot_layout(nrow = 1,
+                                          widths = c(2,1,1,3),
+                                heights = c(2,1,1,2))
+
+top / bottom + plot_layout(nrow = 2) +
   plot_annotation(tag_levels = "A")
@@
 C) The total variance between the uncorrelated and correlated scenerios from A and B.
-D) Histogram of the coefficient of variation of all measured proteins.
-E) Heatmap of the pairwise correlation coefficients between all measured protein pairs.")
+D) To measure protein expression variance and coordinated expression, 25 eggs at 5 time points during the first cell cycle were analyzed with multiplexed shotgun mass spectrometry.
+E) Histogram of the coefficient of variation of all measured proteins.
+F) Histogram of the pairwise correlation coefficients between all measured proteins.
+G) Heatmap of the pairwise correlation coefficients between all measured protein pairs.")
 
@@
 ```{r figure_2,, echo=FALSE, fig.height=8, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_2}
-allGroups <- allGroups %>% mutate(interval = cut_number(normVar, n = 10)) %>%
+library(magick)
+
+allGroups <- allGroups %>% mutate(interval = cut_number(normVar, n = 7)) %>%
   mutate(interval = ifelse(interval == levels(interval)[1],
                            "low",
-                           ifelse(interval == levels(interval)[10],
+                           ifelse(interval == levels(interval)[7],
                                   "high",
@@
 
-ggplot(allGroups, aes(x = normVar)) + geom_histogram(bins = 100, aes(fill = interval)) + theme_minimal() +
+p1 <- ggplot(allGroups, aes(x = normVar, fill = interval)) + 
+  geom_histogram(bins = 50) + theme_minimal() +
   #scale_fill_viridis_d()
-  scale_fill_brewer(palette = "PiYG")
+  scale_fill_manual(values = c("#d53e4f","#3288bd","grey")) +
+  xlab("Normalized Variation") + ylab("Number")
 
-ggplot(allGroups, aes(x = medCors, y = medCV, color = interval)) + 
+interesting_groups_low <- c("03010","00030","GO:0046496","00071",
+                            "GO:0006096","GO:0015934","GO:0019843",
+                            "00051","GO:0042555","04012")
+interesting_groups_high <- c("GO:0042995","03008","GO:0007264","GO:0005885","GO:0007166","04150","GO:0035556","GO:0016791","GO:0007165")
+interesting_groups <- c(interesting_groups_low, interesting_groups_high)
+
+p2 <- ggplot(filter(allGroups, n > 5), aes(x = medCors, y = medCV, color = interval)) + 
+  geom_point(data = filter(allGroups, interval == "mid", n > 4),
+             size = 2, alpha = 0.5, stroke = 0) +
+  geom_point(data = filter(allGroups, interval != "mid", n > 4),
+             size = 3, alpha = 0.75, stroke = 0) +
+  geom_label_repel(data = filter(allGroups, accession %in% interesting_groups), 
+                  aes(label = name), 
+                  colour = 'black', 
+                  size = 2,
+                  force = 5,
+                  box.padding   = 1) +
+  theme_minimal() +
+  theme(legend.position = "none") +
+  scale_color_manual(values = c("#d53e4f","#3288bd","grey")) +
+  xlab("Mean Correlation") + ylab("Mean Coefficient of Variation")
+
+p3 <- ggplot(allGroups, aes(x = medCors, fill = interval)) + 
+  annotate("rect", xmin = 0.15, xmax = Inf, ymin = 0, ymax = Inf, fill = "orange", alpha = 0.5) + 
+  geom_density(data = filter(allGroups, interval != "mid"), alpha = 0.75, color = NA) +
+  geom_density(data = filter(allGroups, interval == "mid"), alpha = 0.5, color = NA) +
+  theme_minimal() +
+  scale_fill_manual(values = c("#d53e4f","#3288bd","grey"))
+
+p4 <- ggplot(allGroups, aes(x = medCV, fill = interval)) + 
+  geom_density(alpha = 0.75, color = NA) +
+  theme_minimal() +
+  scale_fill_manual(values = c("#d53e4f","#3288bd","grey"))
+
+p5 <- rbind(mutate(allGroups, interval = "mid", name = "all"),
+      filter(allGroups, accession %in% interesting_groups)) %>%
+  select(-Uniprot,-covs,-cv,-type,-ontology) %>%
+  unnest(cols = "cors") %>%
+  ggplot(., aes(x = fct_reorder(name,rank(interval)), y = cors, fill = interval)) + 
+  geom_boxplot(notch = TRUE, alpha = 0.75) +
+  scale_fill_manual(values = c("#d53e4f","#3288bd","grey")) +
+  coord_flip() + 
+  scale_x_discrete(labels = function(x) str_wrap(x, width = 15)) +
+  theme_minimal() +
+  theme(legend.position = "none",
+        text = element_text(size = 8),
+        axis.title.y=element_blank())
+
+p6 <- p2 + p5 + plot_layout(widths = c(1,0.5))
+
+(p1 + p3 + p4) / p6 + plot_layout(heights = c(1,4)) +
+  plot_annotation(tag_levels = "A") & 
+  theme(legend.position = "none")
+
+fig_legend_2 <- c("A) Histogram of the normalized variation values across all of the GO terms and KEGG pathways.
+The bottom and top 10% are colored blue and red respectively.
+B) Pairwise plot between the mean correlation and mean coefficient of variaiton within each GO term and KEGG pathway with color coding from A.
+Interestingly, both the top and bottom 10% of normalized variation have groups of proteins with high levels of correlation, despite high correlation adding to total variance.
+In the lowest 10% of normalized variation groups there appears to be a negative correlation between mean correlation and variation.")
+```
+
+```{r sup_fig_1, echo=FALSE, fig.height=2, fig.width=6, message=FALSE, warning=FALSE, fig.cap=sup_fig_legend_1}
+allGroups %>% filter(n > 4) %>%
+  ggplot(., aes(x = medCors, y = medCV, color = interval)) + 
   geom_point() + 
+  scale_color_manual(values = c("#d53e4f","#3288bd","grey")) +
+  stat_summary(fun.data=mean_cl_normal) + 
+  geom_smooth(method='lm', formula= y~x, color = "black") +
+  facet_wrap(~interval, scales = "free", nrow = 1) +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+sup_fig_legend_1 <- c("Pairwise plots of the mean correlation coefficient and mean coefficient of variation of protien groups with linear fit.
+The low variance protein groups have a significant negative correlation between these two variables, showing that for protien modules that require low variance and high coexpression, showing that these two parameters...
+R (high, low, med) = (0.057, -0.485, 0.071)")
+```
+
+```{r figure_3, echo=FALSE, fig.height=8, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_3}
+library(circlize)
+
+allGroups[grep("large ribo",allGroups$name),]
+prots <- allGroups %>% filter(accession == "GO:0015934") %>% pull(Uniprot) %>% unlist() %>% as.vector()
+mat <- cor[prots,prots]
+rownames(mat) <- mapIds(org.Xl.eg.db, keys = rownames(mat), keytype = 'UNIPROT', column = 'SYMBOL')
+
+col_fun = colorRamp2(seq(-1,1,0.01), rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201)))
+
+p1 <- as.ggplot(
+  Heatmap(
+    mat,
+    column_title = "Large ribosomal subunit",
+    column_title_gp = gpar(fontsize = 20, fontface = "bold"),
+    col = col_fun,
+    heatmap_legend_param = list(title = "R"),
+    show_column_dend = TRUE,
+    show_row_dend = TRUE,
+    show_column_names = TRUE,
+    show_row_names = TRUE,
+    use_raster = FALSE,
+    cell_fun = function(j, i, x, y, width, height, fill) {
+        grid.text(sprintf("%.1f", mat[i, j]), x, y, gp = gpar(fontsize = 10))}
+  )
+)
+
+# Plot of max number of complete complexes
+cors <- seq(0,1,0.02)
+vars <- rev(seq(0,1,0.02))
+
+corVarMat <- matrix(data = NA, ncol = length(cors), nrow = length(vars))
+colnames(corVarMat) <- cors
+rownames(corVarMat) <- vars
+
+# Percent of potentially complete complexes
+
+corVarMat <- sapply(1:length(cors), function(i) {
+  sapply(1:length(vars), function(j) {
+    perc <- corProts(nProt = 10, nCell = 10000, R = cors[i], mean = 1000, sd = 1000*vars[j])
+    mean(apply(perc, 1, function(x) min(abs(x)) / max(abs(x))), na.rm = T)
+    #corVarMat[j,i] <- perc
+  })
+})
+
+x <- cbind(vars, corVarMat)
+colnames(x) <- c("vars",as.character(cors))
+x <- x %>% as_tibble() %>% gather(key = R, value = percent, 2:ncol(x)) %>% mutate(R = as.numeric(R)) %>%
+  rename_(CV = vars)
+
+p2 <- ggplot(x, aes(x = CV, y = percent, colour = R, group = as.factor(R))) + 
+  geom_path() + 
+  scale_x_reverse() + 
+  scale_color_viridis_c() +
   theme_minimal()
 
-#fig_legend_2 <- c("A) ")
+cors <- seq(0,1,0.01)
+vars <- rev(seq(0,0.5,0.005))
+
+totVarMat <- sapply(1:length(cors), function(i) {
+  sapply(1:length(vars), function(j) {
+    (10*vars[j]^2 + 10*2*cors[i]*vars[j]) / 10
+    #corVarMat[j,i] <- perc
+  })
+})
+
+rownames(totVarMat) <- vars
+colnames(totVarMat) <- cors
+
+col_fun <- colorRamp2(seq(range(totVarMat)[1],
+                         range(totVarMat)[2],
+                         range(totVarMat)[2] / 150), 
+                     wesanderson::wes_palette("Zissou1", 
+                                              n = 151, 
+                                              type = "continuous"))
+
+Heatmap(totVarMat,
+        col = col_fun,
+        cluster_rows = FALSE,
+        cluster_columns = FALSE,
+        border = NA,
+        column_title = "R",
+        row_title = "CV",
+        heatmap_legend_param = list(title = "Norm Var"),
+        row_names_gp = gpar(fontsize = 5),
+        column_names_gp = gpar(fontsize = 5))
+
+fig_legend_3 <- c("A) Histogram of the normalized variation values across all of the GO terms and KEGG pathways.
+The bottom and top 10% are colored blue and red respectively.
+B) Pairwise plot between the mean correlation and mean coefficient of variaiton within each GO term and KEGG pathway with color coding from A.
+Interestingly, both the top and bottom 10% of normalized variation have groups of proteins with high levels of correlation, despite high correlation adding to total variance.
+In the lowest 10% of normalized variation groups there appears to be a negative correlation between mean correlation and variation.")
 ```
@@
 $$
-R =
 \begin{bmatrix} 
-\rho_{11} & \rho_{12} & \cdots & \rho_{1p} \\
-\rho_{21} & \rho_{22} & \cdots & \rho_{2p} \\
-\vdots & \vdots & \ddots & \vdots \\
-\rho_{p1} & \rho_{p2} & \cdots & \rho_{pp}
+\ \sigma(X_{1})^2 +  \rho(X_1,X_2) \cdot \sigma(X_{1}) \cdot \sigma(X_{2}) \\
+\sigma(X_{1})^2 + \rho(X_1,X_3) \cdot \sigma(X_{1}) \cdot \sigma(X_{3}) \\
+\sigma(X_{1})^2 + \rho(X_1,X_4) \cdot \sigma(X_{1}) \cdot \sigma(X_{4}) \\
+\vdots \\
+\sigma(X_{i})^2 + \rho(X_i,X_j) \cdot \sigma(X_{i}) \cdot \sigma(X_{j}) \\
 \end{bmatrix}
  $$
-
-```{r}
-library(magick)
-
-cowplot::ggdraw() + cowplot::draw_image(magick::image_read_pdf("../images/preview.pdf", density = 600))
-
-```
 

### 2020-03-11 Modeling.Rmd
.

@@
                        nrow = nProt)
-  for (i in 1:nProt) {
-    cor_matrix[i, i] = 1
-  }
 
+  diag(cor_matrix) <- 1
+  # for (i in 1:nProt) {
+  #   cor_matrix[i, i] = 1
+  # }
+  # 
   list_distributions <- list()
@@
 # Percent of potentially complete complexes
-for(i in 1:length(cors)){
-  for(j in 1:length(vars)){
+
+corVarMat <- sapply(1:length(cors), function(i) {
+  sapply(1:length(vars), function(j) {
     perc <- corProts(nProt = 10, nCell = 10000, R = cors[i], mean = 1000, sd = 1000*vars[j])
-    perc <- mean(apply(perc, 1, function(x) min(abs(x)) / max(abs(x))), na.rm = T)
-    corVarMat[j,i] <- perc
-  }
-}
+    mean(apply(perc, 1, function(x) min(abs(x)) / max(abs(x))), na.rm = T)
+    #corVarMat[j,i] <- perc
+  })
+})
+
 pheatmap(corVarMat, cluster_cols = F, cluster_rows = F)
@@
 pheatmap(rotate(corVarMat), cluster_cols = F, cluster_rows = F)
-plot(diag(rotate(corVarMat)), rotate(corVarMat)[])
+#plot(diag(rotate(corVarMat)), rotate(corVarMat)[])
 
@@
 
-x <- as_tibble(cbind(vars, corVarMat))
-x <- x %>% gather(key = R, value = percent, 2:ncol(x)) %>% mutate(R = as.numeric(R)) %>%
+x <- cbind(vars, corVarMat)
+colnames(x) <- c("vars",as.character(cors))
+x <- x %>% as_tibble() %>% gather(key = R, value = percent, 2:ncol(x)) %>% mutate(R = as.numeric(R)) %>%
   rename_(CV = vars)
-ggplot(x, aes(x = CV, y = percent, colour = R, group = as.factor(R))) + geom_path() + scale_x_reverse() + theme_bw()
-
-ggplot(x, aes(x = CV, y = percent, colour = as.factor(R))) + geom_path() + scale_x_reverse(limits = c(0.2,0)) + theme_bw()
+ggplot(x, aes(x = CV, y = percent, colour = R, group = as.factor(R))) + 
+  geom_path() + 
+  scale_x_reverse() + 
+  scale_color_viridis_c() +
+  theme_minimal()
 

### 2020-03-11 functions.R

@@ -18,10 +18,8 @@ corProts <- function(nProt, nCell, R, mean, sd) {
   cor_matrix =  matrix(data = R,
                        ncol = nProt,
                        nrow = nProt)
-  for (i in 1:nProt) {
-    cor_matrix[i, i] = 1
-  }
   
+  diag(cor_matrix) <- 1
   list_distributions <- list()
   for (i in 1:nProt) {
     list_distributions <-
@@ -31,4 +29,44 @@ corProts <- function(nProt, nCell, R, mean, sd) {
   
   mat = fun(cor_matrix, list_distributions, nCell)
   return(mat)
+}
+
+# Rug Plots
+cor_rug_plot <- function(x, color = '#e08214'){
+  temp <- filter(allGroups, accession %in% x) %>%
+    dplyr::select(name, cors) %>% unnest(cols = cors)
+  
+  temp <- allGroups %>% select(name, cors) %>%
+    unnest(cols = cors) %>% mutate(name = "all") %>%
+    rbind(temp,.)
+  
+  p1 <- ggplot(temp, aes(x = cors, fill = name)) + geom_density(alpha = 0.75, color = "transparent") +
+    scale_fill_manual(values = c('#bababa',color), name = "") +
+    geom_rug(data = filter(temp, name != "all"), color = color) +
+    xlab("") + ylab("Density") +
+    theme_bw() +
+    # theme(legend.position = c(.2,.9),
+    #       legend.background = element_rect(fill="transparent")) +
+    theme(legend.position = "none",
+          text = element_text(size = 6)) +
+    xlim(-1,1) +
+    ggtitle(filter(temp, name != "all") %>% pull(name) %>% unique())
+  
+  p2 <- ggplot(temp, aes(x = name, y = cors, fill = name)) + 
+    geom_boxplot(notch = TRUE, alpha = 0.75) +
+    scale_fill_manual(values = c('#bababa',color)) +
+    ylab("Correlation Coefficients") +
+    theme_bw() +
+    theme(legend.position = "none",
+          text = element_text(size = 6),
+          axis.title.y=element_blank(),
+          axis.text.y=element_blank(),
+          axis.ticks.y=element_blank()) +
+    ylim(-1,1) +
+    coord_flip()
+  
+  # p1/p2 + 
+  #   #plot_annotation(tag_levels = "A") +
+  #   plot_layout(heights = c(2,1))
+  plot_grid(p1, p2, ncol = 1, rel_heights = c(2,1), align = "v")
 }

### 2020-03-11 Manuscript/Manuscript.Rmd
.

@@
 fontsize: 11pt
+header-includes: 
+  \usepackage{float} \floatplacement{figure}{H} 
+  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
 spacing: doublespacing
@@
 library(KEGG.db)
+library(magick)
 
@@
 
-```{r group_prots, include=FALSE}
+```{r group_prots, include=FALSE, eval=FALSE}
 # GO Terms
-GO <- protImpute %>% dplyr::select(Uniprot) %>% filter(Uniprot %in% keys) %>% 
-  mutate(GOterms = mapIds(org.Xl.eg.db, 
+keys <- keys(org.Xl.eg.db, keytype = 'UNIPROT')
+
+GO <- protImpute %>% dplyr::select(Uniprot) %>% dplyr::filter(Uniprot %in% keys) %>% 
+  dplyr::mutate(GOterms = mapIds(org.Xl.eg.db, 
                      keys = Uniprot, 
@@
 allKEGG <- unnest(kegg, cols = KEGG) %>% nest(Uniprot = Uniprot) %>%
-  rowwise() %>% mutate(n = length(unlist(Uniprot))) %>% ungroup() %>%
-  mutate(name = kegg_names$kegg_name[match(KEGG,kegg_names$kegg_id)],
+  rowwise() %>% dplyr::mutate(n = length(unlist(Uniprot))) %>% ungroup() %>%
+  dplyr::mutate(name = kegg_names$kegg_name[match(KEGG,kegg_names$kegg_id)],
          type = "KEGG",
@@
 
-```{r total_var, include=FALSE}
+```{r total_var, include=FALSE, eval=FALSE}
 groupCor <- function(x){
@@
   filter(n > 3) %>% ungroup()
+
+allGroups <- allGroups %>% mutate(interval = cut_number(normVar, n = 7)) %>%
+  mutate(interval = ifelse(interval == levels(interval)[1],
+                           "low",
+                           ifelse(interval == levels(interval)[7],
+                                  "high",
+                                  "mid")))
+
+saveRDS(allGroups,"allGroups.RDS")
 ```
@@
 ```{r figure_1, echo=FALSE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}
+allGroups <- readRDS("allGroups.RDS")
 
@@
 
-```{r figure_2,, echo=FALSE, fig.height=8, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_2}
-library(magick)
-
-allGroups <- allGroups %>% mutate(interval = cut_number(normVar, n = 7)) %>%
-  mutate(interval = ifelse(interval == levels(interval)[1],
-                           "low",
-                           ifelse(interval == levels(interval)[7],
-                                  "high",
-                                  "mid")))
-
+```{r figure_2, echo=FALSE, fig.height=7, fig.width=6, message=FALSE, warning=FALSE, eval=TRUE, fig.cap=fig_legend_2}
 p1 <- ggplot(allGroups, aes(x = normVar, fill = interval)) + 
@@
                             "00051","GO:0042555","04012")
-interesting_groups_high <- c("GO:0042995","03008","GO:0007264","GO:0005885","GO:0007166","04150","GO:0035556","GO:0016791","GO:0007165")
+
+interesting_groups_high <- c("GO:0042995","03008","GO:0007264","GO:0005885",
+                             "GO:0007166","04150","GO:0035556","GO:0016791","GO:0007165")
+
 interesting_groups <- c(interesting_groups_low, interesting_groups_high)
@@
                   colour = 'black',
-                  size = 2,
+                  size = 1.5,
                   force = 5,
                   box.padding = 1) +
+  #coord_cartesian(ylim=c(-10,10)) +
   theme_minimal() +
@@
 
-p5 <- rbind(mutate(allGroups, interval = "mid", name = "all"),
-      filter(allGroups, accession %in% interesting_groups)) %>%
-  select(-Uniprot,-covs,-cv,-type,-ontology) %>%
-  unnest(cols = "cors") %>%
-  ggplot(., aes(x = fct_reorder(name,rank(interval)), y = cors, fill = interval)) + 
-  geom_boxplot(notch = TRUE, alpha = 0.75) +
-  scale_fill_manual(values = c("#d53e4f","#3288bd","grey")) +
-  coord_flip() + 
-  scale_x_discrete(labels = function(x) str_wrap(x, width = 15)) +
-  theme_minimal() +
-  theme(legend.position = "none",
-        text = element_text(size = 8),
-        axis.title.y=element_blank())
+# p5 <- rbind(mutate(allGroups, interval = "mid", name = "all"),
+#       filter(allGroups, accession %in% interesting_groups)) %>%
+#   dplyr::select(-Uniprot,-covs,-cv,-type,-ontology) %>%
+#   unnest(cols = "cors") %>%
+#   ggplot(., aes(x = fct_reorder(name,rank(interval)), y = cors, fill = interval)) +
+#   #geom_boxplot(notch = FALSE, alpha = 0.75, outlier.size = 0) +
+#   geom_violin() +
+#   scale_fill_manual(values = c("#d53e4f","#3288bd","grey")) +
+#   coord_flip() +
+#   scale_x_discrete(labels = function(x) str_wrap(x, width = 15)) +
+#   theme_minimal() +
+#   theme(text = element_text(size = 6),
+#         axis.title.y=element_blank())
 
-p6 <- p2 + p5 + plot_layout(widths = c(1,0.5))
+p6 <- p2 + plot_spacer() + plot_layout(widths = c(1,0.5))
 
-(p1 + p3 + p4) / p6 + plot_layout(heights = c(1,4)) +
+(p1 + p3 + p4) / p6 + plot_layout(heights = c(1,5)) +
   plot_annotation(tag_levels = "A") &
   theme(legend.position = "none")
+
+#p1 + p2 + p3 + plot_annotation(tag_levels = "A")
 
@@
 In the lowest 10% of normalized variation groups there appears to be a negative correlation between mean correlation and variation.")
+
 ```
 
-```{r sup_fig_1, echo=FALSE, fig.height=2, fig.width=6, message=FALSE, warning=FALSE, fig.cap=sup_fig_legend_1}
-allGroups %>% filter(n > 4) %>%
-  ggplot(., aes(x = medCors, y = medCV, color = interval)) + 
-  geom_point() + 
-  scale_color_manual(values = c("#d53e4f","#3288bd","grey")) +
-  stat_summary(fun.data=mean_cl_normal) + 
-  geom_smooth(method='lm', formula= y~x, color = "black") +
-  facet_wrap(~interval, scales = "free", nrow = 1) +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-sup_fig_legend_1 <- c("Pairwise plots of the mean correlation coefficient and mean coefficient of variation of protien groups with linear fit.
-The low variance protein groups have a significant negative correlation between these two variables, showing that for protien modules that require low variance and high coexpression, showing that these two parameters...
-R (high, low, med) = (0.057, -0.485, 0.071)")
-```
-
-```{r figure_3, echo=FALSE, fig.height=8, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_3}
+```{r figure_3, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, eval=TRUE, fig.cap=fig_legend_3}
 library(circlize)
 
-allGroups[grep("large ribo",allGroups$name),]
+#allGroups[grep("large ribo",allGroups$name),]
 prots <- allGroups %>% filter(accession == "GO:0015934") %>% pull(Uniprot) %>% unlist() %>% as.vector()
@@
 rownames(mat) <- mapIds(org.Xl.eg.db, keys = rownames(mat), keytype = 'UNIPROT', column = 'SYMBOL')
-
-col_fun = colorRamp2(seq(-1,1,0.01), rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201)))
 
@@
     column_title = "Large ribosomal subunit",
-    column_title_gp = gpar(fontsize = 20, fontface = "bold"),
-    col = col_fun,
+    column_title_gp = gpar(fontsize = 5, fontface = "bold"),
+    col = colorRamp2(seq(-1,1,0.01), rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201))),
     heatmap_legend_param = list(title = "R"),
-    show_column_dend = TRUE,
-    show_row_dend = TRUE,
+    row_names_gp = gpar(fontsize = 4),
+    column_names_gp = gpar(fontsize = 4),
+    show_column_dend = FALSE,
+    show_row_dend = FALSE,
     show_column_names = TRUE,
     show_row_names = TRUE,
-    use_raster = FALSE,
-    cell_fun = function(j, i, x, y, width, height, fill) {
-        grid.text(sprintf("%.1f", mat[i, j]), x, y, gp = gpar(fontsize = 10))}
+    #cell_fun = function(j, i, x, y, width, height, fill) {grid.text(sprintf("%.1f", mat[i, j]), x, y, gp = gpar(fontsize = 1))},
+    use_raster = FALSE
   )
@@
 
-Heatmap(totVarMat,
+p3 <- as.ggplot(Heatmap(totVarMat,
         col = col_fun,
@@
         cluster_columns = FALSE,
+        show_column_names = FALSE,
+        show_row_names = FALSE,
         border = NA,
@@
         row_names_gp = gpar(fontsize = 5),
-        column_names_gp = gpar(fontsize = 5))
+        column_names_gp = gpar(fontsize = 5)))
+
+bottom <- p2 + p3 + plot_layout(widths = c(1,2))
+(p1 + plot_spacer()) / bottom + 
+  plot_annotation(tag_levels = "A")
 
@@
 $$\sigma_{{}_{norm}} = \dfrac{\sigma_{{}_{total}}}{n}$$
-
-$$
-\begin{bmatrix} 
-\ \sigma(X_{1})^2 +  \rho(X_1,X_2) \cdot \sigma(X_{1}) \cdot \sigma(X_{2}) \\
-\sigma(X_{1})^2 + \rho(X_1,X_3) \cdot \sigma(X_{1}) \cdot \sigma(X_{3}) \\
-\sigma(X_{1})^2 + \rho(X_1,X_4) \cdot \sigma(X_{1}) \cdot \sigma(X_{4}) \\
-\vdots \\
-\sigma(X_{i})^2 + \rho(X_i,X_j) \cdot \sigma(X_{i}) \cdot \sigma(X_{j}) \\
-\end{bmatrix}
- $$
-
-$$
-\begin{bmatrix}
-\sigma(X_{1}) \\
-\sigma(X_{1}) \\
-\sigma(X_{1}) \\
-\vdots \\
-\sigma(X_{i}) \\
-\end{bmatrix}^2
-+
-\begin{bmatrix} 
-\rho(X_1,X_2) \\
-\rho(X_1,X_3) \\
-\rho(X_1,X_4) \\
-\vdots \\
-\rho(X_i,X_j)
-\end{bmatrix}
-
-\begin{bmatrix}
-\sigma(X_{1}) \\
-\sigma(X_{1}) \\
-\sigma(X_{1}) \\
-\vdots \\
-\sigma(X_{i}) \\
-\end{bmatrix}
-
-\begin{bmatrix}
-\sigma(X_{2}) \\
-\sigma(X_{3}) \\
-\sigma(X_{4}) \\
-\vdots \\
-\sigma(X_{j}) \\
-\end{bmatrix}
- $$
 
@@
 
+\newpage
+# Supplementary material {-}
+
+\beginsupplement
+```{r sup_fig_1, echo=FALSE, fig.height=2, fig.width=6, message=FALSE, warning=FALSE, fig.cap=sup_fig_legend_1}
+allGroups %>% filter(n > 4) %>%
+  ggplot(., aes(x = medCors, y = medCV, color = interval)) + 
+  geom_point() + 
+  scale_color_manual(values = c("#d53e4f","#3288bd","grey")) +
+  stat_summary(fun.data=mean_cl_normal) + 
+  geom_smooth(method='lm', formula= y~x, color = "black") +
+  facet_wrap(~interval, scales = "free", nrow = 1) +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+sup_fig_legend_1 <- c("Pairwise plots of the mean correlation coefficient and mean coefficient of variation of protien groups with linear fit.
+The low variance protein groups have a significant negative correlation between these two variables, showing that for protien modules that require low variance and high coexpression, showing that these two parameters...
+R (high, low, med) = (0.057, -0.485, 0.071)")
+```
+
 \eleft

### 2020-03-11 Manuscript/Manuscript.Rmd
.

@@
 The coordinated expression of proteins is vital for many aspects of cellular function, from maintaining a dynamic steady state to differentiation of cells into specialized types in different tissues.
+These requirements can range from precise stoichiometric coordination for multisubunit complexes and metabolic pathways, to high levels of noise that can result in heterogeneous responses of identical cell populations to identical stimuli `r citep("10.1073/pnas.1615660114")`.
 The ability to regulate coordinated expression has been demonstrated to occur via transcription (transcription factors, chromatin regulation), translation (specialized ribosomes, mRNA structure/modifications), and degradation (E3 ligases).
-These processes, as well as the molecules they target, are all subject to noise, which has been shown to lead to differences in cell behavior and decisions in otherwise identical cells.
+These processes, as well as the molecules they target, are all subject to stochastic expression and interactions, which has been shown to lead to differences in cell behavior and decisions in otherwise identical cells.
 Variability and coordinated expression (correlation) of proteins leads to increased population level control in binary decisions, and a decreased ability to execute analog signaling.
 
-Despite these important insights, a global understanding of the variability of protein expression between single cells and the coordinated expression of groups of proteins is largely an nonexistent.
+There have been numerous targeted studies of coordinated protein expression and variation that have shown the important impacts of these parameters on cellular function.
+However, both of these parameters (variation and correlation) act together to determine the total variance of a system.
+
+there is a gap in the literature of direct measurement of the relationship between variation and coordinated protein expression between single cells.
 To systematically assess these properties of single cells, we carried out a proteomics experiment on exceeding large single cells, Xenopus laevis eggs.
@@
 
-With this dataset we have been able to, for the first time, measured the relationship between protein expression variance and coordinated protein expression on a proteome scale in single cells.
+With this data set we have been able to, for the first time, measured the relationship between protein expression variance and coordinated protein expression on a proteome scale in single cells.
 We have observed that certain classes of proteins, including protein complexes and metabolic pathways, are expressed in such a way that increased coordinated expression is balanced by decreased variation.
@@
 
-The specific requirments for coordinated protein expression of pathways or complexes can vary, with some requiring strict stoichiometric regulation with others having much more relaxed requirements.
-Strict stoichiometry scenerios require coordinated expression of proteins (Figure 1A, top), and others can have uncoordinated expression (Figure 1A bottom).
+The specific requirements for coordinated protein expression of pathways or complexes can vary, with some requiring strict stoichiometric regulation with others having much more relaxed requirements.
+Strict stoichiometry scenarios require coordinated expression of proteins (Figure 1A, top), and others can have uncoordinated expression (Figure 1A bottom).
 At the single cell level, coordinated expression will result in a high correlation coefficient, whereas un-coordinated expression will result in a low correlation coefficient (Figure 1B).
-Even when the expression variation (standard deviation or coefficient of variation) and abundance is identical between these two scenerios, the total variance of a pathway or complex can be significantly different due to the variance sum law: $var\left(\sum_{i=1}^n X_{i}\right) = \sum_{i=1}^{n} var(X_{i}) + 2\sum_{i,j:i<j}^{n} cov(X_{i},X_{j})$ (Figure 1C).
+Even when the expression variation (standard deviation or coefficient of variation) and abundance is identical between these two scenarios, the total variance of a pathway or complex can be significantly different due to the variance sum law: $var\left(\sum_{i=1}^n X_{i}\right) = \sum_{i=1}^{n} var(X_{i}) + 2\sum_{i,j:i<j}^{n} cov(X_{i},X_{j})$ (Figure 1C).
 Our previous work has shown that this effect is important in single cell pathway activation dynamics `r citep("doi.org/10.15252/msb.20177997")`.
 
-In order to study the relationship between protein expression variance and coexpression (correlated expression) of cellular pathways and complexes at the proteome level, we utilized *Xenopus laevis* eggs as a single cell model.
-Activated *Xenopus laevis* eggs were collected at 5 time points across the first cell cycle (0, 20, 40, 60, and 80 minutes), with 5 eggs at each time point.
+In order to study the relationship between protein expression variance and co-expression (correlated expression) of cellular pathways and complexes at the proteome level, we utilized *Xenopus laevis* eggs as a single cell model.
+Activated *Xenopus laevis* eggs were collected at 5 time points across the first cell cycle (0, 20, 40, 60, and 80 minutes), with 5 eggs at each time point (Fig 1D).
 Using TMT multiplexing and mass spectrometry, we were able to determine the relative abundance of more than 1300 proteins.
 Expression of these proteins were largely invariant across the cell cycle, revealing that these highly expressed genes are likely not regulated by cell cycle processes.
-Additionally, a PCA analysis of these eggs showed no discernible clustering on cell cycle time (Fig S1).
+Additionally, a PCA analysis of these eggs showed no discernible clustering on cell cycle time (Fig S2).
 
 To determine the variance across these proteins, we calculated the variance and coefficient of variation (CV) using all time points collected.
-This showed a wide range of variation containing multiple distributions (Fig 1D), many of which are consistent with our previous study of variation using targeted mass spectrometry (Fig S2).
-In order to see if variation was a regulated process, we grouped the protein variation by gene ontology terms (GO terms) and plotted them in ranked order.
-We saw that in general, processes xxx were…
+This showed a wide range of variation containing multiple distributions (Fig 1E), many of which are consistent with our previous study of variation using targeted mass spectrometry (Fig S2).
 
 Since we were able to measure all of these proteins within single cells, we are able to calculate coordinated expression of protein pairs at single cell resolution.
-Using Pearson correlation coefficient, we could determine coordinated expression of nearly 2 million protein pairs (Fig 1C) The distribution of correlation coefficients fit a normal distribution, with the majority of protein pairs appearing to not show significant coregulation.
-However, there appear to be a significant number of protein pairs containing high correlation coefficients, and the heatmap showed a lot of clustering of these highly coregulated pairs (Fig 1D).
+Using Pearson correlation coefficient, we could determine coordinated expression of nearly 2 million protein pairs (Fig 1F-G) The distribution of correlation coefficients fit a normal distribution, with the majority of protein pairs appearing to not show significant co-regulation.
+However, there appear to be a significant number of protein pairs containing high correlation coefficients, and the heat-map showed a lot of clustering of these highly co-regulated pairs (Fig 1G).
 
@@
 fig_legend_1 <- c("A) Simulated expression of two proteins over time in a single cell with a mean of 100 and a standard deviation of 30 with expression correlations of 0 and 0.8.
-B) Pairwise plots of the two proteins in the single cells between uncorrelated and correlated scenerios.
-C) The total variance between the uncorrelated and correlated scenerios from A and B.
+B) Pairwise plots of the two proteins in the single cells between uncorrelated and correlated scenarios.
+C) The total variance between the uncorrelated and correlated scenarios from A and B.
 D) To measure protein expression variance and coordinated expression, 25 eggs at 5 time points during the first cell cycle were analyzed with multiplexed shotgun mass spectrometry.
@@
 F) Histogram of the pairwise correlation coefficients between all measured proteins.
-G) Heatmap of the pairwise correlation coefficients between all measured protein pairs.")
+G) Heat-map of the pairwise correlation coefficients between all measured protein pairs.")
 
@@
 ## Protein modules display different levels of regulation of noise and co-expression of constituents
+
+In order to see if variation was a regulated process, we grouped the protein variation by gene ontology terms (GO terms) and plotted them in ranked order. We saw that in general, processes xxx were…
 
@@
 fig_legend_2 <- c("A) Histogram of the normalized variation values across all of the GO terms and KEGG pathways.
-The bottom and top 10% are colored blue and red respectively.
-B) Pairwise plot between the mean correlation and mean coefficient of variaiton within each GO term and KEGG pathway with color coding from A.
+The bottom and top ~15% are colored blue and red respectively.
+B) Distributions of the mean correlation coefficient across all groups of proteins.
+Both the top and bottom 15% of normalized variation groups had peaks of high levels of mean correlation coefficients.
+C) Distributions of the mean coefficient of variation across all groups of proteins.
+The top and bottom ~15% of normalized variation groups were largely separated.
+D) Pairwise plot between the mean correlation and mean coefficient of variation within each GO term and KEGG pathway with color coding from A.
 Interestingly, both the top and bottom 10% of normalized variation have groups of proteins with high levels of correlation, despite high correlation adding to total variance.
@@
 ```
+
+## Total variance metric reveals certain modules enforce a trade-off between protein co-expression and noise
+
+The sum of variance for a group of proteins $X_{1},\dotsc,X_{n}$ with dependent relationships:
+
+$$\sigma_{{}_{total}}^2 = \sum_{i=1}^{n} \sigma_{X_{i}}^2 + 2\sum_{i,j:i<j}^{n} \rho(X_{i},X_{j})\sigma_{X_i}\sigma_{X_j}$$
+
+$$\sigma_{{}_{norm}} = \dfrac{\sqrt{\sigma_{{}_{total}}}}{n}$$
+
+When calculating the total variance of a pathway or general group of proteins, the total variance is the square root of the sums of the squares of the coefficient of variance.
+However, if any of the proteins are correlated in expression, an extra term must be added to account for this correlation.
+This means that if a group of proteins have a positive correlation, then the total variance will be higher than if there was no correlation.
+In order to study how variance and coordinated expression are related between single cells, we wanted to calculate a normalized total variance metric.
+Since we were not interested in the increase of variance due to the number of proteins in a module, we calculated the total variance using the coefficient of variation and correlation coefficients and normalize to the total number of proteins in the module.
+When total variance is plotted against the number of proteins, there is a clear linear relationship between the two.
+Once we normalize to the total number of proteins this relationship is lost and we are left with the effects of coefficient of variation and correlation.
+
+## Low total variance modules are enriched for metabolic pathways and heteromeric protein complexes
+
+## Enforcement of a trade-off between protein co-expression and noise increases efficiency of metabolic pathways and heteromeric protein complexes
 
@@
 
-fig_legend_3 <- c("A) Histogram of the normalized variation values across all of the GO terms and KEGG pathways.
-The bottom and top 10% are colored blue and red respectively.
-B) Pairwise plot between the mean correlation and mean coefficient of variaiton within each GO term and KEGG pathway with color coding from A.
-Interestingly, both the top and bottom 10% of normalized variation have groups of proteins with high levels of correlation, despite high correlation adding to total variance.
-In the lowest 10% of normalized variation groups there appears to be a negative correlation between mean correlation and variation.")
+fig_legend_3 <- c("A) A representative heat-map of one of the low normalized variation groups with high levels of correlated proteins (Large ribosomal subunit).
+B) A simple model showing the percent of potentially assembled subunits of an idealized heteromeric 10 subunit complex.
+Both the variation of expression between the subunits as well as the coordinated expression between them can effect the percent of maximum assembled complexes.
+C) Heat-map showing the total variance as a function of coefficient of variation and correlation from the idealized model in B).")
 ```
-
-## Total variance metric reveals certain modules enforce a tradeoff between protein co-expression and noise
-
-Variance Sum Law — Dependent Case
-
-The sum of variance for a group of proteins $X_{1},\dotsc,X_{n}$ with dependent relationships:
-
-$$var\left(\sum_{i=1}^n X_{i}\right) = \sum_{i=1}^{n} var(X_{i}) + 2\sum_{i,j:i<j}^{n} cov(X_{i},X_{j})$$
-
-$$\sigma_{{}_{total}} = \sum_{i=1}^{n} \sigma_{X_{i}}^2 + 2\sum_{i,j:i<j}^{n} \rho(X_{i},X_{j})\sigma_{X_i}\sigma_{X_j}$$
-
-$$\sigma_{{}_{total}} = \sqrt{var\left(\sum_{i=1}^n X_{i}\right)}$$
-
-$$\sigma_{{}_{norm}} = \dfrac{\sigma_{{}_{total}}}{n}$$
-
-When calculating the total variance of a pathway or general group of proteins, the total variance is the square root of the sums of the squares of the coefficient of variance.
-However, if any of the proteins are correlated in expression, an extra term must be added to account for this correlation.
-This means that if a group of proteins have a positive correlation, then the total variance will be higher than if there was no correlation.
-In order to study how variance and coordinated expression are related between single cells, we wanted to calculate a normalized total variance metric.
-Since we were not interested in the increase of variance due to the number of proteins in a module, we calculated the total variance using the coefficient of variation and correlation coefficients and normalize to the total number of proteins in the module.
-When total variance is plotted against the number of proteins, there is a clear linear relationship between the two.
-Once we normalize to the total number of proteins this relationship is lost and we are left with the effects of coefficient of variation and correlation.
-
-## Low total variance modules are enriched for metabolic pathways and heteromeric protein complexes
-
-## Enforcement of a tradeoff between protein co-expression and noise increases efficiency of metabolic pathways and heteromeric protein complexes
 
@@
 
-Xenopus egg extracts were prepared based on modifications of a previous protocol (Tsai et al, 2014).
+Xenopus egg extracts were prepared based on modifications of a previous protocol `r citep("10.1371/journal.pbio.1001788")`.
 All of the animal protocols used in this manuscript were approved by the Stanford University Administrative Panel on Laboratory Animal Care.
@@
 To activate the cell cycle, eggs were placed in a solution of 0.5 $\mu$g/ml of calcium ionophore A23187 (Sigma) and 1X MMR buffer for 3 min, after which they were washed four times with 1× MMR buffer.
-Single eggs were collected at their respective timepoints and placed into 600uL tubes and snap frozen in liquid nitrogen before being stored at 80°C.
+Single eggs were collected at their respective time-points and placed into 600uL tubes and snap frozen in liquid nitrogen before being stored at 80°C.
 
@@
 Single eggs were lysed mechanically by pipetting the egg in 100$\mu$L of lysis buffer (100 mM NaCl, 25 mM Tris pH 8.2, Complete EDTA- free protease inhibitor cocktail (Sigma).
-The lysate was then placed in a 400 uLnatural polyethylene microcentrifuge tube (E&K
+The lysate was then placed in a 400 uLnatural polyethylene micro-centrifuge tube (E&K
 Scientific #485050) and spun at 15,000 g in a right angle centrifuge (Beckman Microfuge E) at 4°C for 5 min.
@@
 sup_fig_legend_1 <- c("Pairwise plots of the mean correlation coefficient and mean coefficient of variation of protien groups with linear fit.
-The low variance protein groups have a significant negative correlation between these two variables, showing that for protien modules that require low variance and high coexpression, showing that these two parameters...
+The low variance protein groups have a significant negative correlation between these two variables, showing that for protein modules that require low variance and high co-expression, showing that these two parameters...
 R (high, low, med) = (0.057, -0.485, 0.071)")

### 2020-03-11 Manuscript/Manuscript.Rmd
.

@@
 
-```{r figure_1, echo=FALSE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}
+```{r figure_1, echo=FALSE, fig.height=4.5, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}
 allGroups <- readRDS("allGroups.RDS")
@@
 
-top / bottom + plot_layout(nrow = 2) +
+top / bottom + plot_layout(nrow = 2,
+                           heights = c(1,0.8)) +
   plot_annotation(tag_levels = "A")
@@
   scale_fill_manual(values = c("#d53e4f","#3288bd","grey")) +
-  xlab("Normalized Variation") + ylab("Number")
+  xlab("Normalized Variation") + ylab("Number") +
+  theme(legend.position = "none")
 
@@
   theme_minimal() +
-  theme(legend.position = "none") +
+  theme(legend.position = "bottom") +
   scale_color_manual(values = c("#d53e4f","#3288bd","grey")) +
@@
   theme_minimal() +
-  scale_fill_manual(values = c("#d53e4f","#3288bd","grey"))
+  scale_fill_manual(values = c("#d53e4f","#3288bd","grey")) +
+  theme(legend.position = "none")
 
@@
 
-p6 <- p2 + plot_spacer() + plot_layout(widths = c(1,0.5))
+# p6 <- p2 + plot_spacer() + plot_layout(widths = c(1,0.5))
+#   
+# (p1 + p3 + p4) / p6 + plot_layout(heights = c(1,5)) +
+#   plot_annotation(tag_levels = "A") &
+#   theme(legend.position = "none")
 
-(p1 + p3 + p4) / p6 + plot_layout(heights = c(1,5)) +
-  plot_annotation(tag_levels = "A") &
-  theme(legend.position = "none")
+left <- (p1 / p3 / p4) + plot_layout(ncol = 1)
+
+wrap_plots(left, p2) + plot_layout(widths = c(1,3)) +
+  plot_annotation(tag_levels = "A")
 

### 2020-03-11 Manuscript/Manuscript.Rmd
.

@@
 
-## Protein modules display different levels of regulation of noise and co-expression of constituents
+## Total variance metric reveals certain modules enforce a trade-off between protein co-expression and noise
 
-In order to see if variation was a regulated process, we grouped the protein variation by gene ontology terms (GO terms) and plotted them in ranked order. We saw that in general, processes xxx were…
+The sum of variance for a group of proteins $X_{1},\dotsc,X_{n}$ with dependent relationships:
 
-To investigate whether coordinated expression goes beyond just protein pairs, we wanted to see if pathways, complexes, and other modules had proteins that with coordinated expression.
-To answer this question, we grouped proteins by GO terms and calculated the median correlation coefficient across all pairs of proteins within the term.
-When the ranked media correlation is plotted it’s clear that there are a significant number of GO terms with higher than expected levels of coordinate expression.
+$$\sigma_{{}_{total}}^2 = \sum_{i=1}^{n} \sigma_{X_{i}}^2 + 2\sum_{i,j:i<j}^{n} \rho(X_{i},X_{j})\sigma_{X_i}\sigma_{X_j}$$
+
+$$\sigma_{{}_{norm}} = \dfrac{\sqrt{\sigma_{{}_{total}}}}{n}$$
+
+When calculating the total variance of a pathway or general group of proteins, the total variance is the square root of the sums of the squares of the coefficient of variance.
+However, if any of the proteins are correlated in expression, an extra term must be added to account for this correlation.
+This means that if a group of proteins have a positive correlation, then the total variance will be higher than if there was no correlation.
+In order to study how variance and coordinated expression are related between single cells, we wanted to calculate a normalized total variance metric.
+Since we were not interested in the increase of variance due to the number of proteins in a module, we calculated the total variance using the coefficient of variation and correlation coefficients and normalize to the total number of proteins in the module.
+When total variance is plotted against the number of proteins, there is a clear linear relationship between the two.
+Once we normalize to the total number of proteins this relationship is lost and we are left with the effects of coefficient of variation and correlation.
 
@@
 p1 <- ggplot(allGroups, aes(x = normVar, fill = interval)) + 
-  geom_histogram(bins = 50) + theme_minimal() +
-  #scale_fill_viridis_d()
+  geom_histogram(data = filter(allGroups, interval == "mid"),bins = 50) + 
+  geom_histogram(data = filter(allGroups, interval == "low"),bins = 50) + 
+  geom_histogram(data = filter(allGroups, interval == "high"),bins = 50) + 
+  theme_minimal() +
   scale_fill_manual(values = c("#d53e4f","#3288bd","grey")) +
@@
 
-#p1 + p2 + p3 + plot_annotation(tag_levels = "A")
-
 fig_legend_2 <- c("A) Histogram of the normalized variation values across all of the GO terms and KEGG pathways.
@@
 
-## Total variance metric reveals certain modules enforce a trade-off between protein co-expression and noise
+## Protein modules display different levels of regulation of noise and co-expression of constituents
 
-The sum of variance for a group of proteins $X_{1},\dotsc,X_{n}$ with dependent relationships:
+In order to see if variation was a regulated process, we grouped the protein variation by gene ontology terms (GO terms) and plotted them in ranked order. We saw that in general, processes xxx were…
 
-$$\sigma_{{}_{total}}^2 = \sum_{i=1}^{n} \sigma_{X_{i}}^2 + 2\sum_{i,j:i<j}^{n} \rho(X_{i},X_{j})\sigma_{X_i}\sigma_{X_j}$$
+To investigate whether coordinated expression goes beyond just protein pairs, we wanted to see if pathways, complexes, and other modules had proteins that with coordinated expression.
+To answer this question, we grouped proteins by GO terms and calculated the median correlation coefficient across all pairs of proteins within the term.
+When the ranked media correlation is plotted it’s clear that there are a significant number of GO terms with higher than expected levels of coordinate expression.
 
-$$\sigma_{{}_{norm}} = \dfrac{\sqrt{\sigma_{{}_{total}}}}{n}$$
+```{r figure_3, echo=FALSE, fig.height=6, fig.width=7, message=FALSE, warning=FALSE, eval=TRUE, fig.cap=fig_legend_3}
+plots <- 1:4
+for(i in 1:length(plots)){
+  assign(paste0("p",plots[i]), cor_rug_plot(interesting_groups_low[i], "#3288bd"))
+}
 
-When calculating the total variance of a pathway or general group of proteins, the total variance is the square root of the sums of the squares of the coefficient of variance.
-However, if any of the proteins are correlated in expression, an extra term must be added to account for this correlation.
-This means that if a group of proteins have a positive correlation, then the total variance will be higher than if there was no correlation.
-In order to study how variance and coordinated expression are related between single cells, we wanted to calculate a normalized total variance metric.
-Since we were not interested in the increase of variance due to the number of proteins in a module, we calculated the total variance using the coefficient of variation and correlation coefficients and normalize to the total number of proteins in the module.
-When total variance is plotted against the number of proteins, there is a clear linear relationship between the two.
-Once we normalize to the total number of proteins this relationship is lost and we are left with the effects of coefficient of variation and correlation.
+plots <- 5:8
+for(i in 1:length(plots)){
+  assign(paste0("p",plots[i]), cor_rug_plot(interesting_groups_high[i], "#d53e4f"))
+}
+
+left <- p1 + p2 + p3 + p4 + plot_layout(nrow = 1)
+
+right <- p5 + p6 + p7 + p8 + plot_layout(nrow = 1)
+
+wrap_plots(left, right) + plot_annotation(tag_levels = "A") + plot_layout(nrow = 2)
+
+fig_legend_3 <- c("Density plots of correlation coefficients select protein groups color coded by low (blue) and high (red) normalized protein variance, with the distribution of all measured correlation coefficients colored in grey (rug plot to show the measured correlations coefficents shown on plot).
+Below each density plot is a box plot.")
+```
 
@@
 
-```{r figure_3, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, eval=TRUE, fig.cap=fig_legend_3}
+```{r figure_4, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, eval=TRUE, fig.cap=fig_legend_3}
 library(circlize)

### 2020-03-11 functions.R

@@ -38,9 +38,11 @@ cor_rug_plot <- function(x, color = '#e08214'){
   
   temp <- allGroups %>% select(name, cors) %>%
     unnest(cols = cors) %>% mutate(name = "all") %>%
+    filter(!duplicated(cors)) %>%
     rbind(temp,.)
   
-  p1 <- ggplot(temp, aes(x = cors, fill = name)) + geom_density(alpha = 0.75, color = "transparent") +
+  p1 <- ggplot(temp, aes(x = cors, fill = name)) + 
+    geom_density(alpha = 0.75, color = "transparent") +
     scale_fill_manual(values = c('#bababa',color), name = "") +
     geom_rug(data = filter(temp, name != "all"), color = color) +
     xlab("") + ylab("Density") +

### 2020-03-11 scratch.R

@@ -49,3 +49,8 @@ pca = prcomp(as.matrix(glyc[,2:ncol(glyc)]))
 pcaDF = tibble(PC1 = pca$rotation[,1], PC2 = pca$rotation[,2])
 ggplot(pcaDF, aes(x = PC1, y = PC2)) + geom_point()
 
+int <- matrix(data = c(0,1,0,
+                       0,0,1,
+                       1,0,0),
+              nrow = 3)
+mat <- cor[1:3,1:3]

### 2020-03-11 Manuscript/Manuscript.Rmd
.

@@
 
-```{r figure_3, echo=FALSE, fig.height=6, fig.width=7, message=FALSE, warning=FALSE, eval=TRUE, fig.cap=fig_legend_3}
+```{r figure_3, echo=FALSE, fig.height=4, fig.width=7, message=FALSE, warning=FALSE, eval=TRUE, fig.cap=fig_legend_3}
 plots <- 1:4
@@
 
-wrap_plots(left, right) + plot_annotation(tag_levels = "A") + plot_layout(nrow = 2)
+wrap_plots(left, right) + 
+  plot_annotation(tag_levels = "A") + 
+  plot_layout(nrow = 2)
 

### 2020-03-11 functions.R

@@ -46,7 +46,7 @@ cor_rug_plot <- function(x, color = '#e08214'){
     scale_fill_manual(values = c('#bababa',color), name = "") +
     geom_rug(data = filter(temp, name != "all"), color = color) +
     xlab("") + ylab("Density") +
-    theme_bw() +
+    theme_minimal() +
     # theme(legend.position = c(.2,.9),
     #       legend.background = element_rect(fill="transparent")) +
     theme(legend.position = "none",
@@ -58,7 +58,7 @@ cor_rug_plot <- function(x, color = '#e08214'){
     geom_boxplot(notch = TRUE, alpha = 0.75) +
     scale_fill_manual(values = c('#bababa',color)) +
     ylab("Correlation Coefficients") +
-    theme_bw() +
+    theme_minimal() +
     theme(legend.position = "none",
           text = element_text(size = 6),
           axis.title.y=element_blank(),

### 2020-03-11 Manuscript/Manuscript.Rmd
.

@@
 
+\newpage
+
 ```{r figure_1, echo=FALSE, fig.height=4.5, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}
@@
 
+\newpage
+
 ## Total variance metric reveals certain modules enforce a trade-off between protein co-expression and noise
@@
 Once we normalize to the total number of proteins this relationship is lost and we are left with the effects of coefficient of variation and correlation.
+
+\newpage
 
@@
 
+\newpage
+
 ## Protein modules display different levels of regulation of noise and co-expression of constituents
@@
 When the ranked media correlation is plotted it’s clear that there are a significant number of GO terms with higher than expected levels of coordinate expression.
+
+\newpage
 
@@
 
+\newpage
+
 ## Low total variance modules are enriched for metabolic pathways and heteromeric protein complexes
@@
 
-```{r figure_4, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, eval=TRUE, fig.cap=fig_legend_3}
+\newpage
+
+```{r figure_4, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, eval=TRUE, fig.cap=fig_legend_4}
 library(circlize)
@@
 
+p4 <- cowplot::ggdraw() + cowplot::draw_image(magick::image_read_svg("../images/complex_cartoon.svg"))
+
 bottom <- p2 + p3 + plot_layout(widths = c(1,2))
-(p1 + plot_spacer()) / bottom + 
+(p1 + p4) / bottom + 
   plot_annotation(tag_levels = "A")
 
-fig_legend_3 <- c("A) A representative heat-map of one of the low normalized variation groups with high levels of correlated proteins (Large ribosomal subunit).
+fig_legend_4 <- c("A) A representative heat-map of one of the low normalized variation groups with high levels of correlated proteins (Large ribosomal subunit).
 B) A simple model showing the percent of potentially assembled subunits of an idealized heteromeric 10 subunit complex.
@@
 ```
+
+\newpage
 

### 2020-03-11 Untitled.Rmd
.

@@
 cor_rug_plot <- function(x){
-  temp <- filter(allGO, GOterms %in% x) %>%
-    dplyr::select(GO, cors) %>% unnest(cols = cors)
+  temp <- filter(allGroups, accession %in% x) %>%
+    dplyr::select(name, cors) %>% unnest(cols = cors)
 
-  temp <- allGO %>% select(GO, cors) %>%
-    unnest(cols = cors) %>% mutate(GO = "all") %>%
+  temp <- allGroups %>% select(name, cors) %>%
+    unnest(cols = cors) %>% mutate(name = "all") %>%
     rbind(temp,.)
 
-  ggplot(temp, aes(x = cors, fill = GO, color = GO)) +
+  ggplot(temp, aes(x = cors, fill = name, color = name)) +
     geom_density(alpha = 0.25) + 
-    geom_rug(data = filter(temp, GO != "all")) + 
+    geom_rug(data = filter(temp, name != "all")) + 
     theme_bw()
 }
-
+cor_rug_plot("GO:0006739")
 ```

### 2020-03-11 functions.R

@@ -50,17 +50,17 @@ cor_rug_plot <- function(x, color = '#e08214'){
     # theme(legend.position = c(.2,.9),
     #       legend.background = element_rect(fill="transparent")) +
     theme(legend.position = "none",
-          text = element_text(size = 6)) +
+          text = element_text(size = 4)) +
     xlim(-1,1) +
-    ggtitle(filter(temp, name != "all") %>% pull(name) %>% unique())
+    ggtitle(str_wrap(filter(temp, name != "all") %>% pull(name) %>% unique(), width = 20))
   
   p2 <- ggplot(temp, aes(x = name, y = cors, fill = name)) + 
-    geom_boxplot(notch = TRUE, alpha = 0.75) +
+    geom_boxplot(notch = TRUE, alpha = 0.75, outlier.size = 0.25) +
     scale_fill_manual(values = c('#bababa',color)) +
     ylab("Correlation Coefficients") +
     theme_minimal() +
     theme(legend.position = "none",
-          text = element_text(size = 6),
+          text = element_text(size = 4),
           axis.title.y=element_blank(),
           axis.text.y=element_blank(),
           axis.ticks.y=element_blank()) +

### 2020-03-12 Manuscript/Manuscript.Rmd
.

@@
 output:
-
   pdf_document:
@@
 
+  word_document: null
   html_document: null
 
-  word_document: null
 documentclass: article
@@
 
-## Total variance metric reveals certain modules enforce a trade-off between protein co-expression and noise
+## Variance sum law reveals modules that utilize a trade-off between protein co-expression and noise
 
-The sum of variance for a group of proteins $X_{1},\dotsc,X_{n}$ with dependent relationships:
-
+Multicomponent modules such as heteromeric protein complexes and metabolic and signaling pathways have a certain tolerance for noise.
+Two components of the noise of these systems are the expression variance of the individual proteins, as well as the correlation or covariance of expression between the proteins.
+This means that if a group of proteins have a positive correlation, then the total variance will be higher than if there was no correlation.
+This single cell proteomics dataset gave us an excellent opportunity to see how variance and co-expression of proteins are related in low variance and high variance modules.
+The total variance of a module can be calculated using the variance sum law, where for a group of proteins $X_{1},\dotsc,X_{n}$, the total variance of the module is:
 $$\sigma_{{}_{total}}^2 = \sum_{i=1}^{n} \sigma_{X_{i}}^2 + 2\sum_{i,j:i<j}^{n} \rho(X_{i},X_{j})\sigma_{X_i}\sigma_{X_j}$$
 
-$$\sigma_{{}_{norm}} = \dfrac{\sqrt{\sigma_{{}_{total}}}}{n}$$
+To categorize the measured protein expression data into modules, we utilized GO term and KEGG pathway annotations.
+Since pathways and complexes have varying numbers of proteins, and the number of proteins in a module is highly influential on the total variance sum, we normalized the total variance sum by the number of components (Fig.
+xA and xB).
+With this normalized total variance metric, we could separate modules into high and low variance categories (Fig 2A).
+Since co-expression of proteins (positive correlation between protein expression) increases the total variance, we expected to see the high variance modules to be enriched for high levels of both protein expression variance and co-expression (Fig 1B and 1C, red).
+However, we were suprised to see that the low variance modules were also enriched for high levels of co-expression of proteins, with some of the mean correlation coefficients of modules as high as those in the high variance category (Fig 1B, orange shaded region).
 
-When calculating the total variance of a pathway or general group of proteins, the total variance is the square root of the sums of the squares of the coefficient of variance.
-However, if any of the proteins are correlated in expression, an extra term must be added to account for this correlation.
-This means that if a group of proteins have a positive correlation, then the total variance will be higher than if there was no correlation.
-In order to study how variance and coordinated expression are related between single cells, we wanted to calculate a normalized total variance metric.
-Since we were not interested in the increase of variance due to the number of proteins in a module, we calculated the total variance using the coefficient of variation and correlation coefficients and normalize to the total number of proteins in the module.
-When total variance is plotted against the number of proteins, there is a clear linear relationship between the two.
-Once we normalize to the total number of proteins this relationship is lost and we are left with the effects of coefficient of variation and correlation.
+We wondered how these protein modules dealted with the increased variance load caused by the high numbers of correlated proteins, so we looked at the pairwise relationship between the mean variance of the modules and the mean correlations (Fig S1).
+The low variance modules uniquely showed a significant negative correlation (R = 0.49) between these two metrics, meaning that for modules that enforce a low level of total variance, co-expression of proteins within that module must also have low expression variance.
 
@@
 
-In order to see if variation was a regulated process, we grouped the protein variation by gene ontology terms (GO terms) and plotted them in ranked order. We saw that in general, processes xxx were…
-
-To investigate whether coordinated expression goes beyond just protein pairs, we wanted to see if pathways, complexes, and other modules had proteins that with coordinated expression.
-To answer this question, we grouped proteins by GO terms and calculated the median correlation coefficient across all pairs of proteins within the term.
-When the ranked media correlation is plotted it’s clear that there are a significant number of GO terms with higher than expected levels of coordinate expression.
+We saw that in general, processes xxx were… (Fig. 2D).
+(Fig. 3A-D) (Fig. 3E-H)
 

### 2020-03-12 Manuscript/Manuscript.Rmd
.

@@
 output:
+  # word_document:
+  # html_document:  
   pdf_document:
@@
       in_header: header.tex
-
-  word_document: null
-  html_document: null
 
@@
 
-The coordinated expression of proteins is vital for many aspects of cellular function, from maintaining a dynamic steady state to differentiation of cells into specialized types in different tissues.
-These requirements can range from precise stoichiometric coordination for multisubunit complexes and metabolic pathways, to high levels of noise that can result in heterogeneous responses of identical cell populations to identical stimuli `r citep("10.1073/pnas.1615660114")`.
-The ability to regulate coordinated expression has been demonstrated to occur via transcription (transcription factors, chromatin regulation), translation (specialized ribosomes, mRNA structure/modifications), and degradation (E3 ligases).
-These processes, as well as the molecules they target, are all subject to stochastic expression and interactions, which has been shown to lead to differences in cell behavior and decisions in otherwise identical cells.
-Variability and coordinated expression (correlation) of proteins leads to increased population level control in binary decisions, and a decreased ability to execute analog signaling.
+The coordinated expression of proteins is vital for cellular function, from maintaining a dynamic steady state to differentiation of cells into specialized types and tissues.
+This coordinated regulation of protein abundance is subject to stochastic expression levels, and modules (pathways, complexes, etc.) have varying tolerances of noise for proper function ranging from precise stoichiometric coordination, to high levels of noise that can result in heterogeneous responses of homogeneous cell populations to identical stimuli `r citep("10.1073/pnas.1615660114")`.
+The ability to regulate coordinated expression has been demonstrated to occur via transcription (transcription factors, chromatin regulation), translation (specialized ribosomes, mRNA structure/modifications), and degradation (E3 ligases), all of which are subject to stochastic expression and interactions.
+This noise has been shown be utilized by cells for bet hedging strategies as well as tissue size maintenance (more concrete examples to come).
+Additionally, the combination of variability and coordinated expression (correlation) of proteins can lead to increased population level control in binary decisions, and a decreased ability to execute analog signaling `r citep(c("doi.org/10.15252/msb.20177997", "10.1073/pnas.1615660114"))`.
 
-There have been numerous targeted studies of coordinated protein expression and variation that have shown the important impacts of these parameters on cellular function.
+There have been numerous studies, both targeted and unbiased, of coordinated protein expression or expression variation independently that have shown the important impacts of these parameters on cellular function.
 However, both of these parameters (variation and correlation) act together to determine the total variance of a system.
+To systematically assess these properties of single cells at a large scale, we conducted shotgun proteomics analyses on *Xenopus laevis* eggs during their first cell cycle.
+We have found this model to be ideal at this nascent stage of the single cell proteomics field because of their large size and tractability.
+An advantage of this approach is that we can get around much of the signal to noise issues that accompany studying single cells caused by low sample protein levels.
+Additionally, at this stage of development transcription is restricted, allowing for insights into non-transcriptional control of protein expression.
+By utilizing isobaric tagging and shotgun mass spectrometry we were able to measure >1000 proteins across 25 single cells in 5 mass spectrometry runs.
 
-there is a gap in the literature of direct measurement of the relationship between variation and coordinated protein expression between single cells.
-To systematically assess these properties of single cells, we carried out a proteomics experiment on exceeding large single cells, Xenopus laevis eggs.
-An advantage of this approach is that we can get around much of the signal to noise issues that accompany studying single cells due to low sample amounts.
-Additionally, at this stage of development transcription is restricted so we are able to gain insights into non-transcriptional control of protein expression.
-We also utilized isobaric tagging in order to measure 25 individual eggs in 5 mass spectrometry runs.
-In this study, we were able to measure the relative abundance >1000 proteins across single cells in order to better understand the properties of stochastic and coordinated expression of proteins.
-
-With this data set we have been able to, for the first time, measured the relationship between protein expression variance and coordinated protein expression on a proteome scale in single cells.
-We have observed that certain classes of proteins, including protein complexes and metabolic pathways, are expressed in such a way that increased coordinated expression is balanced by decreased variation.
-By doing this, cells are able to decrease the total variance of a given complex or pathway, an elegant balancing act that allows for finer control of metabolic throughput and controlling the number of potentially formed complexes though stoichiometric control.
-Though this kind of coordinated expression leads to an increase in variance in a population of cells, it can reduce variation within a cell.
+With this data set we have been able to, for the first time, gain insights into the relationship between protein expression variance, coordinated protein expression, and total variance within modules at a proteomic scale in single cells.
+We identified classes of proteins that include heteromeric complexes and metabolic pathways that are expressed in such a way that the increase in total variance of the system caused by high coordinated expression of proteins is offset by decreased variation, allowing for modules that rely on stoichiometric expression and low level of variance to maintain a low level of total system variance.
+By counter balancing coordinated expression (correlation) of proteins in a module with low expression variance, cells are able to decrease the total variance of a given complex or pathway.
+This elegant balancing act may allow for finer control of metabolic throughput and controlling the number of potentially formed complexes though stoichiometric control.
 
@@
 
-The specific requirements for coordinated protein expression of pathways or complexes can vary, with some requiring strict stoichiometric regulation with others having much more relaxed requirements.
+The specific requirements for coordinated protein expression of pathways or complexes can vary, with some requiring strict stoichiometric regulation and others having much more relaxed requirements.
 Strict stoichiometry scenarios require coordinated expression of proteins (Figure 1A, top), and others can have uncoordinated expression (Figure 1A bottom).
@@
 Activated *Xenopus laevis* eggs were collected at 5 time points across the first cell cycle (0, 20, 40, 60, and 80 minutes), with 5 eggs at each time point (Fig 1D).
-Using TMT multiplexing and mass spectrometry, we were able to determine the relative abundance of more than 1300 proteins.
-Expression of these proteins were largely invariant across the cell cycle, revealing that these highly expressed genes are likely not regulated by cell cycle processes.
+Using TMT multiplexing and mass spectrometry, we were able to determine the relative abundance of more than 1300 proteins in single cells.
+Expression of these proteins across the time course of the cell cycle showed no dynamic patterns, revealing that these highly expressed proteins are likely not regulated by cell cycle processes.
 Additionally, a PCA analysis of these eggs showed no discernible clustering on cell cycle time (Fig S2).
 
-To determine the variance across these proteins, we calculated the variance and coefficient of variation (CV) using all time points collected.
+To determine the expression variability for the measured proteins between single cells, we calculated the variance and coefficient of variation (CV) across all samples.
 This showed a wide range of variation containing multiple distributions (Fig 1E), many of which are consistent with our previous study of variation using targeted mass spectrometry (Fig S2).
+An added benefit to measuring these proteins in parallel is that we are able to calculate coordinated expression of protein pairs at single cell resolution.
+Using the Pearson correlation coefficient, we were able to determine the coordinated expression of nearly 2 million protein pairs (Fig 1F-G).
+The distribution of correlation coefficients fit a normal distribution centered around 0, with the majority of protein pairs appearing to not show significant co-regulation.
+However, there appear to be a significant number of protein pairs containing high correlation coefficients, and a clustered heat map shows that many highly co-regulated pairs cluster together (Fig 1G).
 
-Since we were able to measure all of these proteins within single cells, we are able to calculate coordinated expression of protein pairs at single cell resolution.
-Using Pearson correlation coefficient, we could determine coordinated expression of nearly 2 million protein pairs (Fig 1F-G) The distribution of correlation coefficients fit a normal distribution, with the majority of protein pairs appearing to not show significant co-regulation.
-However, there appear to be a significant number of protein pairs containing high correlation coefficients, and the heat-map showed a lot of clustering of these highly co-regulated pairs (Fig 1G).
+## Variance sum law reveals modules that utilize a trade-off between protein co-expression and noise
+
+Multicomponent modules such as heteromeric protein complexes and metabolic and signaling pathways have a certain tolerance for noise.
+Two components of the noise of these systems are the expression variance of the individual proteins, as well as the correlation or co-variance of expression between the proteins.
+This means that if a group of proteins have a positive correlation, then the total variance will be higher than if there was no correlation.
+This single cell proteomics data set gave us an excellent opportunity to see how variance and co-expression of proteins are related in low variance and high variance modules.
+The total variance of a module can be calculated using the variance sum law, where for a group of proteins $X_{1},\dotsc,X_{n}$, the total variance of the group is:
+$$\sigma_{{}_{total}}^2 = \sum_{i=1}^{n} \sigma_{X_{i}}^2 + 2\sum_{i,j:i<j}^{n} \rho(X_{i},X_{j})\sigma_{X_i}\sigma_{X_j}$$
+
+To categorize the measured protein expression data into modules, we utilized GO term and KEGG pathway annotations.
+Since pathways and complexes have varying numbers of proteins, and the number of proteins in a module is highly influential on the total variance sum, we normalized the total variance sum by the number of components (Fig.
+xA and xB) to eliminate this confounding dependency.
+With the normalized total variance metric, we could separate modules into high and low variance categories (Fig 2A).
+Since co-expression of proteins (positive correlation between protein expression) increases the total variance, we expected to see the high variance modules to be enriched for high levels of both protein expression variance and co-expression (Fig 1B and 1C, red).
+However, we were surprised to see that the low variance modules were also enriched for high levels of co-expression of proteins, with some of the mean correlation coefficients of modules as high as those in the high variance category (Fig 1B, orange shaded region).
+
+We wondered how these protein modules dealt with the increased variance load caused by the high numbers of correlated proteins, so we looked at the pairwise relationship between the mean variance of the modules and the mean correlation coefficients (Fig 2D and S1).
+The low variance modules uniquely showed a significant negative correlation (R = 0.49) between these two metrics, showing that for modules that enforce a low level of total variance, co-expression of proteins within that module is offset by lower expression variance.
+
+## Enforcement of a trade-off between protein co-expression and noise increases efficiency of heteromeric protein complexes and metabolic pathways
+
+Interestingly, we saw the low total variance and high coordinated expression sub population contained many metabolic pathways as well as multisubunit complexes, while the high total variance and high coordinated expression sub-population contained many signalling pathways (Fig.
+2D, Fig.
+3E-H).
+Previous studies have shown that protein complex components are produced stoichiometricly `r citep("doi.org/10.1016/j.cels.2018.11.003")` across eukaryotic species and that co-expression of pathway specific enzymes are tightly regulated post-transcriptionally across species `r citep("doi.org/10.1016/j.cell.2018.03.007")`.
+This led us to wondered what the implications were for the synergy of maintaing low total variance with high levels of co-expression of the members of protein complexes and metabolic pathways.
+
+To test why complexes observed in our data set and others, (i.e.
+core members of the ribosome (Fig 4A)), coordinate co-expression and expression variation, we imagined an idealized 10 member heteromeric complex (Fig 4B).
+Coordinated expression could allow for tight control of the total number of assembled complexes.
+However, this on it's own is not sufficient since high levels of individual protein variance would result in an increased number of partially assembled complexes.
+In this idealized complex, we titrated the correlation coefficients between all 10 members along with the individual expression variation and saw that increased correlations and decreasing variance had a synergistic effect in allowing for the control of the total number of fully assembled complexes (Fig 4C.).
+Additionally, there is a large regime with high correlations and low variation that allow for low total variance of the system.
+
+We also observed a number of metabolic pathways that had similar variation and correlation properties as the protein complexes, such as the pentose phosphate pathway and glycolysis.
+Both of these pathways have a large number of reversible enzymes.
+For a reversible enzymatic pathway, the concentrations of the enzymes dictate the rate of the reaction, and the concentrations of the reaction products relative to the substrates can dictate the direction of the reaction equilibrium.
+Therefore, strict balance of the concentration of enzymes using low levels of variance and strict control of the relative abundances of the enzymes is a potentially important step in regulating metabolic pathway output.
+To test this hypothesis, we constructed an idealized metabolic model (or a physiological model??) and tested the ability of the pathway to maximize output (Fig 5x).
+(haven't done this, would like feedback / help).
+
+## Discussion
 
 \newpage
+
+# Figures
 
@@
 
-## Variance sum law reveals modules that utilize a trade-off between protein co-expression and noise
-
-Multicomponent modules such as heteromeric protein complexes and metabolic and signaling pathways have a certain tolerance for noise.
-Two components of the noise of these systems are the expression variance of the individual proteins, as well as the correlation or covariance of expression between the proteins.
-This means that if a group of proteins have a positive correlation, then the total variance will be higher than if there was no correlation.
-This single cell proteomics dataset gave us an excellent opportunity to see how variance and co-expression of proteins are related in low variance and high variance modules.
-The total variance of a module can be calculated using the variance sum law, where for a group of proteins $X_{1},\dotsc,X_{n}$, the total variance of the module is:
-$$\sigma_{{}_{total}}^2 = \sum_{i=1}^{n} \sigma_{X_{i}}^2 + 2\sum_{i,j:i<j}^{n} \rho(X_{i},X_{j})\sigma_{X_i}\sigma_{X_j}$$
-
-To categorize the measured protein expression data into modules, we utilized GO term and KEGG pathway annotations.
-Since pathways and complexes have varying numbers of proteins, and the number of proteins in a module is highly influential on the total variance sum, we normalized the total variance sum by the number of components (Fig.
-xA and xB).
-With this normalized total variance metric, we could separate modules into high and low variance categories (Fig 2A).
-Since co-expression of proteins (positive correlation between protein expression) increases the total variance, we expected to see the high variance modules to be enriched for high levels of both protein expression variance and co-expression (Fig 1B and 1C, red).
-However, we were suprised to see that the low variance modules were also enriched for high levels of co-expression of proteins, with some of the mean correlation coefficients of modules as high as those in the high variance category (Fig 1B, orange shaded region).
-
-We wondered how these protein modules dealted with the increased variance load caused by the high numbers of correlated proteins, so we looked at the pairwise relationship between the mean variance of the modules and the mean correlations (Fig S1).
-The low variance modules uniquely showed a significant negative correlation (R = 0.49) between these two metrics, meaning that for modules that enforce a low level of total variance, co-expression of proteins within that module must also have low expression variance.
-
-\newpage
-
 ```{r figure_2, echo=FALSE, fig.height=7, fig.width=6, message=FALSE, warning=FALSE, eval=TRUE, fig.cap=fig_legend_2}
@@
 
-## Protein modules display different levels of regulation of noise and co-expression of constituents
-
-We saw that in general, processes xxx were… (Fig. 2D).
-(Fig. 3A-D) (Fig. 3E-H)
-
-\newpage
-
 ```{r figure_3, echo=FALSE, fig.height=4, fig.width=7, message=FALSE, warning=FALSE, eval=TRUE, fig.cap=fig_legend_3}
@@
 
-fig_legend_3 <- c("Density plots of correlation coefficients select protein groups color coded by low (blue) and high (red) normalized protein variance, with the distribution of all measured correlation coefficients colored in grey (rug plot to show the measured correlations coefficents shown on plot).
+fig_legend_3 <- c("Density plots of correlation coefficients select protein groups color coded by low (blue) and high (red) normalized protein variance, with the distribution of all measured correlation coefficients colored in grey (rug plot to show the measured correlations coefficients shown on plot).
 Below each density plot is a box plot.")
 ```
-
-\newpage
-
-## Low total variance modules are enriched for metabolic pathways and heteromeric protein complexes
-
-## Enforcement of a trade-off between protein co-expression and noise increases efficiency of metabolic pathways and heteromeric protein complexes
 
@@
 
+# Supplementary material {-}
+
+\beginsupplement
+```{r sup_fig_1, echo=FALSE, fig.height=2, fig.width=6, message=FALSE, warning=FALSE, fig.cap=sup_fig_legend_1}
+allGroups %>% filter(n > 4) %>%
+  ggplot(., aes(x = medCors, y = medCV, color = interval)) + 
+  geom_point() + 
+  scale_color_manual(values = c("#d53e4f","#3288bd","grey")) +
+  stat_summary(fun.data=mean_cl_normal) + 
+  geom_smooth(method='lm', formula= y~x, color = "black") +
+  facet_wrap(~interval, scales = "free", nrow = 1) +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+sup_fig_legend_1 <- c("Pairwise plots of the mean correlation coefficient and mean coefficient of variation of protein groups with linear fit.
+The low variance protein groups have a significant negative correlation between these two variables, showing that for protein modules that require low variance and high co-expression, showing that these two parameters...
+R (high, low, med) = (0.057, -0.485, 0.071)")
+```
+
+\newpage
+
 # Methods
@@
 
-\newpage
-# Supplementary material {-}
-
-\beginsupplement
-```{r sup_fig_1, echo=FALSE, fig.height=2, fig.width=6, message=FALSE, warning=FALSE, fig.cap=sup_fig_legend_1}
-allGroups %>% filter(n > 4) %>%
-  ggplot(., aes(x = medCors, y = medCV, color = interval)) + 
-  geom_point() + 
-  scale_color_manual(values = c("#d53e4f","#3288bd","grey")) +
-  stat_summary(fun.data=mean_cl_normal) + 
-  geom_smooth(method='lm', formula= y~x, color = "black") +
-  facet_wrap(~interval, scales = "free", nrow = 1) +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-sup_fig_legend_1 <- c("Pairwise plots of the mean correlation coefficient and mean coefficient of variation of protien groups with linear fit.
-The low variance protein groups have a significant negative correlation between these two variables, showing that for protein modules that require low variance and high co-expression, showing that these two parameters...
-R (high, low, med) = (0.057, -0.485, 0.071)")
-```
-
 \eleft

### 2020-03-14 Manuscript/Manuscript.Rmd
added marginal histograms to fig 2

@@
   mutate(abundance = abundance / mean(abundance, na.rm = T)) %>% 
+  filter(n() > 15) %>% # comment out if using KNN
   group_by(cell) %>% 
@@
   dplyr::select(cell, Uniprot, abundance) %>% 
-  spread(cell, abundance) %>%
-  kNN(data = ., k = 1, impNA = TRUE) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)
+  spread(cell, abundance)
+  # spread(cell, abundance) %>%
+  # kNN(data = ., k = 1, impNA = TRUE) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)
 
@@
   filter(condition != 'ctrl') %>% group_by(Uniprot) %>%
-  summarise(var = var(abundance),
-            sd = sd(abundance),
-            mean = mean(abundance),
+  summarise(var = var(abundance, na.rm = TRUE),
+            sd = sd(abundance, na.rm = TRUE),
+            mean = mean(abundance, na.rm = TRUE),
             cv = sd / mean)
@@
 allGroups <- allGroups %>% mutate(interval = cut_number(normVar, n = 7)) %>%
-  mutate(interval = ifelse(interval == levels(interval)[1],
+  mutate(int_range = interval,
+         interval = ifelse(interval == levels(interval)[1],
                            "low",
@@
 
-saveRDS(allGroups,"allGroups.RDS")
+#saveRDS(allGroups,"allGroups.RDS")
 ```
@@
 ```{r figure_1, echo=FALSE, fig.height=4.5, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}
-allGroups <- readRDS("allGroups.RDS")
+#allGroups <- readRDS("allGroups.RDS")
 
@@
   scale_color_manual(values = c("#d53e4f","#3288bd","grey")) +
-  xlab("Mean Correlation") + ylab("Mean Coefficient of Variation")
+  xlab("Mean Correlation") + ylab("Mean Coefficient of Variation") +
+  xlim(-0.15,0.5) + 
+  ylim(0.08,0.32)
 
@@
   scale_fill_manual(values = c("#d53e4f","#3288bd","grey")) +
-  theme(legend.position = "none")
+  theme(legend.position = "none",
+        axis.title.x=element_blank(),
+        axis.text.x=element_blank(),
+        axis.ticks.x=element_blank()) +
+  xlim(-0.15,0.5)
 
@@
   scale_fill_manual(values = c("#d53e4f","#3288bd","grey")) +
-  theme(legend.position = "none")
+  theme(legend.position = "none",
+        axis.title.y=element_blank(),
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank()) +
+  xlim(0.08,0.32)
 
@@
 
-# p6 <- p2 + plot_spacer() + plot_layout(widths = c(1,0.5))
+# left <- (p1 / p3 / p4) + plot_layout(ncol = 1)
 # 
-# (p1 + p3 + p4) / p6 + plot_layout(heights = c(1,5)) +
-#   plot_annotation(tag_levels = "A") &
-#   theme(legend.position = "none")
+# wrap_plots(left, p2) + plot_layout(widths = c(1,3)) +
+#   plot_annotation(tag_levels = "A")
 
-left <- (p1 / p3 / p4) + plot_layout(ncol = 1)
-
-wrap_plots(left, p2) + plot_layout(widths = c(1,3)) +
-  plot_annotation(tag_levels = "A")
+p3 + plot_spacer() + p2 + (p4 + coord_flip()) +
+  plot_layout(widths = c(4,1),
+              heights = c(1,4))
 

### 2020-03-18 Manuscript/Manuscript.Rmd
.

@@
 p3 <- ggplot(allGroups, aes(x = medCors, fill = interval)) +
-  annotate("rect", xmin = 0.15, xmax = Inf, ymin = 0, ymax = Inf, fill = "orange", alpha = 0.5) +
+  #annotate("rect", xmin = 0.15, xmax = Inf, ymin = 0, ymax = Inf, fill = "orange", alpha = 0.5) +
   geom_density(data = filter(allGroups, interval != "mid"), alpha = 0.75, color = NA) +

### 2020-03-18 Manuscript/Manuscript.tex
.

@@
 \newcommand{\eleft}{\end{flushleft}}
+\usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
 
@@
 
-The coordinated expression of proteins is vital for many aspects of
-cellular function, from maintaining a dynamic steady state to
-differentiation of cells into specialized types in different tissues.
-The ability to regulate coordinated expression has been demonstrated to
-occur via transcription (transcription factors, chromatin regulation),
-translation (specialized ribosomes, mRNA structure/modifications), and
-degradation (E3 ligases). These processes, as well as the molecules they
-target, are all subject to noise, which has been shown to lead to
-differences in cell behavior and decisions in otherwise identical cells.
-Variability and coordinated expression (correlation) of proteins leads
-to increased population level control in binary decisions, and a
-decreased ability to execute analog signaling.
+The coordinated expression of proteins is vital for cellular function,
+from maintaining a dynamic steady state to differentiation of cells into
+specialized types and tissues. This coordinated regulation of protein
+abundance is subject to stochastic expression levels, and modules
+(pathways, complexes, etc.) have varying tolerances of noise for proper
+function ranging from precise stoichiometric coordination, to high
+levels of noise that can result in heterogeneous responses of
+homogeneous cell populations to identical stimuli (Suderman et al.,
+2017). The ability to regulate coordinated expression has been
+demonstrated to occur via transcription (transcription factors,
+chromatin regulation), translation (specialized ribosomes, mRNA
+structure/modifications), and degradation (E3 ligases), all of which are
+subject to stochastic expression and interactions. This noise has been
+shown be utilized by cells for bet hedging strategies as well as tissue
+size maintenance (more concrete examples to come). Additionally, the
+combination of variability and coordinated expression (correlation) of
+proteins can lead to increased population level control in binary
+decisions, and a decreased ability to execute analog signaling (Kovary
+et al., 2018; Suderman et al., 2017).
 
-Despite these important insights, a global understanding of the
-variability of protein expression between single cells and the
-coordinated expression of groups of proteins is largely an nonexistent.
-To systematically assess these properties of single cells, we carried
-out a proteomics experiment on exceeding large single cells, Xenopus
-laevis eggs. An advantage of this approach is that we can get around
-much of the signal to noise issues that accompany studying single cells
-due to low sample amounts. Additionally, at this stage of development
-transcription is restricted so we are able to gain insights into
-non-transcriptional control of protein expression. We also utilized
-isobaric tagging in order to measure 25 individual eggs in 5 mass
-spectrometry runs. In this study, we were able to measure the relative
-abundance \textgreater{}1000 proteins across single cells in order to
-better understand the properties of stochastic and coordinated
-expression of proteins.
+There have been numerous studies, both targeted and unbiased, of
+coordinated protein expression or expression variation independently
+that have shown the important impacts of these parameters on cellular
+function. However, both of these parameters (variation and correlation)
+act together to determine the total variance of a system. To
+systematically assess these properties of single cells at a large scale,
+we conducted shotgun proteomics analyses on \emph{Xenopus laevis} eggs
+during their first cell cycle. We have found this model to be ideal at
+this nascent stage of the single cell proteomics field because of their
+large size and tractability. An advantage of this approach is that we
+can get around much of the signal to noise issues that accompany
+studying single cells caused by low sample protein levels. Additionally,
+at this stage of development transcription is restricted, allowing for
+insights into non-transcriptional control of protein expression. By
+utilizing isobaric tagging and shotgun mass spectrometry we were able to
+measure \textgreater{}1000 proteins across 25 single cells in 5 mass
+spectrometry runs.
 
-With this dataset we have been able to, for the first time, measured the
-relationship between protein expression variance and coordinated protein
-expression on a proteome scale in single cells. We have observed that
-certain classes of proteins, including protein complexes and metabolic
-pathways, are expressed in such a way that increased coordinated
-expression is balanced by decreased variation. By doing this, cells are
-able to decrease the total variance of a given complex or pathway, an
-elegant balancing act that allows for finer control of metabolic
-throughput and controlling the number of potentially formed complexes
-though stoichiometric control. Though this kind of coordinated
-expression leads to an increase in variance in a population of cells, it
-can reduce variation within a cell.
+With this data set we have been able to, for the first time, gain
+insights into the relationship between protein expression variance,
+coordinated protein expression, and total variance within modules at a
+proteomic scale in single cells. We identified classes of proteins that
+include heteromeric complexes and metabolic pathways that are expressed
+in such a way that the increase in total variance of the system caused
+by high coordinated expression of proteins is offset by decreased
+variation, allowing for modules that rely on stoichiometric expression
+and low level of variance to maintain a low level of total system
+variance. By counter balancing coordinated expression (correlation) of
+proteins in a module with low expression variance, cells are able to
+decrease the total variance of a given complex or pathway. This elegant
+balancing act may allow for finer control of metabolic throughput and
+controlling the number of potentially formed complexes though
+stoichiometric control.
 
@@
 
-The specific requirments for coordinated protein expression of pathways
+The specific requirements for coordinated protein expression of pathways
 or complexes can vary, with some requiring strict stoichiometric
-regulation with others having much more relaxed requirements. Strict
-stoichiometry scenerios require coordinated expression of proteins
+regulation and others having much more relaxed requirements. Strict
+stoichiometry scenarios require coordinated expression of proteins
 (Figure 1A, top), and others can have uncoordinated expression (Figure
@@
 expression variation (standard deviation or coefficient of variation)
-and abundance is identical between these two scenerios, the total
+and abundance is identical between these two scenarios, the total
 variance of a pathway or complex can be significantly different due to
@@
 In order to study the relationship between protein expression variance
-and coexpression (correlated expression) of cellular pathways and
+and co-expression (correlated expression) of cellular pathways and
 complexes at the proteome level, we utilized \emph{Xenopus laevis} eggs
@@
 collected at 5 time points across the first cell cycle (0, 20, 40, 60,
-and 80 minutes), with 5 eggs at each time point. Using TMT multiplexing
-and mass spectrometry, we were able to determine the relative abundance
-of more than 1300 proteins. Expression of these proteins were largely
-invariant across the cell cycle, revealing that these highly expressed
-genes are likely not regulated by cell cycle processes. Additionally, a
-PCA analysis of these eggs showed no discernible clustering on cell
-cycle time (Fig S1).
+and 80 minutes), with 5 eggs at each time point (Fig 1D). Using TMT
+multiplexing and mass spectrometry, we were able to determine the
+relative abundance of more than 1300 proteins in single cells.
+Expression of these proteins across the time course of the cell cycle
+showed no dynamic patterns, revealing that these highly expressed
+proteins are likely not regulated by cell cycle processes. Additionally,
+a PCA analysis of these eggs showed no discernible clustering on cell
+cycle time (Fig S2).
 
-To determine the variance across these proteins, we calculated the
-variance and coefficient of variation (CV) using all time points
-collected. This showed a wide range of variation containing multiple
-distributions (Fig 1D), many of which are consistent with our previous
-study of variation using targeted mass spectrometry (Fig S2). In order
-to see if variation was a regulated process, we grouped the protein
-variation by gene ontology terms (GO terms) and plotted them in ranked
-order. We saw that in general, processes xxx were\ldots{}
+To determine the expression variability for the measured proteins
+between single cells, we calculated the variance and coefficient of
+variation (CV) across all samples. This showed a wide range of variation
+containing multiple distributions (Fig 1E), many of which are consistent
+with our previous study of variation using targeted mass spectrometry
+(Fig S2). An added benefit to measuring these proteins in parallel is
+that we are able to calculate coordinated expression of protein pairs at
+single cell resolution. Using the Pearson correlation coefficient, we
+were able to determine the coordinated expression of nearly 2 million
+protein pairs (Fig 1F-G). The distribution of correlation coefficients
+fit a normal distribution centered around 0, with the majority of
+protein pairs appearing to not show significant co-regulation. However,
+there appear to be a significant number of protein pairs containing high
+correlation coefficients, and a clustered heat map shows that many
+highly co-regulated pairs cluster together (Fig 1G).
 
-Since we were able to measure all of these proteins within single cells,
-we are able to calculate coordinated expression of protein pairs at
-single cell resolution. Using Pearson correlation coefficient, we could
-determine coordinated expression of nearly 2 million protein pairs (Fig
-1C) The distribution of correlation coefficients fit a normal
-distribution, with the majority of protein pairs appearing to not show
-significant coregulation. However, there appear to be a significant
-number of protein pairs containing high correlation coefficients, and
-the heatmap showed a lot of clustering of these highly coregulated pairs
-(Fig 1D).
+\hypertarget{variance-sum-law-reveals-modules-that-utilize-a-trade-off-between-protein-co-expression-and-noise}{%
+\subsection{Variance sum law reveals modules that utilize a trade-off
+between protein co-expression and
+noise}\label{variance-sum-law-reveals-modules-that-utilize-a-trade-off-between-protein-co-expression-and-noise}}
+
+Multicomponent modules such as heteromeric protein complexes and
+metabolic and signaling pathways have a certain tolerance for noise. Two
+components of the noise of these systems are the expression variance of
+the individual proteins, as well as the correlation or co-variance of
+expression between the proteins. This means that if a group of proteins
+have a positive correlation, then the total variance will be higher than
+if there was no correlation. This single cell proteomics data set gave
+us an excellent opportunity to see how variance and co-expression of
+proteins are related in low variance and high variance modules. The
+total variance of a module can be calculated using the variance sum law,
+where for a group of proteins \(X_{1},\dotsc,X_{n}\), the total variance
+of the group is:
+\[\sigma_{{}_{total}}^2 = \sum_{i=1}^{n} \sigma_{X_{i}}^2 + 2\sum_{i,j:i<j}^{n} \rho(X_{i},X_{j})\sigma_{X_i}\sigma_{X_j}\]
+
+To categorize the measured protein expression data into modules, we
+utilized GO term and KEGG pathway annotations. Since pathways and
+complexes have varying numbers of proteins, and the number of proteins
+in a module is highly influential on the total variance sum, we
+normalized the total variance sum by the number of components (Fig. xA
+and xB) to eliminate this confounding dependency. With the normalized
+total variance metric, we could separate modules into high and low
+variance categories (Fig 2A). Since co-expression of proteins (positive
+correlation between protein expression) increases the total variance, we
+expected to see the high variance modules to be enriched for high levels
+of both protein expression variance and co-expression (Fig 1B and 1C,
+red). However, we were surprised to see that the low variance modules
+were also enriched for high levels of co-expression of proteins, with
+some of the mean correlation coefficients of modules as high as those in
+the high variance category (Fig 1B, orange shaded region).
+
+We wondered how these protein modules dealt with the increased variance
+load caused by the high numbers of correlated proteins, so we looked at
+the pairwise relationship between the mean variance of the modules and
+the mean correlation coefficients (Fig 2D and S1). The low variance
+modules uniquely showed a significant negative correlation (R = 0.49)
+between these two metrics, showing that for modules that enforce a low
+level of total variance, co-expression of proteins within that module is
+offset by lower expression variance.
+
+\hypertarget{enforcement-of-a-trade-off-between-protein-co-expression-and-noise-increases-efficiency-of-heteromeric-protein-complexes-and-metabolic-pathways}{%
+\subsection{Enforcement of a trade-off between protein co-expression and
+noise increases efficiency of heteromeric protein complexes and
+metabolic
+pathways}\label{enforcement-of-a-trade-off-between-protein-co-expression-and-noise-increases-efficiency-of-heteromeric-protein-complexes-and-metabolic-pathways}}
+
+Interestingly, we saw the low total variance and high coordinated
+expression sub population contained many metabolic pathways as well as
+multisubunit complexes, while the high total variance and high
+coordinated expression sub-population contained many signalling pathways
+(Fig. 2D, Fig. 3E-H). Previous studies have shown that protein complex
+components are produced stoichiometricly (Taggart and Li, 2018) across
+eukaryotic species and that co-expression of pathway specific enzymes
+are tightly regulated post-transcriptionally across species
+(Jean-Beno\emph{Lalanne and James C. Taggart and Monica S. Guo and Lydia
+Herzel and Ariel Schieler and Gene-Wei Li}, 2018). This led us to
+wondered what the implications were for the synergy of maintaing low
+total variance with high levels of co-expression of the members of
+protein complexes and metabolic pathways.
+
+To test why complexes observed in our data set and others, (i.e.~core
+members of the ribosome (Fig 4A)), coordinate co-expression and
+expression variation, we imagined an idealized 10 member heteromeric
+complex (Fig 4B). Coordinated expression could allow for tight control
+of the total number of assembled complexes. However, this on it's own is
+not sufficient since high levels of individual protein variance would
+result in an increased number of partially assembled complexes. In this
+idealized complex, we titrated the correlation coefficients between all
+10 members along with the individual expression variation and saw that
+increased correlations and decreasing variance had a synergistic effect
+in allowing for the control of the total number of fully assembled
+complexes (Fig 4C.). Additionally, there is a large regime with high
+correlations and low variation that allow for low total variance of the
+system.
+
+We also observed a number of metabolic pathways that had similar
+variation and correlation properties as the protein complexes, such as
+the pentose phosphate pathway and glycolysis. Both of these pathways
+have a large number of reversible enzymes. For a reversible enzymatic
+pathway, the concentrations of the enzymes dictate the rate of the
+reaction, and the concentrations of the reaction products relative to
+the substrates can dictate the direction of the reaction equilibrium.
+Therefore, strict balance of the concentration of enzymes using low
+levels of variance and strict control of the relative abundances of the
+enzymes is a potentially important step in regulating metabolic pathway
+output. To test this hypothesis, we constructed an idealized metabolic
+model (or a physiological model??) and tested the ability of the pathway
+to maximize output (Fig 5x). (haven't done this, would like feedback /
+help).
+
+\hypertarget{discussion}{%
+\subsection{Discussion}\label{discussion}}
+
+\newpage
+
+\hypertarget{figures}{%
+\section{Figures}\label{figures}}
 
@@
 correlations of 0 and 0.8. B) Pairwise plots of the two proteins in the
-single cells between uncorrelated and correlated scenerios. C) The total
-variance between the uncorrelated and correlated scenerios from A and
-B.}
+single cells between uncorrelated and correlated scenarios. C) The total
+variance between the uncorrelated and correlated scenarios from A and B.
+D) To measure protein expression variance and coordinated expression, 25
+eggs at 5 time points during the first cell cycle were analyzed with
+multiplexed shotgun mass spectrometry. E) Histogram of the coefficient
+of variation of all measured proteins. F) Histogram of the pairwise
+correlation coefficients between all measured proteins. G) Heat-map of
+the pairwise correlation coefficients between all measured protein
+pairs.}
 \end{figure}
 
-\hypertarget{protein-modules-display-different-levels-of-regulation-of-noise-and-co-expression-of-constituents}{%
-\subsection{Protein modules display different levels of regulation of
-noise and co-expression of
-constituents}\label{protein-modules-display-different-levels-of-regulation-of-noise-and-co-expression-of-constituents}}
+\newpage
 
-To investigate whether coordinated expression goes beyond just protein
-pairs, we wanted to see if pathways, complexes, and other modules had
-proteins that with coordinated expression. To answer this question, we
-grouped proteins by GO terms and calculated the median correlation
-coefficient across all pairs of proteins within the term. When the
-ranked media correlation is plotted it's clear that there are a
-significant number of GO terms with higher than expected levels of
-coordinate expression.
+\begin{figure}
+\centering
+\includegraphics{output/figures/figure_2-1.pdf}
+\caption{A) Histogram of the normalized variation values across all of
+the GO terms and KEGG pathways. The bottom and top \textasciitilde{}15\%
+are colored blue and red respectively. B) Distributions of the mean
+correlation coefficient across all groups of proteins. Both the top and
+bottom 15\% of normalized variation groups had peaks of high levels of
+mean correlation coefficients. C) Distributions of the mean coefficient
+of variation across all groups of proteins. The top and bottom
+\textasciitilde{}15\% of normalized variation groups were largely
+separated. D) Pairwise plot between the mean correlation and mean
+coefficient of variation within each GO term and KEGG pathway with color
+coding from A. Interestingly, both the top and bottom 10\% of normalized
+variation have groups of proteins with high levels of correlation,
+despite high correlation adding to total variance. In the lowest 10\% of
+normalized variation groups there appears to be a negative correlation
+between mean correlation and variation.}
+\end{figure}
 
-\hypertarget{total-variance-metric-reveals-certain-modules-enforce-a-tradeoff-between-protein-co-expression-and-noise}{%
-\subsection{Total variance metric reveals certain modules enforce a
-tradeoff between protein co-expression and
-noise}\label{total-variance-metric-reveals-certain-modules-enforce-a-tradeoff-between-protein-co-expression-and-noise}}
+\newpage
 
-Variance Sum Law --- Dependent Case
+\begin{figure}
+\centering
+\includegraphics{output/figures/figure_3-1.pdf}
+\caption{Density plots of correlation coefficients select protein groups
+color coded by low (blue) and high (red) normalized protein variance,
+with the distribution of all measured correlation coefficients colored
+in grey (rug plot to show the measured correlations coefficients shown
+on plot). Below each density plot is a box plot.}
+\end{figure}
 
-The sum of variance for a group of proteins \(X_{1},\dotsc,X_{n}\) with
-dependent relationships:
+\newpage
 
-\[\sigma_{{}_{total}} = \sqrt{var\left(\sum_{i=1}^n X_{i}\right)}\]
+\begin{figure}
+\centering
+\includegraphics{output/figures/figure_4-1.pdf}
+\caption{A) A representative heat-map of one of the low normalized
+variation groups with high levels of correlated proteins (Large
+ribosomal subunit). B) A simple model showing the percent of potentially
+assembled subunits of an idealized heteromeric 10 subunit complex. Both
+the variation of expression between the subunits as well as the
+coordinated expression between them can effect the percent of maximum
+assembled complexes. C) Heat-map showing the total variance as a
+function of coefficient of variation and correlation from the idealized
+model in B).}
+\end{figure}
 
-\[\sigma_{{}_{norm}} = \dfrac{\sigma_{{}_{total}}}{n}\]
+\newpage
 
-When calculating the total variance of a pathway or general group of
-proteins, the total variance is the square root of the sums of the
-squares of the coefficient of variance. However, if any of the proteins
-are correlated in expression, an extra term must be added to account for
-this correlation. This means that if a group of proteins have a positive
-correlation, then the total variance will be higher than if there was no
-correlation. In order to study how variance and coordinated expression
-are related between single cells, we wanted to calculate a normalized
-total variance metric. Since we were not interested in the increase of
-variance due to the number of proteins in a module, we calculated the
-total variance using the coefficient of variation and correlation
-coefficients and normalize to the total number of proteins in the
-module. When total variance is plotted against the number of proteins,
-there is a clear linear relationship between the two. Once we normalize
-to the total number of proteins this relationship is lost and we are
-left with the effects of coefficient of variation and correlation.
+\hypertarget{supplementary-material}{%
+\section*{Supplementary material}\label{supplementary-material}}
+\addcontentsline{toc}{section}{Supplementary material}
 
-\hypertarget{low-total-variance-modules-are-enriched-for-metabolic-pathways-and-heteromeric-protein-complexes}{%
-\subsection{Low total variance modules are enriched for metabolic
-pathways and heteromeric protein
-complexes}\label{low-total-variance-modules-are-enriched-for-metabolic-pathways-and-heteromeric-protein-complexes}}
+\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}
 
-\hypertarget{enforcement-of-a-tradeoff-between-protein-co-expression-and-noise-increases-efficiency-of-metabolic-pathways-and-heteromeric-protein-complexes}{%
-\subsection{Enforcement of a tradeoff between protein co-expression and
-noise increases efficiency of metabolic pathways and heteromeric protein
-complexes}\label{enforcement-of-a-tradeoff-between-protein-co-expression-and-noise-increases-efficiency-of-metabolic-pathways-and-heteromeric-protein-complexes}}
+\begin{figure}
+\centering
+\includegraphics{output/figures/sup_fig_1-1.pdf}
+\caption{Pairwise plots of the mean correlation coefficient and mean
+coefficient of variation of protein groups with linear fit. The low
+variance protein groups have a significant negative correlation between
+these two variables, showing that for protein modules that require low
+variance and high co-expression, showing that these two
+parameters\ldots{} R (high, low, med) = (0.057, -0.485, 0.071)}
+\end{figure}
+
+\newpage
 
@@
 Xenopus egg extracts were prepared based on modifications of a previous
-protocol (Tsai et al, 2014). All of the animal protocols used in this
+protocol (Tsai et al., 2014). All of the animal protocols used in this
 manuscript were approved by the Stanford University Administrative Panel
@@
 for 3 min, after which they were washed four times with 1× MMR buffer.
-Single eggs were collected at their respective timepoints and placed
+Single eggs were collected at their respective time-points and placed
 into 600uL tubes and snap frozen in liquid nitrogen before being stored
@@
 protease inhibitor cocktail (Sigma). The lysate was then placed in a 400
-uLnatural polyethylene microcentrifuge tube (E\&K Scientific \#485050)
+uLnatural polyethylene micro-centrifuge tube (E\&K Scientific \#485050)
 and spun at 15,000 g in a right angle centrifuge (Beckman Microfuge E)
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Lalanne_2018}{}%
+Jean-Beno\emph{Lalanne and James C. Taggart and Monica S. Guo and Lydia
+Herzel and Ariel Schieler and Gene-Wei Li} (2018). Evolutionary
+convergence of pathway-specific enzyme expression stoichiometry. Cell
+\emph{173}, 749--761.e38.
+
 \leavevmode\hypertarget{ref-Kovary_2018}{}%
@@
 signaling control. Molecular Systems Biology \emph{14}.
+
+\leavevmode\hypertarget{ref-Suderman_2017}{}%
+Suderman, R., Bachman, J.A., Smith, A., Sorger, P.K., and Deeds, E.J.
+(2017). Fundamental trade-offs between information flow in single cells
+and cellular populations. Proceedings of the National Academy of
+Sciences \emph{114}, 5755--5760.
+
+\leavevmode\hypertarget{ref-Taggart_2018}{}%
+Taggart, J.C., and Li, G.-W. (2018). Production of protein-complex
+components is stoichiometric and lacks general feedback regulation in
+eukaryotes. Cell Systems \emph{7}, 580--589.e4.
+
+\leavevmode\hypertarget{ref-Tsai_2014}{}%
+Tsai, T.Y.-C., Theriot, J.A., and Ferrell, J.E. (2014). Changes in
+oscillatory dynamics in the cell cycle of early xenopus laevis embryos.
+PLoS Biology \emph{12}, e1001788.
 
@@
 attached base packages:
-[1] stats     graphics  grDevices utils     datasets  methods   base
+ [1] parallel  stats4    grid      stats     graphics  grDevices utils    
+ [8] datasets  methods   base
 
 other attached packages:
-[1] knitcitations_1.0.10 knitr_1.28
+ [1] circlize_0.4.8       pheatmap_1.0.12      magick_2.3          
+ [4] KEGG.db_3.2.3        latex2exp_0.4.0      ggplotify_0.0.4     
+ [7] knitcitations_1.0.10 mvtnorm_1.1-0        patchwork_1.0.0     
+[10] cowplot_1.0.0        plotly_4.9.2         ggrepel_0.8.1       
+[13] GO.db_3.10.0         org.Xl.eg.db_3.10.0  AnnotationDbi_1.48.0
+[16] IRanges_2.20.2       S4Vectors_0.24.3     Biobase_2.46.0      
+[19] BiocGenerics_0.32.0  VIM_5.1.0            data.table_1.12.8   
+[22] colorspace_1.4-1     GGally_1.4.0         ComplexHeatmap_2.2.0
+[25] RColorBrewer_1.1-2   forcats_0.5.0        stringr_1.4.0       
+[28] dplyr_0.8.5          purrr_0.3.3          readr_1.3.1         
+[31] tidyr_1.0.2          tibble_2.1.3         ggplot2_3.3.0.9000  
+[34] tidyverse_1.3.0      knitr_1.28
 
 loaded via a namespace (and not attached):
- [1] Rcpp_1.0.3        lubridate_1.7.4   digest_0.6.24     plyr_1.8.5       
- [5] R6_2.4.1          jsonlite_1.6.1    formatR_1.7       magrittr_1.5     
- [9] evaluate_0.14     bibtex_0.4.2.2    httr_1.4.1        rlang_0.4.4      
-[13] stringi_1.4.6     curl_4.3          xml2_1.2.2        rmarkdown_2.1    
-[17] tools_3.6.1       stringr_1.4.0     RefManageR_1.2.12 xfun_0.12        
-[21] yaml_2.2.1        compiler_3.6.1    htmltools_0.4.0  
+  [1] readxl_1.3.1        backports_1.1.5     Hmisc_4.3-1        
+  [4] plyr_1.8.5          lazyeval_0.2.2      splines_3.6.1      
+  [7] sp_1.4-1            digest_0.6.25       htmltools_0.4.0    
+ [10] wesanderson_0.3.6   fansi_0.4.1         checkmate_2.0.0    
+ [13] magrittr_1.5        memoise_1.1.0       cluster_2.1.0      
+ [16] openxlsx_4.1.4      modelr_0.1.6        jpeg_0.1-8.1       
+ [19] blob_1.2.1          rvest_0.3.5         haven_2.2.0        
+ [22] xfun_0.12           crayon_1.3.4        jsonlite_1.6.1     
+ [25] survival_3.1-8      zoo_1.8-7           glue_1.3.1         
+ [28] gtable_0.3.0        GetoptLong_0.1.8    car_3.0-6          
+ [31] shape_1.4.4         DEoptimR_1.0-8      abind_1.4-5        
+ [34] scales_1.1.0        DBI_1.1.0           bibtex_0.4.2.2     
+ [37] Rcpp_1.0.3          htmlTable_1.13.3    viridisLite_0.3.0  
+ [40] laeken_0.5.1        clue_0.3-57         gridGraphics_0.5-0 
+ [43] foreign_0.8-75      bit_1.1-15.2        Formula_1.2-3      
+ [46] rsvg_1.3            vcd_1.4-5           htmlwidgets_1.5.1  
+ [49] httr_1.4.1          acepack_1.4.1       ellipsis_0.3.0     
+ [52] pkgconfig_2.0.3     reshape_0.8.8       farver_2.0.3       
+ [55] nnet_7.3-13         dbplyr_1.4.2        tidyselect_1.0.0   
+ [58] labeling_0.3        rlang_0.4.4         munsell_0.5.0      
+ [61] cellranger_1.1.0    tools_3.6.1         cli_2.0.2          
+ [64] generics_0.0.2      RSQLite_2.2.0       ranger_0.12.1      
+ [67] broom_0.5.5         evaluate_0.14       yaml_2.2.1         
+ [70] RefManageR_1.2.12   bit64_0.9-7         fs_1.3.1           
+ [73] zip_2.0.4           robustbase_0.93-5   nlme_3.1-144       
+ [76] formatR_1.7         xml2_1.2.2          compiler_3.6.1     
+ [79] rstudioapi_0.11     curl_4.3            png_0.1-7          
+ [82] e1071_1.7-3         reprex_0.3.0        stringi_1.4.6      
+ [85] highr_0.8           lattice_0.20-40     Matrix_1.2-18      
+ [88] vctrs_0.2.3         pillar_1.4.3        lifecycle_0.1.0    
+ [91] BiocManager_1.30.10 lmtest_0.9-37       GlobalOptions_0.1.1
+ [94] latticeExtra_0.6-29 R6_2.4.1            gridExtra_2.3      
+ [97] rio_0.5.16          codetools_0.2-16    boot_1.3-24        
+[100] MASS_7.3-51.5       assertthat_0.2.1    rjson_0.2.20       
+[103] withr_2.1.2         mgcv_1.8-31         hms_0.5.3          
+[106] rpart_4.1-15        class_7.3-15        rmarkdown_2.1      
+[109] rvcheck_0.1.8       carData_3.0-3       base64enc_0.1-3    
+[112] lubridate_1.7.4    
 \end{verbatim}

### 2020-03-18 Manuscript/test-tikz.tex
.

% Created by tikzDevice version 0.12.3 on 2020-03-05 17:30:43
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{xcolor}
\usetikzlibrary{calc}
\usepackage[active,tightpage,psfixbb]{preview}
\PreviewEnvironment{pgfpicture}\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,505.89);
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (480.69,456.69);
\definecolor{drawColor}{RGB}{0,0,255}

\path[draw=drawColor,line width= 0.8pt,line join=round,line cap=round] ( 65.18,442.04) --
	( 78.50,436.06) --
	( 91.82,430.08) --
	(105.13,424.10) --
	(118.45,418.12) --
	(131.77,412.14) --
	(145.09,406.16) --
	(158.40,400.17) --
	(171.72,394.19) --
	(185.04,388.21) --
	(198.36,382.23) --
	(211.67,376.25) --
	(224.99,370.27) --
	(238.31,364.29) --
	(251.63,358.31) --
	(264.94,352.32) --
	(278.26,346.34) --
	(291.58,340.36) --
	(304.90,334.38) --
	(318.22,328.40) --
	(331.53,322.42) --
	(344.85,316.44) --
	(358.17,310.46) --
	(371.49,304.48) --
	(384.80,298.49) --
	(398.12,292.51) --
	(411.44,286.53) --
	(424.76,280.55) --
	(438.07,274.57) --
	(451.39,268.59) --
	(464.71,262.61);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,505.89);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 51.86, 61.20) -- (451.39, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 51.86, 61.20) -- ( 51.86, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (118.45, 61.20) -- (118.45, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.04, 61.20) -- (185.04, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.63, 61.20) -- (251.63, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (318.22, 61.20) -- (318.22, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.80, 61.20) -- (384.80, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (451.39, 61.20) -- (451.39, 55.20);
\end{scope}
\end{tikzpicture}

\end{document}

### 2020-03-18 scratch.R

@@ -54,3 +54,20 @@ int <- matrix(data = c(0,1,0,
                        1,0,0),
               nrow = 3)
 mat <- cor[1:3,1:3]
+
+
+
+
+
+piris <- ggplot(iris, aes(Sepal.Length, Sepal.Width, colour = Species)) +
+  geom_point() + 
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+ggMarginal(piris, groupColour = F, groupFill = TRUE, color = "transparent")
+
+ggMarginal(p2, groupColour = F, groupFill = TRUE, color = "transparent")
+
+p3 + plot_spacer() + p2 + (p4 + coord_flip()) +
+  plot_layout(widths = c(4,1),
+              heights = c(1,4))

### 2020-03-18 Manuscript/Manuscript.Rmd
.

@@
   mutate(abundance = abundance / mean(abundance, na.rm = T)) %>% 
-  filter(n() > 15) %>% # comment out if using KNN
+  # filter(n() > 15) %>% # comment out if using KNN
   group_by(cell) %>% 
@@
   spread(cell, abundance)
-  # spread(cell, abundance) %>%
-  # kNN(data = ., k = 1, impNA = TRUE) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)
+  spread(cell, abundance) %>%
+  kNN(data = ., k = 1, impNA = TRUE) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)
 

### 2020-03-19 Manuscript/Manuscript.Rmd
.

@@
   dplyr::select(cell, Uniprot, abundance) %>% 
-  spread(cell, abundance)
+  # spread(cell, abundance)
   spread(cell, abundance) %>%
@@
 ```{r figure_1, echo=FALSE, fig.height=4.5, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}
-#allGroups <- readRDS("allGroups.RDS")
+allGroups <- readRDS("allGroups.RDS")
 
@@
 
-p2 <- ggplot(filter(allGroups, n > 5), aes(x = medCors, y = medCV, color = interval)) +
+p2 <- ggplot(filter(allGroups, n > 5), aes(x = medCors, y = medCV, color = interval, shape = type)) +
   geom_point(data = filter(allGroups, interval == "mid", n > 4),

### 2020-03-19 Manuscript/Manuscript.tex
.

@@
 other attached packages:
- [1] circlize_0.4.8       pheatmap_1.0.12      magick_2.3          
- [4] KEGG.db_3.2.3        latex2exp_0.4.0      ggplotify_0.0.4     
- [7] knitcitations_1.0.10 mvtnorm_1.1-0        patchwork_1.0.0     
-[10] cowplot_1.0.0        plotly_4.9.2         ggrepel_0.8.1       
-[13] GO.db_3.10.0         org.Xl.eg.db_3.10.0  AnnotationDbi_1.48.0
-[16] IRanges_2.20.2       S4Vectors_0.24.3     Biobase_2.46.0      
-[19] BiocGenerics_0.32.0  VIM_5.1.0            data.table_1.12.8   
-[22] colorspace_1.4-1     GGally_1.4.0         ComplexHeatmap_2.2.0
-[25] RColorBrewer_1.1-2   forcats_0.5.0        stringr_1.4.0       
-[28] dplyr_0.8.5          purrr_0.3.3          readr_1.3.1         
-[31] tidyr_1.0.2          tibble_2.1.3         ggplot2_3.3.0.9000  
-[34] tidyverse_1.3.0      knitr_1.28
+ [1] circlize_0.4.8       magick_2.3           KEGG.db_3.2.3       
+ [4] latex2exp_0.4.0      ggplotify_0.0.4      knitcitations_1.0.10
+ [7] mvtnorm_1.1-0        patchwork_1.0.0      cowplot_1.0.0       
+[10] plotly_4.9.2         ggrepel_0.8.1        GO.db_3.10.0        
+[13] org.Xl.eg.db_3.10.0  AnnotationDbi_1.48.0 IRanges_2.20.2      
+[16] S4Vectors_0.24.3     Biobase_2.46.0       BiocGenerics_0.32.0 
+[19] VIM_5.1.0            data.table_1.12.8    colorspace_1.4-1    
+[22] GGally_1.4.0         ComplexHeatmap_2.2.0 RColorBrewer_1.1-2  
+[25] forcats_0.5.0        stringr_1.4.0        dplyr_0.8.5         
+[28] purrr_0.3.3          readr_1.3.1          tidyr_1.0.2         
+[31] tibble_2.1.3         ggplot2_3.3.0.9000   tidyverse_1.3.0     
+[34] knitr_1.28
 
@@
   [1] readxl_1.3.1        backports_1.1.5     Hmisc_4.3-1        
-  [4] plyr_1.8.5          lazyeval_0.2.2      splines_3.6.1      
-  [7] sp_1.4-1            digest_0.6.25       htmltools_0.4.0    
+  [4] plyr_1.8.5          lazyeval_0.2.2      sp_1.4-1           
+  [7] splines_3.6.1       digest_0.6.25       htmltools_0.4.0    
  [10] wesanderson_0.3.6   fansi_0.4.1         checkmate_2.0.0    

### 2020-05-03 Manuscript/Manuscript.Rmd
updated

--- 2020-03-18 Manuscript/Manuscript.Rmd
@@
 output:
-  # word_document:
+  word_document:
   # html_document:  
-  pdf_document:
-    fig_caption: yes
-    fig_width: 7
-    fig_height: 7
-    keep_tex: yes
-    number_sections: no
-    template: manuscript.latex
-    includes:
-      in_header: header.tex
+  # pdf_document:
+  #   fig_caption: yes
+  #   fig_width: 7
+  #   fig_height: 7
+  #   keep_tex: yes
+  #   number_sections: no
+  #   template: manuscript.latex
+  #   includes:
+  #     in_header: header.tex
 
@@
 Here, we measure cell-to-cell variability of relative protein abundance using quantitative proteomics of individual *Xenopus laevis* eggs and show that proteins involved in metabolic pathways or members of heteromeric complexes tend to have high correlations with other members of those pathways/complexes.
-Our previous work highlighted the fact that correlated expression increases the total variation of a pathway, so one would reason that certain pathways or complexes would need to compensate of this extra source of by reducing the variation of expression of these pathways or complexes.
-To test this we computed total variance score that took into account both the coefficient of variance and correlations between proteins in a pathway and found that the lower 10% of GO terms were highly enriched for metabolic pathways.
+Our previous work highlighted the fact that correlated expression increases the total variation of a pathway, so one would reason that certain pathways or complexes would need to compensate this extra source of variation for these pathways or complexes.
+To test this we computed total variance score that took into account both the coefficients of variation and correlation coefficients between proteins in a pathway and found that the lower 10% of GO terms were highly enriched for metabolic pathways.
 When we looked at the relationship between CV and R between these GO terms we found a negative relationship between them, demonstrating that increased correlation needs to come at the expense of decreased variance.
@@
 library(ComplexHeatmap)
+library(circlize)
 library(GGally)
@@
   mutate(abundance = abundance / mean(abundance, na.rm = T)) %>%
-  # filter(n() > 15) %>% # comment out if using KNN
+  #filter(n() >= 10) %>% # comment out if using KNN
+  filter(n() > 5) %>%
   group_by(cell) %>%
@@
   dplyr::select(cell, Uniprot, abundance) %>%
+  # spread(cell, abundance)
+  #spread(cell, abundance) %>%
+  spread(Uniprot, abundance) %>%
+  kNN(data = ., k = 3, impNA = TRUE) %>% as_tibble() %>%
+  #dplyr::select(Uniprot:e_131_80)
+  dplyr::select(cell:UBE2I_HUMAN) %>%
+  pivot_longer(2:ncol(.), names_to = "Uniprot", values_to = "abundance") %>%
   spread(cell, abundance)
-  spread(cell, abundance) %>%
-  kNN(data = ., k = 1, impNA = TRUE) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)
+
+# protImpute <- pepTidy %>%
+#   group_by(peptide,run) %>%
+#   mutate(intensity = intensity / intensity[label == "126"]) %>%
+#   ungroup() %>%
+#   filter(run != "ctrl", condition != "ctrl") %>%
+#   dplyr::filter(abs((intensity)) < 2*sd((pepTidy$intensity), na.rm = T)) %>%
+#   group_by(cell, run, label, condition, Uniprot) %>%
+#   summarise(abundance = mean(intensity, na.rm = T)) %>% ungroup() %>%
+#   group_by(Uniprot) %>%
+#   mutate(abundance = abundance / mean(abundance, na.rm = T)) %>%
+#   #filter(n() >= 10) %>% # comment out if using KNN
+#   group_by(cell) %>%
+#   mutate(abundance = abundance / mean(abundance, na.rm = T)) %>%
+#   ungroup() %>%
+#   dplyr::select(cell, Uniprot, abundance) %>%
+#   spread(cell, abundance) %>%
+#   kNN(data = ., k = 1, impNA = TRUE) %>%
+#   as_tibble() %>%
+#   dplyr::select(Uniprot:e_131_80)
 
@@
   rowwise() %>% mutate(medCors = mean(cors, na.rm = TRUE),
+                       sdCors = sd(cors),
                        medCV = mean(cv, na.rm = TRUE),
+                       sdCV = sd(cv),
                        medSD = mean(sd, na.rm = TRUE)) %>%
@@
 
+# 3 Takeaway Points
+
+* To understand the total variance of a protein module such as a metabolic pathway or protein complex, a single cell approach capable of measuring the varaince and correlation of proteins is necessary.
+* For core housekeeping modules such as metabolic pathways and heteromeric complexes like the ribosome, individual cells enforce a balance between protein expression variance and co-expression variance.
+* For certain modules increasing co-expression and reducing expression variance may allow for better control and more efficient opperation.
+
 # Introduction
@@
 The coordinated expression of proteins is vital for cellular function, from maintaining a dynamic steady state to differentiation of cells into specialized types and tissues.
+This is especailly critical for heteromeric protein complexes and metabolic and signaling pathways where many proteins are required to work in coordination in order for the system to function.
+For prokaryotes (and in rare instances eukaryotes), this is accomplished in a rather uncomplicated fashion through the use of opperons that can be co-translated.
+Eukaryotes for the most part rely on transcriptional activators and repressors to coordinate the expression of these genes that may exist in clusters or scattered throughout the genome.
+The bursty nature of transcriptional dynamics, which is less of an issue in the case of opperons, leads to stochastic production of mRNA of each component of the pathway.
+There exist buffering mechanisms such as bottlenecking in post-processing of nacient mRNA and nuclear export.
+There also exists evidence of co-translation through specialized ribosomes in eukaryotes that can preferentially translate members of the same pathway.
+
 This coordinated regulation of protein abundance is subject to stochastic expression levels, and modules (pathways, complexes, etc.) have varying tolerances of noise for proper function ranging from precise stoichiometric coordination, to high levels of noise that can result in heterogeneous responses of homogeneous cell populations to identical stimuli `r citep("10.1073/pnas.1615660114")`.
 The ability to regulate coordinated expression has been demonstrated to occur via transcription (transcription factors, chromatin regulation), translation (specialized ribosomes, mRNA structure/modifications), and degradation (E3 ligases), all of which are subject to stochastic expression and interactions.
-This noise has been shown be utilized by cells for bet hedging strategies as well as tissue size maintenance (more concrete examples to come).
+This noise has been shown to be utilized by cells for bet hedging strategies as well as tissue size maintenance (more concrete examples to come).
 Additionally, the combination of variability and coordinated expression (correlation) of proteins can lead to increased population level control in binary decisions, and a decreased ability to execute analog signaling `r citep(c("doi.org/10.15252/msb.20177997", "10.1073/pnas.1615660114"))`.
@@
 However, both of these parameters (variation and correlation) act together to determine the total variance of a system.
+Additionally, the nature of these two sources of variance can have very different impacts on the behavior or function of different systems.
+For example, high amounts of variance can decrease the output of a metabolic pathway but can increase the population level control of a binary signalling pathway.
+On the other hand, high levels of correlated expression, while technically increasing the variance of a system, can increase the output of metabolic pathways.
+We wanted to study these two sources of variation to identify instances where there may be a trade off between them or if they're used in coordination with one another.
+
 To systematically assess these properties of single cells at a large scale, we conducted shotgun proteomics analyses on *Xenopus laevis* eggs during their first cell cycle.
@@
 By counter balancing coordinated expression (correlation) of proteins in a module with low expression variance, cells are able to decrease the total variance of a given complex or pathway.
-This elegant balancing act may allow for finer control of metabolic throughput and controlling the number of potentially formed complexes though stoichiometric control.
+This elegant balancing act may allow for finer control of metabolic throughput and controlling the number of potentially formed complexes through stoichiometric control.
 
@@
 At the single cell level, coordinated expression will result in a high correlation coefficient, whereas un-coordinated expression will result in a low correlation coefficient (Figure 1B).
-Even when the expression variation (standard deviation or coefficient of variation) and abundance is identical between these two scenarios, the total variance of a pathway or complex can be significantly different due to the variance sum law: $var\left(\sum_{i=1}^n X_{i}\right) = \sum_{i=1}^{n} var(X_{i}) + 2\sum_{i,j:i<j}^{n} cov(X_{i},X_{j})$ (Figure 1C).
+Even when the expression variation (standard deviation or coefficient of variation) and abundances are identical between these two scenarios, the total variance of a pathway or complex can be significantly different due to the variance sum law: $var\left(\sum_{i=1}^n X_{i}\right) = \sum_{i=1}^{n} var(X_{i}) + 2\sum_{i,j:i<j}^{n} cov(X_{i},X_{j})$ (Figure 1C).
+The nature of these two sources of variation that contribute to the variance sum law can have very different impacts on the behavior of systems.
+This means that for a given pathway with fixed total variance, the variance and correlation of the members of that system can be changed and the behavior can be impacted (needs big time re-wording, and first figure should show this...).
 Our previous work has shown that this effect is important in single cell pathway activation dynamics `r citep("doi.org/10.15252/msb.20177997")`.
@@
 To determine the expression variability for the measured proteins between single cells, we calculated the variance and coefficient of variation (CV) across all samples.
-This showed a wide range of variation containing multiple distributions (Fig 1E), many of which are consistent with our previous study of variation using targeted mass spectrometry (Fig S2).
+This showed a wide range of variation containing multiple distributions (Fig SXX), many of which are consistent with our previous study of variation using targeted mass spectrometry (Fig SXX).
 An added benefit to measuring these proteins in parallel is that we are able to calculate coordinated expression of protein pairs at single cell resolution.
-Using the Pearson correlation coefficient, we were able to determine the coordinated expression of nearly 2 million protein pairs (Fig 1F-G).
+Using the Pearson correlation coefficient, we were able to determine the coordinated expression of nearly 2 million protein pairs (Fig SXX).
 The distribution of correlation coefficients fit a normal distribution centered around 0, with the majority of protein pairs appearing to not show significant co-regulation.
-However, there appear to be a significant number of protein pairs containing high correlation coefficients, and a clustered heat map shows that many highly co-regulated pairs cluster together (Fig 1G).
+However, there appear to be a significant number of protein pairs containing high correlation coefficients, and a clustered heat map shows that many highly co-regulated pairs cluster together (Fig SXX).
+
+In order to study how protein expression variation, co-expression, and total variance are related in protein modules, we utilized the KEGG ID and GO Term databases to categorize the proteins into modules (Fig 1E).
+For each module the coefficient of variance and pairwise correlation coefficients for the proteins were used to calculate the normalized total variance of the module.
 
@@
 Multicomponent modules such as heteromeric protein complexes and metabolic and signaling pathways have a certain tolerance for noise.
-Two components of the noise of these systems are the expression variance of the individual proteins, as well as the correlation or co-variance of expression between the proteins.
+Two components of the noise of these systems are the expression variance of the individual proteins, as well as the correlation of expression between the proteins.
 This means that if a group of proteins have a positive correlation, then the total variance will be higher than if there was no correlation.
@@
 Since co-expression of proteins (positive correlation between protein expression) increases the total variance, we expected to see the high variance modules to be enriched for high levels of both protein expression variance and co-expression (Fig 1B and 1C, red).
-However, we were surprised to see that the low variance modules were also enriched for high levels of co-expression of proteins, with some of the mean correlation coefficients of modules as high as those in the high variance category (Fig 1B, orange shaded region).
+However, we were surprised to see that the low variance modules were also enriched for high levels of co-expression of proteins, with some of the mean correlation coefficients of modules as high as those in the high variance category (Fig 2B).
 
@@
 Previous studies have shown that protein complex components are produced stoichiometricly `r citep("doi.org/10.1016/j.cels.2018.11.003")` across eukaryotic species and that co-expression of pathway specific enzymes are tightly regulated post-transcriptionally across species `r citep("doi.org/10.1016/j.cell.2018.03.007")`.
-This led us to wondered what the implications were for the synergy of maintaing low total variance with high levels of co-expression of the members of protein complexes and metabolic pathways.
+This led us to wonder what the implications were for the synergy of maintaing low total variance with high levels of co-expression of the members of protein complexes and metabolic pathways.
 
@@
 
-\newpage
+#\newpage
 
@@
 
-```{r figure_1, echo=FALSE, fig.height=4.5, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}
+```{r figure_1, echo=FALSE, fig.height=9, fig.width=8.5, message=FALSE, warning=FALSE, fig.cap=fig_legend_1}
 #allGroups <- readRDS("allGroups.RDS")
@@
   theme_minimal() +
-  theme(legend.position = "none")
+  theme(legend.position = "bottom")
 
@@
 
-top / bottom + plot_layout(nrow = 2,
-                           heights = c(1,0.8)) +
-  plot_annotation(tag_levels = "A")
+# top / bottom + plot_layout(nrow = 2,
+#                            heights = c(1,0.8)) +
+#   plot_annotation(tag_levels = "A")
 
-fig_legend_1 <- c("A) Simulated expression of two proteins over time in a single cell with a mean of 100 and a standard deviation of 30 with expression correlations of 0 and 0.8.
+cowplot::ggdraw() + cowplot::draw_image(magick::image_read_pdf("../images/Fig_1_new.pdf"))
+
+fig_legend_1 <- c("Single cell approach to study expression variation and correlation at a proteomic scale: A) Simulated expression of two proteins over time in a single cell with means and standard deviations of 100 and 30 respectively and correlations of 0 and 0.8.
 B) Pairwise plots of the two proteins in the single cells between uncorrelated and correlated scenarios.
@@
 D) To measure protein expression variance and coordinated expression, 25 eggs at 5 time points during the first cell cycle were analyzed with multiplexed shotgun mass spectrometry.
-E) Histogram of the coefficient of variation of all measured proteins.
-F) Histogram of the pairwise correlation coefficients between all measured proteins.
-G) Heat-map of the pairwise correlation coefficients between all measured protein pairs.")
+E) Proteins were grouped by KEGG IDs and GO terms.
+The expression variation and pairwise correlation coefficients were calculated for all proteins within each group (4 of 1225 shown) which were then used to calcuate the normalized total variance of each group.")
 
@@
 
-```{r figure_2, echo=FALSE, fig.height=7, fig.width=6, message=FALSE, warning=FALSE, eval=TRUE, fig.cap=fig_legend_2}
+# New Figure 1 sub-plots
+```{r eval=FALSE, include=FALSE}
+#hm_groups <- c("03010","00030","GO:0046496","00071","GO:0006096")
+hm_groups <- c("GO:0015934", "04012", "GO:0007264", "GO:0010608", "GO:0031329", "GO:0000786", "00030")
+#hm_groups <- allGroups$accession
+
+rm(prot_hm)
+for(i in seq_along(hm_groups)){
+  if(!exists("prot_hm")){
+    temp_prots <- allGroups %>% 
+      filter(accession == hm_groups[i]) %>% 
+      pull(Uniprot) %>% unlist()
+
+    prot_hm <- protImpute %>% 
+      filter(Uniprot %in% temp_prots) %>% 
+      select(-Uniprot) %>%
+      as.matrix()
+
+    Group <- rep(filter(allGroups, 
+                        accession == hm_groups[i])$name,
+                 times = length(temp_prots))
+
+    mat_labels <- temp_prots
+  } else{
+    temp_prots <- allGroups %>% 
+      filter(accession == hm_groups[i]) %>% 
+      pull(Uniprot) %>% unlist()
+
+    temp_prot_hm <- protImpute %>% 
+      filter(Uniprot %in% temp_prots) %>% 
+      select(-Uniprot) %>%
+      as.matrix()
+
+    prot_hm <- rbind(prot_hm, temp_prot_hm)
+
+    Group <- append(Group,
+                    rep(filter(allGroups, 
+                        accession == hm_groups[i])$name,
+                 times = length(temp_prots)),
+                 after = length(Group))
+
+    mat_labels <- append(mat_labels, temp_prots, after = length(mat_labels))
+  }
+}
+mat_labels <- mapIds(org.Xl.eg.db, 
+       keys = mat_labels, 
+       keytype = 'UNIPROT', 
+       column = 'SYMBOL')
+
+rownames(prot_hm) = mat_labels
+
+df <- tibble(run = rep(LETTERS[1:5], each = 5), 
+  minutes = as.character(rep(seq(0,80,20), times = 5))) %>%
+  unite(run,minutes, col = "x")
+
+col <- c(brewer.pal(6,"Blues")[2:6],
+         brewer.pal(6,"Greens")[2:6],
+         brewer.pal(6,"Oranges")[2:6],
+         brewer.pal(6,"Purples")[2:6],
+         brewer.pal(6,"Reds")[2:6])
+names(col) <- df$x
+
+column_ha <- HeatmapAnnotation(
+  sample = df$x,
+  col = list(sample = col)
+)
+
+# column_ha = HeatmapAnnotation(
+#   run = rep(LETTERS[1:5], each = 5), 
+#   minutes = as.character(rep(seq(0,80,20), times = 5)),
+#   col = list(
+#     run = c("A" = "#e41a1c", "B" = "#377eb8", "C" = "#4daf4a", "D" = "#984ea3", "E" = "#ff7f00"),
+#     minutes = c("0" = "#ffffcc", "20" = "#a1dab4", "40" = "#41b6c4", "60" = "#2c7fb8", "80" = "#253494")
+#   )
+# )
+row_ha = rowAnnotation(Group = Group)
+Heatmap(
+  log2(prot_hm),
+  name = "log2()",
+  #border = "black",
+  col = colorRamp2(seq(-3,3,0.01), rev(colorRampPalette(brewer.pal(11, 'RdBu'))(601))),
+  #col = wesanderson::wes_palette("Zissou1", 100, type = "continuous"),
+  top_annotation = column_ha,
+  right_annotation = row_ha,
+  cluster_columns = FALSE,
+  cluster_rows = FALSE,
+  show_column_names = FALSE,
+  row_names_side = "left",
+  row_names_gp = gpar(fontsize = 7)
+)
+
+df <- prot_hm %>% as_tibble %>% cbind(Group, .) %>% as_tibble() %>%
+  mutate(Prot = as.factor(rownames(prot_hm))) %>%
+  group_by(Group) %>% 
+  #mutate(Prot = 1:n()) %>% 
+  select(Group, Prot, everything()) %>%
+  ungroup() %>%
+  pivot_longer(3:ncol(.), names_to = "cell", values_to = "abundance")
+
+ggplot(df, aes(x = log2(abundance), y = as.factor(Prot), fill = Group)) + 
+  ggridges::geom_density_ridges(scale = 2, size = 0.1) +
+  facet_wrap(~Group, scales = "free_y", ncol = 1) +
+  xlim(-1,1) +
+  theme_minimal() +
+  theme(legend.position = "none",
+        axis.title.y=element_blank(),
+        #axis.text.y=element_blank(),
+        axis.ticks.y=element_blank(),
+        text = element_text(size = 8))
+
+df %>% group_by(Group, Prot) %>%
+  summarise(mean = mean(abundance),
+            sd = sd(abundance),
+            cv = sd / mean) %>%
+  #mutate(order = rank(sd)) %>%
+  ggplot(., aes(x = fct_reorder(Prot, rank(cv)), y = cv, fill = Group)) +
+  geom_bar(stat = "identity") + 
+  geom_hline(yintercept = median(allGroups$medCV), linetype = "dashed") +
+  facet_wrap(~Group, scales = "free_y", ncol = 1) +
+  theme_minimal() +
+  theme(legend.position = "none",
+         axis.title.y=element_blank(),
+        text = element_text(size = 8)) +
+  coord_flip()
+
+test_prots <- allGroups %>% filter(accession == hm_groups[7]) %>% pull(Uniprot) %>% unlist()
+test_cor_hm <- cor[test_prots, test_prots]
+test_prots <- mapIds(org.Xl.eg.db, 
+       keys = test_prots, 
+       keytype = 'UNIPROT', 
+       column = 'SYMBOL')
+rownames(test_cor_hm) = test_prots
+colnames(test_cor_hm) = test_prots
+
+Heatmap(test_cor_hm,
+        col = colorRamp2(seq(-1,1,0.01), rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201))),
+        show_row_names = TRUE,
+        show_column_names = TRUE,
+        show_heatmap_legend = FALSE,
+        show_column_dend = TRUE,
+        show_row_dend = TRUE)
+
+allGroups %>% filter(accession %in% hm_groups) %>% 
+  ggplot(., aes(x = fct_reorder(name, -medCV), y = medCV, fill = name)) + 
+  geom_bar(stat = "identity") +
+  geom_hline(yintercept = median(allGroups$medCV), linetype = "dashed") +
+  coord_flip() +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+allGroups %>% filter(accession %in% hm_groups) %>% 
+  ggplot(., aes(x = fct_reorder(name, medCors), y = medCors, fill = name)) + 
+  geom_bar(stat = "identity") +
+  geom_hline(yintercept = median(allGroups$medCors), linetype = "dashed") +
+  coord_flip() +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+# Histogram w/ variance groups of interest
+allGroups %>% 
+  ggplot(., aes(x = medCV)) +
+  geom_histogram(bins = 50) +
+  geom_histogram(data = filter(allGroups,
+                               accession %in% hm_groups),
+                 aes(x = medCV, fill = name),
+                 bins = 50) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+allGroups %>% 
+  ggplot(., aes(x = medCors)) +
+  geom_histogram(bins = 50) +
+  geom_histogram(data = filter(allGroups,
+                               accession %in% hm_groups),
+                 aes(x = medCors, fill = name),
+                 bins = 50) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+```
+
+# Gini Analysis
+
+Goal: do a positive and negatvie correlation Gini analysis. Since this analysis only works with positive values, and because
+
+```{r}
+library(diptest)
+
+for(i in seq_along(hm_groups)){
+  test_prots <- allGroups %>% 
+    filter(accession == hm_groups[i]) %>% 
+    pull(Uniprot) %>% 
+    unlist()
+
+  test_cor_hm <- cor[test_prots, test_prots]
+
+  print(dip.test(test_cor_hm[lower.tri(test_cor_hm)])$p.value)
+}
+
+test_prots <- allGroups %>% 
+  filter(accession == hm_groups[8]) %>% pull(Uniprot) %>% unlist()
+
+test_cor_hm <- cor[test_prots, test_prots]
+
+Heatmap(test_cor_hm,
+        col = colorRamp2(seq(-1,1,0.01), 
+                         rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201))),
+        show_row_names = FALSE,
+        show_column_names = FALSE,
+        show_heatmap_legend = FALSE,
+        show_column_dend = FALSE,
+        show_row_dend = FALSE)
+
+ggplot(allGroups, aes(x = medCV, y = sdCV)) + 
+  geom_point(alpha = 0.5, color = "grey") +
+  geom_point(data = filter(allGroups, accession %in% hm_groups),
+             color = "orange",
+             size = 2) +
+  geom_text_repel(data = filter(allGroups, accession %in% hm_groups),
+                  aes(label = name),
+                  force = 10,
+                  box.padding = 1) +
+  theme_minimal()
+
+#### KEGG
+  data <- kegg_genes %>% filter(kegg_pathway %in% pathway) %>%
+    left_join(joined_wat_results, by = 'ext_gene') %>%
+    filter(lrt_qval < 0.05)
+
+  genes <- as.vector(log2(data$fc))
+  names(genes) <- data$entrez_id
+  pv.out <- pathview(genes, 
+                     pathway.id = pathway, 
+                     species = 'mmu', 
+                     same.layer = F,
+                     low = list(gene = 'blue'),
+                     mid = list(gene = 'transparent'),
+                     high = list(gene = 'red'),
+                     limit = list(gene = c(-2,2))
+                     )
+
+```
+
+```{r figure_2, echo=FALSE, fig.height=9, fig.width=8.5, message=FALSE, warning=FALSE, eval=TRUE, fig.cap=fig_legend_2}
 p1 <- ggplot(allGroups, aes(x = normVar, fill = interval)) + 
@@
                             "GO:0006096","GO:0015934","GO:0019843",
-                            "00051","GO:0042555","04012")
+                            "00051","GO:0042555","04012","GO:0006098")
 
@@
 
-p2 <- ggplot(filter(allGroups, n > 5), aes(x = medCors, y = medCV, color = interval)) +
-  geom_point(data = filter(allGroups, interval == "mid", n > 4),
+p2 <- ggplot(filter(allGroups, n > 3), aes(x = medCors, y = medCV, color = interval, shape = type)) +
+  geom_point(data = filter(allGroups, interval == "mid", n > 3),
              size = 2, alpha = 0.5, stroke = 0) +
-  geom_point(data = filter(allGroups, interval != "mid", n > 4),
+  geom_point(data = filter(allGroups, interval != "mid", n > 3),
              size = 3, alpha = 0.75, stroke = 0) +
-  geom_label_repel(data = filter(allGroups, accession %in% interesting_groups),
+  geom_label_repel(
+    data = filter(allGroups, accession %in% interesting_groups),
+    #data = filter(allGroups, medCors > 0.15, interval != "mid", n > 4),
     aes(label = name),
@@
     force = 5,
-                  box.padding = 1) +
+    box.padding = 1
+  ) +
   #coord_cartesian(ylim=c(-10,10)) +
@@
         axis.ticks.x=element_blank()) +
-  xlim(-0.15,0.5)
+  xlim(-0.15,0.5) +
+  ylab("Density")
 
@@
         axis.ticks.y=element_blank()) +
-  xlim(0.08,0.32)
+  xlim(0.08,0.32) +
+  ylab("Density")
+
+plots <- 5:8
+for(i in 1:length(plots)){
+  assign(paste0("p",plots[i]), cor_rug_plot(interesting_groups_low[i], "#3288bd"))
+}
+
+plots <- 9:12
+for(i in 1:length(plots)){
+  assign(paste0("p",plots[i]), cor_rug_plot(interesting_groups_high[i], "#d53e4f"))
+}
+
+left <- p5 + p6 + p7 + p8 + plot_layout(nrow = 1)
+
+right <- p9 + p10 + p11 + p12 + plot_layout(nrow = 1)
+
+bottom <- wrap_plots(left, right) + 
+  plot_annotation(tag_levels = "A") + 
+  plot_layout(nrow = 2)
 
@@
 
-p3 + plot_spacer() + p2 + (p4 + coord_flip()) +
+mid <- p3 + plot_spacer() + p2 + (p4 + coord_flip()) +
   plot_layout(widths = c(4,1),
               heights = c(1,4))
+p1 / mid / bottom +
+  plot_layout(heights = c(0.5,4,3)) +
+  plot_annotation(tag_levels = "A")
 
-fig_legend_2 <- c("A) Histogram of the normalized variation values across all of the GO terms and KEGG pathways.
+fig_legend_2 <- c("Tradeoff between expression variation and co-expression in low variation modules: A) Histogram of the normalized variation values across all of the GO terms and KEGG pathways.
 The bottom and top ~15% are colored blue and red respectively.
@@
 Both the top and bottom 15% of normalized variation groups had peaks of high levels of mean correlation coefficients.
-C) Distributions of the mean coefficient of variation across all groups of proteins.
+C) Pairwise plot between the mean correlation and mean coefficient of variation within each GO term and KEGG pathway with color coding from A.
+Interestingly, both the top and bottom 10% of normalized variation have groups of proteins with high levels of correlation, despite high correlation adding to total variance.
+In the lowest 10% of normalized variation groups there appears to be a negative correlation between mean correlation and variation.
+D) Distributions of the mean coefficient of variation across all groups of proteins.
 The top and bottom ~15% of normalized variation groups were largely separated.
-D) Pairwise plot between the mean correlation and mean coefficient of variation within each GO term and KEGG pathway with color coding from A.
-Interestingly, both the top and bottom 10% of normalized variation have groups of proteins with high levels of correlation, despite high correlation adding to total variance.
-In the lowest 10% of normalized variation groups there appears to be a negative correlation between mean correlation and variation.")
+E:L) Density plots of correlation coefficients select protein groups color coded by low (blue) and high (red) normalized protein variance, with the distribution of all measured correlation coefficients colored in grey (rug plot to show the measured correlations coefficients shown on plot).
+Below each density plot is a box plot.")
 
@@
 
-\newpage
-
-```{r figure_3, echo=FALSE, fig.height=4, fig.width=7, message=FALSE, warning=FALSE, eval=TRUE, fig.cap=fig_legend_3}
-plots <- 1:4
-for(i in 1:length(plots)){
-  assign(paste0("p",plots[i]), cor_rug_plot(interesting_groups_low[i], "#3288bd"))
-}
-
-plots <- 5:8
-for(i in 1:length(plots)){
-  assign(paste0("p",plots[i]), cor_rug_plot(interesting_groups_high[i], "#d53e4f"))
-}
-
-left <- p1 + p2 + p3 + p4 + plot_layout(nrow = 1)
-
-right <- p5 + p6 + p7 + p8 + plot_layout(nrow = 1)
-
-wrap_plots(left, right) + 
-  plot_annotation(tag_levels = "A") + 
-  plot_layout(nrow = 2)
+```{r figure_3, echo=FALSE, fig.height=4, fig.width=7, message=FALSE, warning=FALSE, eval=FALSE, fig.cap=fig_legend_3}
 
@@
 
-fig_legend_4 <- c("A) A representative heat-map of one of the low normalized variation groups with high levels of correlated proteins (Large ribosomal subunit).
+fig_legend_4 <- c("Co-expression and low variance leads to better control of assembled heteromeric protein complexes: A) A representative heat-map of one of the low normalized variation groups with high levels of correlated proteins (Large ribosomal subunit).
 B) A simple model showing the percent of potentially assembled subunits of an idealized heteromeric 10 subunit complex.
@@
 
+```{r}
+pca = prcomp(log2(protImpute[,2:ncol(protImpute)]))
+
+pcaDF = tibble(cell = rownames(pca$rotation), PC1 = pca$rotation[,1], PC2 = pca$rotation[,2])
+pcaDF = pcaDF %>% separate(cell, into = c('run','label',"time"), remove = F)
+ggplot(pcaDF, aes(x = PC1, y = PC2, colour = time)) + geom_point(size = 5) +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+```
+
 # Methods
@@
 Single eggs were lysed mechanically by pipetting the egg in 100$\mu$L of lysis buffer (100 mM NaCl, 25 mM Tris pH 8.2, Complete EDTA- free protease inhibitor cocktail (Sigma).
-The lysate was then placed in a 400 uLnatural polyethylene micro-centrifuge tube (E&K
+The lysate was then placed in a 400 uL natural polyethylene micro-centrifuge tube (E&K
 Scientific #485050) and spun at 15,000 g in a right angle centrifuge (Beckman Microfuge E) at 4°C for 5 min.
